### 2020-03-30 docs/draft_chapter.Rmd
start

---
title: "Topic 6.2: Open Data"
author: "Samuel Langton and Reka Solymosi"
date: "1 June 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

- Topic definition
- Chapter structure

## Background

- What is open data?
- What types of organisations release open data?
  - Public sector
  - Private sector
- What are the strengths/limitations of open data?

## Surveying environmental features using Open Street Map

- Explain OSM
- Note important technical details including key, value and elements
- Explain APIs
- Overpass queries and overpass-turbo
- But better: use OSM API to download some features in R
- Note the timeout error
- Map features
- Download crime data to compare with features (eg violence ~ bars)

## Future of open data
- Open topics in open data
- Further applications to crime and place research

## Conclusion

- Re-cap
- Wrap-up

### 2020-03-30 scripts/osm_explore.R

@@ -1,0 +1,28 @@
+library(osmdata)
+library(ggplot2)
+library(sf)
+
+# All queries begin with a bounding box specification i.e. the study region.
+# This can be obtained manually, which requires some existing knowledge about
+# an area using the latitude and longitude coordinates, but it is generally
+# easier to use a search term.
+
+bb <- getbb(place_name = "manchester united kingdom", format_out = "sf_polygon") 
+
+ggplot(data = bb) +
+  geom_sf()
+
+
+p <- opq(bbox = bb) %>%                                  # select bounding box
+  add_osm_feature(key = 'amenity', value = 'pub') %>%    # select features
+  osmdata_sf() %>%                                       # specify class
+  trim_osmdata (bb)                                            
+
+class(p)
+p.sf <- p$osm_points
+
+p.clip.sf <- st_intersection(bb, p.sf)
+
+ggplot() +
+  geom_sf(data = bb) +
+  geom_sf(data = p.df)

### 2020-03-30 scripts/osm_explore.R
bike theft example

@@ -1,6 +1,9 @@
 library(osmdata)
 library(ggplot2)
 library(sf)
+library(readr)
+library(dplyr)
+library(tidyr)
 
 # All queries begin with a bounding box specification i.e. the study region.
 # This can be obtained manually, which requires some existing knowledge about
@@ -9,20 +12,38 @@ library(sf)
 
 bb <- getbb(place_name = "manchester united kingdom", format_out = "sf_polygon") 
 
-ggplot(data = bb) +
-  geom_sf()
+p <- opq(bbox = bb) %>%                                           # select bounding box
+  add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
+  osmdata_sf()                                                    # specify class (sf or sp)
 
-
-p <- opq(bbox = bb) %>%                                  # select bounding box
-  add_osm_feature(key = 'amenity', value = 'pub') %>%    # select features
-  osmdata_sf() %>%                                       # specify class
-  trim_osmdata (bb)                                            
-
-class(p)
-p.sf <- p$osm_points
-
-p.clip.sf <- st_intersection(bb, p.sf)
+p
+p.sf <- p$osm_points # extract point only
 
 ggplot() +
   geom_sf(data = bb) +
-  geom_sf(data = p.df)
+  geom_sf(data = p.sf)
+
+p.sf  <- st_transform(p.sf, 27700)
+bb.sf <- st_transform(bb, 27700)
+p.clip.sf <- st_intersection(bb.sf, p.sf)
+
+ggplot() +
+  geom_sf(data = bb) +
+  geom_sf(data = p.clip.sf, size = 0.5)
+
+# crime data
+crime.df <- read_csv("data/2019-01-greater-manchester-street.csv")
+
+bike.crime.sf <- crime.df %>% 
+  filter(`Crime type` == "Bicycle theft") %>% 
+  drop_na(Longitude, Latitude) %>% 
+  st_as_sf(coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
+  st_transform(27700) %>% 
+  st_intersection(bb.sf)
+
+# plot
+ggplot() +
+  geom_sf(data = bb) +
+  geom_sf(data = p.clip.sf, size = 0.8) +
+  geom_sf(data = bike.crime.sf, size = 0.8, col = "red", alpha = 0.6)
+

### 2020-03-31 scripts/osm_explore.R
More tests

@@ -10,7 +10,7 @@ library(tidyr)
 # an area using the latitude and longitude coordinates, but it is generally
 # easier to use a search term.
 
-bb <- getbb(place_name = "manchester united kingdom", format_out = "sf_polygon") 
+bb <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon") 
 
 p <- opq(bbox = bb) %>%                                           # select bounding box
   add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
@@ -27,12 +27,12 @@ p.sf  <- st_transform(p.sf, 27700)
 bb.sf <- st_transform(bb, 27700)
 p.clip.sf <- st_intersection(bb.sf, p.sf)
 
-ggplot() +
-  geom_sf(data = bb) +
-  geom_sf(data = p.clip.sf, size = 0.5)
+# ggplot() +
+#   geom_sf(data = bb) +
+#   geom_sf(data = p.clip.sf, size = 0.5)
 
 # crime data
-crime.df <- read_csv("data/2019-01-greater-manchester-street.csv")
+crime.df <- read_csv("data/2020-01-metropolitan-street.csv")
 
 bike.crime.sf <- crime.df %>% 
   filter(`Crime type` == "Bicycle theft") %>% 
@@ -45,5 +45,19 @@ bike.crime.sf <- crime.df %>%
 ggplot() +
   geom_sf(data = bb) +
   geom_sf(data = p.clip.sf, size = 0.8) +
-  geom_sf(data = bike.crime.sf, size = 0.8, col = "red", alpha = 0.6)
+  geom_sf(data = bike.crime.sf, size = 0.8, col = "red")
 
+# polygons
+bb <- getbb(place_name = "coventry united kingdom", format_out = "sf_polygon") 
+
+p <- opq(bbox = bb) %>%                                        
+  add_osm_feature(key = 'public_transport', value = 'station') %>% 
+  osmdata_sf() %>% 
+  trim_osmdata(bb_poly = bb)
+
+p
+p.sf <- p$osm_polygons
+
+ggplot() +
+  geom_sf(data = bb) +
+  geom_sf(data = p.sf, fill = "black")

### 2020-04-08 draft_chapter.Rmd
bike crime example

---
title: "Topic 6.2: Open Data"
author: "Samuel Langton and Reka Solymosi"
date: "1 June 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(osmdata)
library(ggplot2)
library(sf)
library(readr)
library(dplyr)
library(tidyr)
```

## Introduction

- Topic definition
- Chapter structure

## Background

- What is open data?
- What types of organisations release open data?
  - Public sector
  - Private sector
- What are the strengths/limitations of open data?

## Surveying environmental features using Open Street Map

### What is Open Street Map?

### Downloading Open Street Map Data
- APIs
- Overpass queries and overpass-turbo

### Using Open Street Map data in R

osmdata R pacakge.

All queries begin with a bounding box specification to define the study region. This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, but it is generally easier to use a search term. Here, we select Greater London in the United Kingdom using the `getbb()` function, specifying that we want the content as a simple features (sf) object.

```{r}
bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon")
```

We now have our study region defined as the administrative boundaries of Greater London.

```{r}
ggplot(data = bb_sf) +
  geom_sf()
```

Now we have our study region, we can scrape data from the OSM API using the `opq()` function. This allows you to build an Overpass query, outlined in the previous section, from within the R environment. We specify the bounding box defined earlier, and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define what we want to pull from the API. We noted earlier, features recorded on OSM with various tags: keys and values. Here, we specify that we want amenities (the key) defined as bicycle parking (the value). The next function `osmdata_sf()` ensures that the resulting object is a simple features class for easy plotting with `ggplot`. We trim the features pulled from the API based on the Greater London boundaries, otherwise we would end up with information which extends beyond the study region.

```{r}
bikes_sf <- opq(bbox = bb_sf) %>%                                 # select bounding box
  add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
  osmdata_sf() %>%                                                # specify class
  trim_osmdata(bb_poly = bb_sf)                                   # trim to bounding box
```

The resulting object `bikes_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the Console.

```{r}
bikes_sf
```

This confirms details like the bounding box, but also provides information on the simple features collected from the query. As one might expect, most information relating to bicyle parking has been recorded using points (i.e. two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing. We also have around one hundred polygons. For now, let's extract the point information only and then transform the CRS to the BNG.

```{r}
bikes_points_sf <- bikes_sf$osm_points 
```

We can then plot these points over our original boundaries of Greater London, once we have projected it as the BNG too. We reduce the default size of the point to ensure that we avoid too much overlap between bicycle parking locations.

```{r}
ggplot() +
  geom_sf(data = bb_sf) +
  geom_sf(data = bikes_points_sf, size = 0.3)
```

As we can see, most bicycle parking spaces are clustered around the city centre, especially just north of the river Thames. It is also possible to make out key roads flowing in and out of the city centre, which contain bicycle parking all along the street.

Using open police recorded crime data we can then plot actual incidences of bicycle theft to explore whether there is a spatial relationship between bike theft and parking spots in Greater London. For this example, we just use crime recorded as occurring in January 2020. First, let's load in the data as it downloaded raw from https://data.police.uk/data/.

```{r}
crime.df <- read_csv("data/2020-01-metropolitan-street.csv")
```

We then need conduct a bit of prliminary data handling: filter crimes which were tagged as bicycle theft, convert the latitude and longitude columns to coordinates with simple features, assign a CRS and then clip the points by our study region. Finally, we transform the CRS to the British National Grid so we can accurately plot these points over our bicycle parking locations.

```{r}
bike.crime.sf <- crime.df %>% 
  filter(`Crime type` == "Bicycle theft") %>% 
  drop_na(Longitude, Latitude) %>% 
  st_as_sf(coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
  st_intersection(bb_sf)
```

Then, finally plot everything together, with a few adjustments to the appearance.

```{r}
ggplot() +
  geom_sf(data = bb_sf) +
  geom_sf(data = bikes_points_sf, aes(colour = "bike_park"), size = 0.4, alpha = 0.5) +
  geom_sf(data = bike.crime.sf, aes(colour = "bike_crime"), size = 0.3, alpha = 0.5) +
  scale_colour_manual(name = NULL, values = c(bike_park = "black", bike_crime = "red")) +
  theme(legend.position = "bottom")
```

## Future of open data

- Open topics in open data
- Further applications to crime and place research

## Conclusion

- Re-cap
- Wrap-up

### 2020-04-08 scripts/osm_explore.R

@@ -8,7 +8,7 @@ library(tidyr)
 # All queries begin with a bounding box specification i.e. the study region.
 # This can be obtained manually, which requires some existing knowledge about
 # an area using the latitude and longitude coordinates, but it is generally
-# easier to use a search term.
+# easier to use a search term. For instance
 
 bb <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon") 
 

### 2020-04-08 scripts/osm_explore.R
Chapter structure

@@ -10,10 +10,10 @@ library(tidyr)
 # an area using the latitude and longitude coordinates, but it is generally
 # easier to use a search term. For instance
 
-bb <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon") 
+bb <- getbb(place_name = "washington dc", format_out = "sf_polygon") 
 
 p <- opq(bbox = bb) %>%                                           # select bounding box
-  add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
+  add_osm_feature(key = 'amenity', value = 'bar') %>% # select features
   osmdata_sf()                                                    # specify class (sf or sp)
 
 p

### 2020-04-09 draft_chapter.Rmd
walkthru prep

@@
 title: "Topic 6.2: Open Data"
-author: "Samuel Langton and Reka Solymosi"
-date: "1 June 2020"
+author: "Samuel Langton (MMU) & Reka Solymosi (University of Manchester)"
+date: "June 2020"
 output: pdf_document
@@
 
-- Topic definition
-- Chapter structure
+- Topic definition.
+- Advantages and exciting prospects in open data.
+- Chapter structure outline.
+- Prerequisites (reference _Geocomputation with R_).
 
@@
   - Private sector
-- What are the strengths/limitations of open data?
+- Why is it needed?
+- What are the strengths?
+- What are the limitations?
+- Substantive examples in criminology (focus on crime and place research).
+- Conclude with: a key one is Open Street Map.
 
-## Surveying environmental features using Open Street Map
+## Open Street Map
 
@@
 
+- Brief history.
+- Motivations for and advantages of an open web mapping platform.
+- Primary features (keys, values, elements).
+- Contributing to the data yourself.
+
 ### Downloading Open Street Map Data
-- APIs
-- Overpass queries and overpass-turbo
+
+- Define and explain APIs.
+- For OSM, describe overpass queries and overpass-turbo.
+- Query example (brief).
 
@@
 
-osmdata R pacakge.
+- Made easy through the `osmdata` package in R.
+- More straightforward than overpass-turbo.
+- Compatible with both `sp` and `sf` classes of spatial data.
+- We focus on `sf` as its compatible with the `tidyverse`.
+- Outline the key documentation and references.
+
+## Walk-through
+
+First, ensure that you have the relevant packages installed in R.
+Although the main package used in this demonstration is `osmdata`, for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation.
+If you don't have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
+
+```{r, eval = F}
+library(osmdata)
+library(ggplot2)
+library(sf)
+library(readr)
+library(dplyr)
+library(tidyr)
+```
 
@@
 This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, but it is generally easier to use a search term.
-Here, we select Greater London in the United Kingdom using the `getbb()` function, specifying that we want the content as a simple features (sf) object.
+Here, we select Greater London in the United Kingdom using the `getbb()` function, specifying that we want the content as a simple features (sf) polygon.
 
@@
 We now have our study region defined as the administrative boundaries of Greater London.
+This can be visualised using `ggplot2` and the _simple features_ geometry `geom_sf()` available with `sf`.
+Note that by default, the Coordinate Reference System (CRS) is the World Geodetic System 84.
 
@@
 
-Now we have our study region, we can scrape data from the OSM API using the `opq()` function.
+Now we have our study region, we can scrape data from the OSM API using the `opq()` function, which is short for 'Overpass query'.
 This allows you to build an Overpass query, outlined in the previous section, from within the R environment.
-We specify the bounding box defined earlier, and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define what we want to pull from the API.
-We noted earlier, features recorded on OSM with various tags: keys and values.
+We specify the bounding box object which is our study area, and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define what we want to pull from the API.
+As we noted earlier, features in OSM have are defined through keys and values.
 Here, we specify that we want amenities (the key) defined as bicycle parking (the value).
-The next function `osmdata_sf()` ensures that the resulting object is a simple features class for easy plotting with `ggplot`.
-We trim the features pulled from the API based on the Greater London boundaries, otherwise we would end up with information which extends beyond the study region.
+This query is then piped through to `osmdata_sf()` which ensures that the resulting object is a _simple features_ class for easy plotting with `ggplot2`.
+We trim the features pulled from the API using `tim_osmdata()` to ensure that everything stays within the boundaries of our study region.
 
@@
   osmdata_sf() %>%                                                # specify class
-  trim_osmdata(bb_poly = bb_sf)                                   # trim to bounding box
+  trim_osmdata(bb_poly = bb_sf)                                   # trim to region
 ```
@@
 This confirms details like the bounding box, but also provides information on the simple features collected from the query.
-As one might expect, most information relating to bicyle parking has been recorded using points (i.e.
+As one might expect, most information relating to bicycle parking has been recorded using points (i.e.
 two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing.
@@
 
-We can then plot these points over our original boundaries of Greater London, once we have projected it as the BNG too.
-We reduce the default size of the point to ensure that we avoid too much overlap between bicycle parking locations.
+We can then plot these points over our original boundaries of Greater London. We reduce the default size of the point to ensure that we avoid too much overlap between bicycle parking locations.
 
@@
 
-We then need conduct a bit of prliminary data handling: filter crimes which were tagged as bicycle theft, convert the latitude and longitude columns to coordinates with simple features, assign a CRS and then clip the points by our study region.
-Finally, we transform the CRS to the British National Grid so we can accurately plot these points over our bicycle parking locations.
+We then need to conduct a bit of preliminary data handling: filter crimes which were tagged as bicycle theft, convert the latitude and longitude columns to coordinates with _simple features_, state the WGS 84 CRS and then clip the points by our study region.
 
@@
 
-Then, finally plot everything together, with a few adjustments to the appearance.
+To demonstrate the data in its entirety, we can plot the Greater London boundaries, overlayed with the bicycle parking space locations, and open crime data about bicycle thefts.
+It is worth clarifying that open police recorded crime data in England and Wales is spatially anonymised by a process of snapping points to a pre-defined grid (see Tompson et al., 2015).
+For that reason, many of these points overlap, and thus a degree of transparency is used for the points.
 
@@
 
-- Open topics in open data
-- Further applications to crime and place research
+- Threats to its sustainability (e.g. licence expiry).
+- Prospects in crime of place research.
+- Examples of cool projects (e.g. Colouring London).
+- Suggestions for new avenues which can expand the field.
 
@@
 
-- Re-cap
-- Wrap-up
+- Re-cap on what we've covered.
+- Wrap-up the key points.
+
+## References
+
+Tompson, L., Johnson, S., Ashby, M., Perkins, C., & Edwards, P.
+(2015).
+UK open source crime data: accuracy and possibilities for research.
+_Cartography and geographic information science_, 42(2), 97-111.

### 2020-04-15 draft_chapter.Rmd
added some literature

@@
   chunk_output_type: console
+bibliography: refs.bib
 ---
@@
 
-## Introduction
+# Introduction
 
-- Topic definition.
-- Advantages and exciting prospects in open data.
-- Chapter structure outline.
+Access to most data sets have traditionally been restricted to those who have the means to collect these themselves or to pay for, or otherwise gain access to them.
+Even when access is granted, this is often with conditions that circumscribe how the data can be used through licensing or policy [@kitchin2014data].
+This lead to unequal access to data and the subsequent insight that can be attained from them.
+
+Open data are a response to these restrictions, seeking to open up data without restrictions, to broaden access and participation in research deriving insight from all sorts of data sets, removing the need to ask permission or negotiate access [@manovich2011trending].
+Not only that, but using open data can lead to all sorts of novel insight within the domain of criminology and crime analysis specifically by tapping into constructs and processes difficult to capture through surveys, interviews, and other traditional measures [@solymosi2018role].
+Therefore, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the social world around them have the skills and know-how to access and accurately interpret, assess and analyse such data sets.
+
+This chapter aims to lay the groundwork for developing such skills, providing both a framework to approach and meainingfully interpret open data regarding understanding its origin and any limitations associated with that, and a practical hands-on guide to accessing, wrangling, and analysting open data to draw conclusions about crime and place.
+We achieve this by first giving a background on open data, discussing the key types of open data out there, critically assessing the strengths and limiations of open data, and building a how-to guide checklist for researchers to follow when approaching an open data source.
+Finally, we work through an example that shows how to access, wrangle, link, and interpret open data from various sources, providing a template to follow.
+
 - Prerequisites (reference _Geocomputation with R_).
 
-## Background
+# Background
 
-- What is open data?
-- What types of organisations release open data?
-  - Public sector
-  - Private sector
-- Why is it needed?
-- What are the strengths?
-- What are the limitations?
-- Substantive examples in criminology (focus on crime and place research).
-- Conclude with: a key one is Open Street Map.
+## What is open data?
+
+What do we mean when we refer to **open data**?
+According to @dietrich2009open in the Open Data Handbook, open data is "data that can be freely used, re-used and redistributed by anyone - subject only, at most, to the requirement to attribute and sharealike."
+
+They summarise the content of the Open Definition document published by the @knowledge2016open into three key tenants of open data:
+
+- **Availability and Access:** the data must be available in the publica domain, accessible at no more than a reasonable reproduction cost.
+It should be in a machine readable form that is easily accessed and modified.
+For example, it should be downloadable as a spreadsheet, or json file, rather than presented as a summary table in a PDF document.
+- **Re-use and Redistribution:** the data must be provided under terms that allow free use, including modification, separation, or compilation of the data.
+permit re-use and redistribution including the intermixing with other datasets.
+- **Universal Participation:** everyone must be able to use, re-use and redistribute the data and its derivatives without restriction (eg: 'non-commercial' restrictions or use in education purposes only)
+
+---
+In short, open data should be data that is available in a useful format for analysis that can be used by anyone for any purpose. 
+---
+
+Open data practices vary within and between countires.
+For example, the United States of America have a history of making public sector data sets available, while the United Kingdom is much more strict in releasing their data under a licence or for a fee [@kitchin2014data].
+The Global Open Data Index [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by countries, providing a rank which shows how well governments across the world follow open data practices.
+
+## What are types of open data?
+
+Here we formulate a typology of open data sets based on their origin.
+
+### Public sector (including publicly funded research)
+
+Much of the open data movement has focused on data generated by state agencies (often termed public sector information) or publicly funded research, given that these have been funded by the public purse for the public's benefit  [@kitchin2014data].
+For example, the Global Open Data Index introduced earlier focuses specifically on governmental bodies' data publishers, identifying data gaps in these organisations, and promoting them to think about how to make their data more useable and eventually more impactful [@globopendata].
+This is important as open public sector data can be ....
+
+<!-- MORE HERE basically the paragraph above should really say what is public sector data, why its important (ie what can we do with it as researhcers) -->
+
+Regarding opening up research data, the argument focuses not only on the need to make open data where its collection was funded by public funds, but also on the need for transparency and reproducibility.
+Replication is a hallmark of such open science practice, and opening up data sets used in publications in criminology and wider social sciences contributes to foster a culture of replication [@pridemore2018replication].
+Open data and replications fit into the wider discourse in social sciences around open data, open review and open dialogue, which may even become (with time) policy requirements regardless of the research funding source [@vuong2017open].
+There are repositories where open research data is available (osf.io, uk data service, others??).
+
+<!-- Somehow I want to also say "keep in mind open research practices when collecting your own data, although we dont cover how to actually release and make open your own data in this chapter, read more about this somewhere...
+idk?
+-->
+
+### Private sector
+
+A more limited focus of open data activists lays on  opening up data generated by private industry, which might have more proprietary value to its creators [@kitchin2014data].
+While datasets are released by private sector companies, they often are only a subset selected by the organisation.
+However they might still provide valuable insight.
+Indeed many interesting papers have been written using the data released by large private sector organisations such as Google [@stephens2017everybody], Instagram [@reece2017instagram], Uber [@davidson2017interactive] and many more.
+In another chapter in this book you will learn to use the free Twitter API to do just the same (see Nick and Dan's chapter link to that).
+
+### Open crowdsourced data
+
+Finally, there is a specific group of open data sources that collate data collectively generated by large groups of individuals who do not specifically belong to any organisation, but instead work together on a collaborative project.
+These are crowdsourced data sets.
+Examples include Wikipedia (www.wikipedia.org), an online encyclopedia where anyone can contribute, and Flickr (www.flickr.com) an online photo gallery where anyone can upload and tag their photos.
+Often the organisations who maintain and monitor these data collection activities are charities or non govermental organisations that operate not for profit, but instead to provide some sort of social good.
+An example is the online problem reportin platform fixmystreet, where people report instances of graffiti or other environmental complaints, which can offer insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
+Another chapter in this book (REF CHAPTER WITH DAVID) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analysing open data of this type.
+
+## What are strengths and limitations
+
+Each one of the above categories might have its own strengths and limitations, and it is very important to consdier the source of the data you are working with, and what might mean for what and who is represented or excluded from these data.
+Here we will provide a general overview, and in the next section some guidance on how to make sure you take all necessary precautions when analysing open data of any kind.
+
+Open data is hailed for its ease of use afforded by its availability and access.
+This strength is emphasised throughout the chapter.
+However another advantage and motivation for using open data is its potential to address many limitations associated with traditional surveying methods, such as social desirability bias, or issues associated with memory and recall [@mayer2013big].
+
+However this also creates a limitation for such data, in that because it is already available, the researcher has no say into what gets collected, and instead must make do with what is available.
+There are limits as to what questions researchers can ask of a data set, and what interpretations are appropriate [@boyd2012critical].
+
+Despite these limitation, the social sciences must embrace these new forms of data that, although messy, biased and noisy, have the potential to describe social phenomena better than well-organized small surveys or even national censuses [@savage2007coming].
+
+## What do researchers need to know to make good use of open data
+
+in the case of open data, there is no need to ask for permissions or negotiate data access in order to use it for research [@manovich2011trending].
+In this sense, the great advantage of this data is that it is already out there.
+To collect it, all that is needed is a way to be able to interpret what this data means (for example by applying a framework for its analysis), and some skill in data scraping, wrangling and cleaning, in order to be able to transform it into a usable format for research [@boyd2012critical].
+
+necessary to ask critical questions about what all this data means, who gets access to what data, how data analysis is deployed, and to what ends [@boyd2012critical].
+
+be careful in interpreting people's communication online as authentic [@manovich2011trending].
+
+The key concerns about these dara relate to issues about sampling, and ultimately the generalisability of any findings that can be drawn from them.
+In the case of crowdsourced data, where the sample of the population who contributes to such data is self-selected, giving way for people more motivated to speak about the issue, which means the data are likely to contain inherent bias [@longley2012geodemographics].
+Specifically it tends to be men, between ages of 20-50, with a college or university degree who are most likely contributors [@budhathoki2010participants; @haklay2010good].
+
+Looking into what contextual factors influence participation in Open Street Map, @mashhadi2013putting find that, socio-economic factors such as population density, dynamic population, distance from the centre and poverty all play an important role.
+These are important to keep in mind when reporting findings based on analysis of such data.
+
+even with this limitation, the inclusion of crowdsourced data which are both up-to-date and specific to the problem at hand can still provide new insight in addition to the knowledge from established sources of data collected in traditional methods [@birkin2011calibration].
+All of this helps us to understand the relations between the events and the occurrences that come together in unique places and provides a framework for understanding how places change over different time periods [@lonlongley2012geodemographics].
 
@@
 To demonstrate the data in its entirety, we can plot the Greater London boundaries, overlayed with the bicycle parking space locations, and open crime data about bicycle thefts.
-It is worth clarifying that open police recorded crime data in England and Wales is spatially anonymised by a process of snapping points to a pre-defined grid (see Tompson et al., 2015).
+It is worth clarifying that open police recorded crime data in England and Wales is spatially anonymised by a process of snapping points to a pre-defined grid (see @tompson2015uk).
 For that reason, many of these points overlap, and thus a degree of transparency is used for the points.
@@
   theme(legend.position = "bottom")
+```
+
+### Bus stop example
+
+```{r}
+bus_sf <- opq(bbox = bb_sf) %>%                                 # select bounding box
+  add_osm_feature(key = 'highway', value = 'bus_stop') %>%      # select features
+  osmdata_sf() %>%                                              # specify class
+  trim_osmdata(bb_poly = bb_sf)  
 ```
@@
 
-## References
-
-Tompson, L., Johnson, S., Ashby, M., Perkins, C., & Edwards, P.
-(2015).
-UK open source crime data: accuracy and possibilities for research.
-_Cartography and geographic information science_, 42(2), 97-111.
+# References

### 2020-04-22 draft_chapter.Rmd
finally knit stupid pdf

@@
 date: "June 2020"
-output: pdf_document
+output:
+  pdf_document:
+    latex_engine: xelatex
 editor_options: 
@@
 knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
+library(jsonlite)
 library(osmdata)
@@
 permit re-use and redistribution including the intermixing with other datasets.
-- **Universal Participation:** everyone must be able to use, re-use and redistribute the data and its derivatives without restriction (eg: 'non-commercial' restrictions or use in education purposes only)
+- **Universal Participation:** everyone must be able to use, re-use and redistribute the data and its derivatives without restriction (eg: 'non-commercia' restrictions or use in education purposes only)
 
@@
 Each one of the above categories might have its own strengths and limitations, and it is very important to consdier the source of the data you are working with, and what might mean for what and who is represented or excluded from these data.
-Here we will provide a general overview, and in the next section some guidance on how to make sure you take all necessary precautions when analysing open data of any kind.
+Here we will provide a general overview, and some guidance on how to make sure you take all necessary precautions when analysing open data of any kind.
 
-Open data is hailed for its ease of use afforded by its availability and access.
-This strength is emphasised throughout the chapter.
-However another advantage and motivation for using open data is its potential to address many limitations associated with traditional surveying methods, such as social desirability bias, or issues associated with memory and recall [@mayer2013big].
-
-However this also creates a limitation for such data, in that because it is already available, the researcher has no say into what gets collected, and instead must make do with what is available.
-There are limits as to what questions researchers can ask of a data set, and what interpretations are appropriate [@boyd2012critical].
-
-Despite these limitation, the social sciences must embrace these new forms of data that, although messy, biased and noisy, have the potential to describe social phenomena better than well-organized small surveys or even national censuses [@savage2007coming].
-
-## What do researchers need to know to make good use of open data
-
-in the case of open data, there is no need to ask for permissions or negotiate data access in order to use it for research [@manovich2011trending].
+Open data is hailed for its ease of use afforded by its availability and access; there is no need to ask for permissions or negotiate data access in order to use it for research [@manovich2011trending].
 In this sense, the great advantage of this data is that it is already out there.
@@
 
-necessary to ask critical questions about what all this data means, who gets access to what data, how data analysis is deployed, and to what ends [@boyd2012critical].
+But besides being easy to access, there are further advantages to using such data.
+One specific motivation for using open data comes from its potential to address many limitations associated with traditional surveying methods, such as social desirability bias, or issues associated with memory and recall [@mayer2013big].
+By generating data often as a byproduct of other activities, we might gain more honest insight into people's everyday lives and associated social processes [@solymosi2018role].
+At the same time, there are new threats to validity that may emerge in these sorts of data, and researchers should be careful in interpreting people's communication online as authentic [@manovich2011trending].
 
-be careful in interpreting people's communication online as authentic [@manovich2011trending].
-
-The key concerns about these dara relate to issues about sampling, and ultimately the generalisability of any findings that can be drawn from them.
+A related, and key concern about these dara relates to issues about sampling, and ultimately the generalisability of any findings that can be drawn from them.
 In the case of crowdsourced data, where the sample of the population who contributes to such data is self-selected, giving way for people more motivated to speak about the issue, which means the data are likely to contain inherent bias [@longley2012geodemographics].
@@
 
-Looking into what contextual factors influence participation in Open Street Map, @mashhadi2013putting find that, socio-economic factors such as population density, dynamic population, distance from the centre and poverty all play an important role.
+Looking into what contextual factors influence participation in Open Street Map, @mashhadi2013putting found that, socio-economic factors such as population density, dynamic population, distance from the centre and poverty all play an important role.
 These are important to keep in mind when reporting findings based on analysis of such data.
 
-even with this limitation, the inclusion of crowdsourced data which are both up-to-date and specific to the problem at hand can still provide new insight in addition to the knowledge from established sources of data collected in traditional methods [@birkin2011calibration].
-All of this helps us to understand the relations between the events and the occurrences that come together in unique places and provides a framework for understanding how places change over different time periods [@lonlongley2012geodemographics].
+Despite these limitation, the social sciences must embrace these new forms of data that, although messy, biased and noisy, have the potential to describe social phenomena better than well-organized small surveys or even national censuses [@savage2007coming].
+The inclusion of crowdsourced data which are both up-to-date and specific to the problem at hand can still provide new insight in addition to the knowledge from established sources of data collected in traditional methods [@birkin2011calibration].
+All of this helps us to understand the relations between the events and the occurrences that come together in unique places and provides a framework for understanding how places change over different time periods [@longley2012geodemographics].
 
-## Open Street Map
+---
 
-### What is Open Street Map?
+In sum, researchers and analysts using open data must first ask critical questions about:
 
-- Brief history.
-- Motivations for and advantages of an open web mapping platform.
-- Primary features (keys, values, elements).
-- Contributing to the data yourself.
+- **where** does this data come from? 
+- **why** was the data collected in the first place?
+- **who** is represented in the data, and who is excluded?
+- **what** concepts and constructs can and cannot be operationalised with this data?
+- **when** did data collection take place, and how might have that influenced results?
 
-### Downloading Open Street Map Data
+And meaningfully engage with the answers to these questions during conceptualisation and operationalisation phases of any research project.
 
-- Define and explain APIs.
-- For OSM, describe overpass queries and overpass-turbo.
-- Query example (brief).
+---
 
-### Using Open Street Map data in R
+# Practical exercise
 
-- Made easy through the `osmdata` package in R.
-- More straightforward than overpass-turbo.
-- Compatible with both `sp` and `sf` classes of spatial data.
-- We focus on `sf` as its compatible with the `tidyverse`.
-- Outline the key documentation and references.
+In this exercise we will demonstrate:
+
+- how to access open data using Application Programming Interfaces (APIs) and through downloading from official sites
+- an example of how to clean and wrangle data for analysis
+- how to link distinct data sets
+- how open data from different sources may tell us different things, and why engaging with the critical questions outline above is crucial to properly use open data in research and analysis.
+
+## Our aim
+
+We are interested in finding out how bus stops might affect crime.
+
+<!-- more about link between bus stops and crime here maybe something from Andy Newton/Vania Ceccato/ Anastasia Loukatou-Sideris/me -->
+
+## Accessing data
+
+We will be using two different types of open data in this exercise, public sector data (crime data and transport authority data) and crowdsourced data (from Open Street Map).
+To access them we will use three different methods: 1) Direct download, 2) Direct request to an API, 3) Request to an API using a wrapper.
+This will give you few different ideas about how you might go about accessing other open data sets relevant to your research.
+
+### Direct download
+
+The simplest way that open data is often made available is through direct download from a website.
+This is the case for open police data made available in the United Kingdom through the [data.police.uk](https://data.police.uk/) webportal.
+To access this data, simply type the url into any web browser ( [https://data.police.uk/](https://data.police.uk/) ), and follow the guidance to acquire data.
+In the case of police.uk you [select the tab which says "Data"](https://data.police.uk/data/), choose your "Date range:" (ie for what months you would like to download crime data for), choose the police forces you want crime data from, choose some options (whether you want to include crime data, outcomes data, and/or stop and search data) and click on 'Generate File' to download.
+
+Once you download this data you will have to save this locally on your computer. We have saved ours in a subfolder called "data" and will be loading the data into R directly from there.
+
+Some data downloads may come bundled with a data dictionary, explaining important information about your data set, which you will need to answer the critical questions outlined in previous sections.
+In other cases this might be found online.
+This is the case with the police.uk dataset, where the data dictionary can be found on the ["About" page](https://data.police.uk/about/).
+Make sure you always read through this correctly so you can answer your critical questions and fully understand your data set.
+
+### Direct request to an API
+
+An application programming interface (API) is a tool which defines an interface for a programme to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
+We are specifically using the term APIs to denote tools created by an open data provider to provide access to different subsets of their content.
+APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
+APIs can take many different forms and be of varying quality and usefulness [@foster2016big].
+For the purposes of accessing open data from the web, we are specifically talking about **RESTful APIs**.
+REST stands for Representational State Transfer.
+These APIs work directly over the web, meaning that we as users can play with the API with relative ease to understand how it works [@foster2016big].
+
+Here we will use the example of the British transit authority Transport for London (TfL) to practice constructing HTTP requests via their API, and then parsing the data that are returned.
+Much like how downloading data directly requires access to a data dictionary to make sense of the data and answer our critical questions, we must find a similar document to understand how the API we plan on using works.
+In the case of the TfL API this is provided via the [TfL Open Data page for their unified API](https://tfl.gov.uk/info-for/open-data-users/unified-api) and the [documentation page](https://api-portal.tfl.gov.uk/docs).
+We will engage more with this in the exercises.
+
+### Request to an API using a wrapper
+
+Often developers who work with APIs will share their code for doing so, and release them in the form of a package or module.
+This is called a **wrapper** because it's code that wraps the API in a nice package, making it easier for future users to access.
+The wrapper can take many forms, it could be a Python module, or an R package, or it can even be a web interface that provides a graphical user interface (GUI) for accessing the API in question.
+
+To demonstrate this, we will be accessing data from Open Street Map (OSM).
+OpenStreetMap is a crowdsourced open data set of spatial information built by a community of mappers that contribute and maintain data about all sorts of environmental features, such as roads, trails, cafes, railway stations, and much more, all over the world.
+The result is a map rich in local knowledge, which prvides valuable information about places and their features.
+
+OSM has two types of wrappers available for its API, a web-based GUI called [Overpass Turbo](https://overpass-turbo.eu/), and an R package called `osmdata`. We will explore using both.
+
+<!-- ### Open Street Map -->
+
+<!-- #### What is Open Street Map? -->
+
+<!-- - Brief history. -->
+<!-- - Motivations for and advantages of an open web mapping platform. -->
+<!-- - Primary features (keys, values, elements). -->
+<!-- - Contributing to the data yourself. -->
+
+<!-- ### data.police.uk -->
+
+<!-- ### Transport for London Open Data Portal -->
+
+<!-- ### Downloading Open Street Map Data -->
+
+<!-- - Define and explain APIs. -->
+<!-- - For OSM, describe overpass queries and overpass-turbo. -->
+<!-- - Query example (brief). -->
+
+<!-- ### Using Open Street Map data in R -->
+
+<!-- - Made easy through the `osmdata` package in R. -->
+<!-- - More straightforward than overpass-turbo. -->
+<!-- - Compatible with both `sp` and `sf` classes of spatial data. -->
+<!-- - We focus on `sf` as its compatible with the `tidyverse`. -->
+<!-- - Outline the key documentation and references. -->
 
@@
 
+Let's get started! Getting back to our problem at hand, crime at bus stops, we want to know how crimes and bus stops interact...
+<!-- clear up what research q we can actually answer....!! -->
+
 First, ensure that you have the relevant packages installed in R.
-Although the main package used in this demonstration is `osmdata`, for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation.
+Although the main packages used in this demonstration are `rjson` for parsing the TfL API results and `osmdata` for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation.
 If you don't have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
@@
 ```{r, eval = F}
+library(jsonlite)
 library(osmdata)
@@
 ```
+
+Now let's focus on acquiring our different data sets
+
+### Direct Dowload of crime data
+
+<!-- IDK how much detail we actually need here, probably not loads, about downloading extracting and importing police.uk data -->
+
+### Direct request to TfL API
+
+Our first step is to read through the .
+The tells us what sort of data is possible to acquire.
+We know that currently we want to find the location of bus stops, and any information that is available from TfL about these bus stops.
+In the documentation is where we find specifically *how* we can request this.
+
+Specifically, we can scroll down to where we find "".
+Next to this we see the example URL: `https://api.tfl.gov.uk/line/24/route/sequence/outbound` This example demonstrates how we can request this information, and gives us a template to alter, but also allows us to take a peak into the result this call will return.
+
+To see, try to take the above URL, and copy and paste it into your browser window. What do you see? It should be something like this:
+
+![](img/tfl_api_browser.png)
+
+This contains the information we requested (bus route 24 stops outbound) in a format called **JSON**, which stands for JavaScript Object Notation.
+JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
+
+Great, but how do we get this into R?
+Well this is where the `rjson` package is handy, and specifically we can use the `fromJSON()` function to parse this JSON p=object, which we can read directly from the URL into R using the `readLines()` function (base R).
+So, if we wanted the results of the bus route 24, we would type:
+
+```{r getrt24}
+
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/24/route/sequence/outbound"))
+
+```
+
+This gives us an object (`api_call`) which contains all the information returned by the TfL API.
+Now JSON is slightly different to dataframes which we are more used to dealing with, in that there is nested information
+
+```{r conferttodf}
+
+json_data_frame <- as.data.frame(api_call)
+
+```
+
+<!-- insert here tutorial on parsing the api data, this is a job for next week i'm tired -->
+
+### Using wrappers to access Open Street Map data
+
+<!-- Maybe start with the GUI-based one, and then do the below in R? -->
 

### 2020-05-05 draft_chapter.Rmd
Edits and additions 2

--- 2020-04-09 draft_chapter.Rmd
@@
 date: "June 2020"
-output: pdf_document
+output:
+  pdf_document:
+    latex_engine: xelatex
 editor_options: 
   chunk_output_type: console
+bibliography: refs.bib
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
+knitr::opts_chunk$set(echo = T, message = F, warning = F)
+library(jsonlite)
 library(osmdata)
@@
 
-## Introduction
+# Introduction
 
-- Topic definition.
-- Advantages and exciting prospects in open data.
-- Chapter structure outline.
-- Prerequisites (reference _Geocomputation with R_).
+Access to data in crime and place research, and even criminology more broadly, has traditionally been reserved to those who have the means to collect fresh data themselves, pay for access, or obtain data through formal data sharing agreements.
+Even when access is granted, the usage of these data often comes with conditions that circumscribe how the data can be used through licensing or policy [@kitchin2014data].
+The public dissemination of findings which emerge from analysis might even be subject to restrictions.
+This can lead to unequal access, controlled usage and curb the diffusion of findings, severely limiting the insight that can be attained from data.
 
-## Background
+Open data are a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements and negotiations [@manovich2011trending].
+Not only this, but open data can also lead to all sorts of novel insight within criminology and cime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional meaures [@solymosi2018role].
+As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the social world around them have the skills and know-how to access, accurately interpret, assess and analyse open data sets.
 
-- What is open data?
-- What types of organisations release open data?
-  - Public sector
-  - Private sector
-- Why is it needed?
-- What are the strengths?
-- What are the limitations?
-- Substantive examples in criminology (focus on crime and place research).
-- Conclude with: a key one is Open Street Map.
+This chapter aims to lay the groundwork for developing such skills by providing a framework to approach and meaningfully interpret open data, in terms of understanding its origin and the implications this might have on the data itself.
+The chapter also offers a practical hands-on guide to accessing, wrangling, and analysing open data in order to draw conclusions about crime and place.
+We achieve this by first giving a background on open data, discussing the key types of data out there, critically assessing the strengths and limitations, and building a how-to guide checklist for researchers to follow when approaching an open data source.
+Finally, we work through an example that shows how to access, wrangle, link, and interpret open data from various sources, providing a template which can be applied in other research areas.
 
-## Open Street Map
+# Background
 
-### What is Open Street Map?
+## What is open data?
 
-- Brief history.
-- Motivations for and advantages of an open web mapping platform.
-- Primary features (keys, values, elements).
-- Contributing to the data yourself.
+First, it is useful to clarify what we mean when we refer to 'open data'.
+The _Open Data Handbook_, compiled by the Open Knowledge Foundation, states that open data are "data that can be freely used, re-used and redistributed by anyone - subject only, at most, to the requirement to attribute and sharealike" [@dietrich2009open].
+Specifically, open data can be defined along three key domains [@knowledge2016open].
 
-### Downloading Open Street Map Data
+- **Availability and Access:** the data must be accessible via a public domain, at no more than a reasonable reproduction cost, and through a convenient medium, such as the internet.
+Ideally, it should be machine-readable and modifiable.
+For example, it should be downloadable from a website as an unlocked spreadsheet, or json file, rather than presented as a summary table in a PDF document.
+- **Re-use and Redistribution:** the data must be provided under conditions that allow free use, such as modification, separation, or compilation of the data and permit re-use and redistribution including the intermixing (merging) with other datasets.
+- **Universal Participation:** everyone should be allowed to use, re-use and redistribute the data and its derivatives without restriction.
+For instance, restrictions that only allow non-commercial or educational usage would not constitute open data.
 
-- Define and explain APIs.
-- For OSM, describe overpass queries and overpass-turbo.
-- Query example (brief).
+In short, 'open data' should be data that are readily and publicly available, in a usable format, and without reistrictions on usage, modification and re-distribution.
 
-### Using Open Street Map data in R
+Open data practices, and the extent to which it is a reality, vary both within and between countries.
+For example, the United States have a history of making public sector data sets openly available.
+The United Kingdom can be more strict, releasing data under a licence ('Open Government Licence'), with some data requiring a fee [@kitchin2014data].
+The _Global Open Data Index_ [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by countries.
+It providing a rank which indicates how well governments across the world follow open data practices on aggregate, but also according to specific domains.
+Many of these have direct relevance to crime and place research, such as administrative boundaries and geographic locations (e.g.
+coordinates).
 
-- Made easy through the `osmdata` package in R.
-- More straightforward than overpass-turbo.
-- Compatible with both `sp` and `sf` classes of spatial data.
-- We focus on `sf` as its compatible with the `tidyverse`.
-- Outline the key documentation and references.
+## What are types of open data?
+
+Now we have a reasonable idea about what open data is in a broad sense, we can consider the different types of open data which you might come across.
+Here, we formulate a typology of open data sets based on their origin to help provide an overview.
+
+### Public sector (including publicly funded research).
+
+Much of the open data movement has focused on information generated by local and national state agencies, or publicly funded research.
+Open data has become a key component of transparency and accountability in government, given that much of the data collected has been funded by the public purse [@kitchin2014data].
+In some countries, the national statistics body is independent from government, which serves to improve accountability and increase public trust in the data being collated and published.
+The _Global Open Data Index_ introduced earlier focuses on identifying the gaps in governmental organisations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
+
+As researchers studying crime and place, public sector data is particularly important.
+Many variables thought to explain the non-random distribution of crime, offenders and victims in space are coallated and published by governmental bodies.
+For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighbourhoods' - a theoretically important scale in criminology.
+Data describing urban structure, such as street networks, are often generated and distributed by public bodies, and can be used in street syntax analysis to assess whether the configeration of streets dictates things like violent crime victimisation [@streetsyntax2017].
+In many countries, victimisation surveys are openly available for download.
+This is not only useful for substantive research (e.g.
+quantifying long-term crime trends), but also scrutiny over reliability and accuracy in other forms of open data, such as police recorded crime records.
+Both of these examples represent an interesting challenge to open data principles, since data tends to be anonymised to some degree before being made available to the public.
+Access to the raw records is dependent on a number of requrements (e.g.
+background checks, contractual agreements, ICT security measures) which might exclude many individuals due to cost and practicality.
+
+The drive towards transparency and scrutiny of public information is further enhanced by an increasing need for replication - a hallmark for open science research practice.
+Opening up data sets used in criminology publications and wider social sciences fosters and facilitates a culture of replication [@pridemore2018replication].
+As such, open data fits into the wider discourse around transparency, open review and open scrutiny, which (in time) may even become a policy requirement irrespective of the research field or the source of funding [@vuong2017open].
+Increasingly, researchers are making use of online repositories for their papers (e.g.
+www.osf.io), data (e.g.
+www.ukdataservice.ac.uk/deposit-data) and code (e.g.
+www.github.com), all in the name of transparency and replication.
+These are guidelines and tools worth considering, not just when conducting analysis on secondary data sources, but also collecting and sharing your own data.
+
+### Private sector
+
+Opening up data generated by the private sector represents a significant challenge, largely due to the proprierary value to its creators [@kitchin2014data].
+The aims and objectives of private companies, who have a duty to shareholders and operate in a competitive market environment, differ considerably from local and central government.
+That said, some datasets are released openly by private sector organisations, albiet often only a subset of what would be available as a paying customer.
+The website for ArcGIS, a piece of software maintained by the Environmental Systems Research Institute (ESRI), a private company, host a number of open geographic datasets on their online cloud platform (www.hub.arcgis.com/search).
+Indeed, many papers have made use of data collected or distributed by private organisations to explore crime and place, such as Google Street View [@langton2017residential] and Twitter [@malleson2015impact].
+In this book you will learn to use the free Twitter API for such purposes in criminology (see Section 3, Topic 7).
+
+### Open crowdsourced data
+
+Finally, there is a specific group of open data sources that collate data collectively, generated by large groups of individuals who do not specifically belong to any organisation, but instead work together on a collaborative project.
+These are crowdsourced data sets.
+Examples include Wikipedia (www.wikipedia.org), an online encyclopedia where anyone can contribute, and Flickr (www.flickr.com) an online photo gallery where people can upload and tag their photos.
+Often, the organisations who maintain and monitor these data collection activities are charities or non-govermental organisations that do not operate for profit, but instead provide some form of social good.
+An example is the online reporting platform FixMyStreet (www.fixmystreet.com), where people can report problems such as instances of graffiti, vandalism or environment issues.
+The data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
+Another chapter in this book (see Section 3, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analysing open data of this type.
+
+## Strengths and limitations
+
+The defining characteristics of open data, namely, that of availability, re-usability and universal participation, outlined earlier, represent its greatest strength.
+But besides from this, there are other advantages.
+One specific motivation for using open data comes from its potential to address many of the limitations associated with traditional surveying methods, such as social desirability bias, or issues associated with memory and recall [@mayer2013big].
+This advantage exists largely due to the organic way in which open data is generated.
+It is often a by-product of other activities, and as such, we can gain an honest insight into people's everyday lives and associated social processes [@solymosi2018role].
+As noted, open data also means open research, facilitating transparency and reproducibility.
+That said, there are a number of shortcomings which are worthy of consideration.
+
+Firstly, one of the biggest obstacles to collecting open data is the ability to interpret what the data means (e.g.
+applying a framework for its analysis), and computational skills in data scraping, wrangling and cleaning, in order to transform it into a usable format for research [@boyd2012critical].
+All sorts of open data remain unanalysed, and or inaccessible, to large portions of the population, simply because they lack the skills and know-how.
+Although this is certainly an obstacle for data usage more generally, it represents a key challenge to the governmental bodies, in particular, on which there is an onus to ensure transparency and faciliate scrutiny.
+Private companies, especially those who generate open data as a by-product of their primary activity (e.g.
+Twitter), have little responsibility or pressure to ensure that their data is accessible and usable to the general public.
+Moreover, in both public and private spheres, licences and conditions on open data are not necessarily concrete, and might be subject to change with little or no notice.
+This is an important consideration when planning and running long-term research projects.
+
+Secondly, despite the merits of open data in terms of remedying pitfalls in traditional survey methods, there are new threats to validity that may emerge.
+For instance, whilst many online resources offer an 'honest' depiction of society, uncaptured by surveys, researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
+This could be a result of individuals willingly managing and 'curating' their online presence, or because of wider issues such as government consorship, or cultural norms around particular topics such as sexual assault.
+Relatedly, researchers should be aware of issues over sampling, and ultimately the generalisability of findings that emerge from the analysis of open data.
+In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduced a degree of bias [@longley2012geodemographics].
+Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
+Contributions to resources such as Open Street Map, which we look at in the practical exercise later, are correlated with contextual characterisics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
+These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
+
+Open data can be messy, biased and noisey, but criminology (and social sciences more generally) can benefit immeasurably from its use.
+Only through open data can public sector bodies be held to account, research be transparent and reproducible, and participation in data analysis universal.
+In crime and place research, both dependent variables (e.g.
+police recorded crime incidents, victimisation rates) and independent variables (e.g.
+demographic characteristics, ambient population estimates) can be sourced from open data, whether public, private or crowdsourced.
+With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science.
+In sum, researchers and analysts using open data must first ask critical questions:
+
+- **Where** does this data come from? 
+- **Why** was the data collected in the first place?
+- **Who** is represented in the data, and who is excluded?
+- **What** concepts and constructs can and cannot be operationalised with this data?
+- **When** did data collection take place, and how might have that influenced results?
+
+With this in mind, the following exercise utilises two different sources of open data on public transport and police recorded crime in London, England.
+
+# Practical exercise
+
+In this exercise we will demonstrate a number of different skills, including:
+
+- Accessing open data using Application Programming Interfaces (APIs) and direct downloads.
+- Cleaning and wrangling data for analysis.
+- Visualising open spatial data.
+- Comparing different sources of open data.
+- Engaging with the critical questions outline above to better understand open data.
+
+## Our aim
+
+In demonstrating these skills, we will be answering two key research questions.
+The first is a substantive empirical investigation into crime in and around London Underground stations, in the spirit of existing research into transport crime (see newton2014above).
+Studies in Sweden, for instance, have shown the importance of environmental and neighbourhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
+Investigations have also taken place into crime concentrations at bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime].
+As auch, we can pose the following research question:
+
+- RQ1: To what extent does crime cluster in and around London underground stations?
+
+The second is more methodological, and relates to our considered and critical usage of open data, by exploring the validity of different open data sources. As such, we can pose:
+
+- RQ2: How comparable are public sector and crowdsourced open data on London underground stations?
+
+First, we will go through how to access the data used in the exercises: an important skill when seeking to use open data.
+
+## Accessing data
+
+We will be using two different types of open data in this exercise, public sector data (crime data and transport authority data) and crowdsourced data (from Open Street Map).
+To access them we will use three different methods: 1) Direct download, 2) Direct request to an API, 3) Request to an API using a wrapper.
+This will give you few different ideas about how you might go about accessing other open data sets relevant to your research.
+
+### Direct download
+
+The simplest way that open data is often made available is through direct download from a website.
+This is the case for open police data made available in the United Kingdom through the [data.police.uk](https://data.police.uk/) webportal.
+To access this data, simply type the url into any web browser ( [https://data.police.uk/](https://data.police.uk/) ), and follow the guidance to acquire data.
+In the case of police.uk you [select the tab which says "Data"](https://data.police.uk/data/), choose your "Date range:" (ie for what months you would like to download crime data for), choose the police forces you want crime data from, choose some options (whether you want to include crime data, outcomes data, and/or stop and search data) and click on 'Generate File' to download.
+
+Once you download this data you will have to save this locally on your computer. We have saved ours in a subfolder called "data" and will be loading the data into R directly from there.
+
+Some data downloads may come bundled with a data dictionary, explaining important information about your data set, which you will need to answer the critical questions outlined in previous sections.
+In other cases this might be found online.
+This is the case with the police.uk dataset, where the data dictionary can be found on the ["About" page](https://data.police.uk/about/).
+Make sure you always read through this correctly so you can answer your critical questions and fully understand your data set.
+
+### Direct request to an API
+
+An application programming interface (API) is a tool which defines an interface for a programme to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
+We are specifically using the term APIs to denote tools created by an open data provider to provide access to different subsets of their content.
+APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
+APIs can take many different forms and be of varying quality and usefulness [@foster2016big].
+For the purposes of accessing open data from the web, we are specifically talking about **RESTful APIs**.
+REST stands for Representational State Transfer.
+These APIs work directly over the web, meaning that we as users can play with the API with relative ease to understand how it works [@foster2016big].
+
+Here we will use the example of the British transit authority Transport for London (TfL) to practice constructing HTTP requests via their API, and then parsing the data that are returned.
+Much like how downloading data directly requires access to a data dictionary to make sense of the data and answer our critical questions, we must find a similar document to understand how the API we plan on using works.
+In the case of the TfL API this is provided via the [TfL Open Data page for their unified API](https://tfl.gov.uk/info-for/open-data-users/unified-api) and the [documentation page](https://api-portal.tfl.gov.uk/docs).
+We will engage more with this in the exercises.
+
+### Request to an API using a wrapper
+
+Often developers who work with APIs will share their code for doing so, and release them in the form of a package or module.
+This is called a **wrapper** because it's code that wraps the API in a nice package, making it easier for future users to access.
+The wrapper can take many forms, it could be a Python module, or an R package, or it can even be a web interface that provides a graphical user interface (GUI) for accessing the API in question.
+
+To demonstrate this, we will be accessing data from Open Street Map (OSM).
+OpenStreetMap is a crowdsourced open data set of spatial information built by a community of mappers that contribute and maintain data about all sorts of environmental features, such as roads, trails, cafes, railway stations, and much more, all over the world.
+The result is a map rich in local knowledge, which prvides valuable information about places and their features.
+
+OSM has two types of wrappers available for its API, a web-based GUI called [Overpass Turbo](https://overpass-turbo.eu/), and an R package called `osmdata`. We will explore using both.
+
+<!-- ### Open Street Map -->
+
+<!-- #### What is Open Street Map? -->
+
+<!-- - Brief history. -->
+<!-- - Motivations for and advantages of an open web mapping platform. -->
+<!-- - Primary features (keys, values, elements). -->
+<!-- - Contributing to the data yourself. -->
+
+<!-- ### data.police.uk -->
+
+<!-- ### Transport for London Open Data Portal -->
+
+<!-- ### Downloading Open Street Map Data -->
+
+<!-- - Define and explain APIs. -->
+<!-- - For OSM, describe overpass queries and overpass-turbo. -->
+<!-- - Query example (brief). -->
+
+<!-- ### Using Open Street Map data in R -->
+
+<!-- - Made easy through the `osmdata` package in R. -->
+<!-- - More straightforward than overpass-turbo. -->
+<!-- - Compatible with both `sp` and `sf` classes of spatial data. -->
+<!-- - We focus on `sf` as its compatible with the `tidyverse`. -->
+<!-- - Outline the key documentation and references. -->
 
@@
 
+Let's get started! Getting back to our problem at hand, crime at bus stops, we want to know how crimes and bus stops interact...
+<!-- clear up what research q we can actually answer....!! -->
+
 First, ensure that you have the relevant packages installed in R.
-Although the main package used in this demonstration is `osmdata`, for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation.
+Although the main packages used in this demonstration are `rjson` for parsing the TfL API results and `osmdata` for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation.
 If you don't have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
 
-```{r, eval = F}
+```{r}
+library(jsonlite)
 library(osmdata)
@@
 ```
+
+Now let's focus on acquiring our different data sets
+
+### Direct Dowload of crime data
+
+<!-- IDK how much detail we actually need here, probably not loads, about downloading extracting and importing police.uk data -->
+
+### Direct request to TfL API
+
+Our first step is to read through the .
+The tells us what sort of data is possible to acquire.
+We know that currently we want to find the location of bus stops, and any information that is available from TfL about these bus stops.
+In the documentation is where we find specifically *how* we can request this.
+
+Specifically, we can scroll down to where we find "".
+Next to this we see the example URL: `https://api.tfl.gov.uk/line/24/route/sequence/outbound` This example demonstrates how we can request this information, and gives us a template to alter, but also allows us to take a peak into the result this call will return.
+
+To see, try to take the above URL, and copy and paste it into your browser window. What do you see? It should be something like this:
+
+```{r tfl, out.width = "\\textwidth", fig.cap="Data from TfL"}
+knitr::include_graphics("img/tfl_screenshot.png")
+```
+
+This contains the information we requested (bus route 24 stops outbound) in a format called **JSON**, which stands for JavaScript Object Notation.
+JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
+
+Great, but how do we get this into R?
+Well this is where the `rjson` package is handy, and specifically we can use the `fromJSON()` function to parse this JSON p=object, which we can read directly from the URL into R using the `readLines()` function (base R).
+So, if we wanted the results of the bus route 24, we would type:
+
+```{r getrt24, eval = F}
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/24/route/sequence/outbound"))
+```
+
+This gives us an object (`api_call`) which contains all the information returned by the TfL API.
+Now JSON is slightly different to dataframes which we are more used to dealing with, in that there is nested information
+
+```{r conferttodf, eval = F}
+json_data_frame <- as.data.frame(api_call)
+```
+
+<!-- insert here tutorial on parsing the api data, this is a job for next week i'm tired -->
+
+### Using wrappers to access Open Street Map data
+
+<!-- Maybe start with the GUI-based one, and then do the below in R? -->
 

### 2020-05-05 scripts/osm_explore_bike.R
Boris bikes

@@ -10,11 +10,12 @@ library(tidyr)
 # an area using the latitude and longitude coordinates, but it is generally
 # easier to use a search term. For instance
 
-bb <- getbb(place_name = "washington dc", format_out = "sf_polygon") 
+bb <- getbb(place_name = "greater london", format_out = "sf_polygon") 
 
 p <- opq(bbox = bb) %>%                                           # select bounding box
-  add_osm_feature(key = 'amenity', value = 'bar') %>% # select features
-  osmdata_sf()                                                    # specify class (sf or sp)
+  add_osm_feature(key = 'amenity', value = 'bicycle_rental') %>%
+  osmdata_sf() %>%                                                    # specify class (sf or sp)
+  trim_osmdata(bb_poly = bb)   
 
 p
 p.sf <- p$osm_points # extract point only
@@ -23,13 +24,27 @@ ggplot() +
   geom_sf(data = bb) +
   geom_sf(data = p.sf)
 
-p.sf  <- st_transform(p.sf, 27700)
-bb.sf <- st_transform(bb, 27700)
-p.clip.sf <- st_intersection(bb.sf, p.sf)
+tfl_bike <- api_call <- fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
 
-# ggplot() +
-#   geom_sf(data = bb) +
-#   geom_sf(data = p.clip.sf, size = 0.5)
+tfl_bike_sf <- tfl_bike %>% 
+  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326)
+
+
+
+p.sf <- st_transform(p.sf, 4326)
+
+ggplot() + 
+  geom_sf(data = bb) +
+  geom_sf(data = tfl_bike_sf) +
+  geom_sf(data = p.sf, col = "Red", alpha = 0.2)
+  
+  
+ggplot(p.sf) + geom_sf()
+
+ggplot() +
+  geom_sf(data = bb) +
+  geom_sf(data = tfl_bike_sf) +
+
 
 # crime data
 crime.df <- read_csv("data/2020-01-metropolitan-street.csv")

### 2020-05-05 scripts/osm_explore_tube.R

@@ -7,6 +7,7 @@ library(dplyr)
 library(tidyr)
 library(purrr)
 library(cowplot) # maybe don't show?
+library(stringr)
 
 # OSM query ==============================================================================================
 
@@ -17,31 +18,39 @@ library(cowplot) # maybe don't show?
 
 bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon") 
 
-osm_bus_sf <- opq(bbox = bb_sf) %>%                                # select bounding box
+osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
   add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
   osmdata_sf() %>%                                                 # specify class (sf or sp)
   trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box  
 
 # View contents.
-osm_bus_sf
+osm_stat_sf
 
 # Extract points only.
-osm_bus_sf <- osm_bus_sf$osm_points 
+osm_stat_sf <- osm_stat_sf$osm_points 
 
-# Filter Northern line only.
-osm_bus_sf <- osm_bus_sf %>% 
-  filter(line == "Northern")
+# Check the network tags.
+table(osm_stat_sf$network)
+table(osm_stat_sf$operator)
 
-osm_bus_sf <- st_transform(osm_bus_sf, 27700)
+# Filter any that include "London Underground".
+osm_tube_sf <- osm_stat_sf %>% 
+  filter(str_detect(network, "London Underground"))
+
+osm_tube_sf <- st_transform(osm_tube_sf, 27700)
 bb_sf      <- st_transform(bb_sf, 27700)
 
 ggplot() +
   geom_sf(data = bb_sf) +
-  geom_sf(data = osm_bus_sf, size = 0.3) 
+  geom_sf(data = osm_tube_sf, mapping = aes(colour = line))
 
 # TfL scrape ==============================================================================================
 
 # Scape bus stops on route 24 from TfL. Warning is fine.
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/Line/Meta/Modes"))
+
+
+
 api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/route/sequence/outbound"))
 
 # Extract bus stop names and the lat-long coordinates, transform to BNG.

### 2020-05-06 draft_chapter.Rmd
Edits and additions 3

@@
 Increasingly, researchers are making use of online repositories for their papers (e.g.
-www.osf.io), data (e.g.
-www.ukdataservice.ac.uk/deposit-data) and code (e.g.
-www.github.com), all in the name of transparency and replication.
+https://osf.io/), data (e.g.
+https://www.ukdataservice.ac.uk/deposit-data) and code (e.g.
+https://github.com/), all in the name of transparency and replication.
 These are guidelines and tools worth considering, not just when conducting analysis on secondary data sources, but also collecting and sharing your own data.
@@
 That said, some datasets are released openly by private sector organisations, albiet often only a subset of what would be available as a paying customer.
-The website for ArcGIS, a piece of software maintained by the Environmental Systems Research Institute (ESRI), a private company, host a number of open geographic datasets on their online cloud platform (www.hub.arcgis.com/search).
+The website for ArcGIS, a piece of software maintained by the Environmental Systems Research Institute (ESRI), a private company, host a number of open geographic datasets on their online cloud platform (https://hub.arcgis.com/search).
 Indeed, many papers have made use of data collected or distributed by private organisations to explore crime and place, such as Google Street View [@langton2017residential] and Twitter [@malleson2015impact].
@@
 These are crowdsourced data sets.
-Examples include Wikipedia (www.wikipedia.org), an online encyclopedia where anyone can contribute, and Flickr (www.flickr.com) an online photo gallery where people can upload and tag their photos.
+Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, and Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
 Often, the organisations who maintain and monitor these data collection activities are charities or non-govermental organisations that do not operate for profit, but instead provide some form of social good.
-An example is the online reporting platform FixMyStreet (www.fixmystreet.com), where people can report problems such as instances of graffiti, vandalism or environment issues.
+An example is the online reporting platform FixMyStreet (https://www.fixmystreet.com/), where people can report problems such as instances of graffiti, vandalism or environment issues.
 The data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
@@
 
-- Accessing open data using Application Programming Interfaces (APIs) and direct downloads.
+- Accessing open data using direct downloads and Application Programming Interfaces (APIs).
 - Cleaning and wrangling data for analysis.
@@
 Studies in Sweden, for instance, have shown the importance of environmental and neighbourhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
-Investigations have also taken place into crime concentrations at bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime].
+Investigations have also taken place into crime concentrations at bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
 As auch, we can pose the following research question:
 
-- RQ1: To what extent does crime cluster in and around London underground stations?
+- RQ1: To what extent does crime cluster in and around London Underground stations?
 
@@
 
-First, we will go through how to access the data used in the exercises: an important skill when seeking to use open data.
+First, we will go through how to access the open data used in the exercises.
 
@@
 
-We will be using two different types of open data in this exercise, public sector data (crime data and transport authority data) and crowdsourced data (from Open Street Map).
-To access them we will use three different methods: 1) Direct download, 2) Direct request to an API, 3) Request to an API using a wrapper.
+We will be using two different types of open data in this exercise: public sector data (police recorded crime data and local transport authority data) and crowdsourced data (from Open Street Map).
+To access them we will use three different methods, namely, (1) direct download, (2) direct request to an API, 3) request to an API using a wrapper.
 This will give you few different ideas about how you might go about accessing other open data sets relevant to your research.
+Locating, identifying and learning to access open data is a skill in itself.
 
@@
 
-The simplest way that open data is often made available is through direct download from a website.
-This is the case for open police data made available in the United Kingdom through the [data.police.uk](https://data.police.uk/) webportal.
-To access this data, simply type the url into any web browser ( [https://data.police.uk/](https://data.police.uk/) ), and follow the guidance to acquire data.
-In the case of police.uk you [select the tab which says "Data"](https://data.police.uk/data/), choose your "Date range:" (ie for what months you would like to download crime data for), choose the police forces you want crime data from, choose some options (whether you want to include crime data, outcomes data, and/or stop and search data) and click on 'Generate File' to download.
-
-Once you download this data you will have to save this locally on your computer. We have saved ours in a subfolder called "data" and will be loading the data into R directly from there.
-
-Some data downloads may come bundled with a data dictionary, explaining important information about your data set, which you will need to answer the critical questions outlined in previous sections.
-In other cases this might be found online.
-This is the case with the police.uk dataset, where the data dictionary can be found on the ["About" page](https://data.police.uk/about/).
-Make sure you always read through this correctly so you can answer your critical questions and fully understand your data set.
+The simplest way that open data can be made available is through direct download from a website.
+This is the case for open police recorded crime data in England and Wales.
+It can be accessed and downloaded using an online web portal (https://data.police.uk/) under Open Government Licence.
+Using the "Data" tab, anyone from researchers to members of the public can select the time period and police force of interest, and the type of information required (e.g.
+crimes, stop and search, outcomes), and then "Generate file" to create a download link.
+Data is downloaded in  comma-separated values format (.csv file extension).
+The website also contains detailed information about data quality, downloads and tools under the "About" tab, which is particularly useful when considering the critical questions posed earlier (where, why, who, what, when).
+For this exercise, we are going to use British Transport Police data, a force which operates on railways and light-rail systems across the country, for the month of January in 2020.
+Select the time period using the dropdown menus, the force using the tickboxes, and then generate and download the file.
+We save this file locally to a subfolder named "data", and during the exercise, the data is loaded into R from this folder.
 
@@
 
-An application programming interface (API) is a tool which defines an interface for a programme to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
-We are specifically using the term APIs to denote tools created by an open data provider to provide access to different subsets of their content.
+Another way of downloading open data is through an application programming interface (API).
+This is a tool which defines an interface for a programme to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
+Here, we are using the term 'API' to denote tools created by an open data provider to provide access to different subsets of their content.
 APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
 APIs can take many different forms and be of varying quality and usefulness [@foster2016big].
-For the purposes of accessing open data from the web, we are specifically talking about **RESTful APIs**.
-REST stands for Representational State Transfer.
-These APIs work directly over the web, meaning that we as users can play with the API with relative ease to understand how it works [@foster2016big].
+For the purposes of accessing open data from the web, we are specifically talking about _RESTful_ APIs.
+The 'REST' stands for Representational State Transfer.
+These APIs work directly over the web, which means users can play with the API with relative ease in order to understand how it works [@foster2016big].
 
-Here we will use the example of the British transit authority Transport for London (TfL) to practice constructing HTTP requests via their API, and then parsing the data that are returned.
-Much like how downloading data directly requires access to a data dictionary to make sense of the data and answer our critical questions, we must find a similar document to understand how the API we plan on using works.
-In the case of the TfL API this is provided via the [TfL Open Data page for their unified API](https://tfl.gov.uk/info-for/open-data-users/unified-api) and the [documentation page](https://api-portal.tfl.gov.uk/docs).
-We will engage more with this in the exercises.
+Here, we will use the example of a local government organisation responsible for transport, Transport for London (TfL).
+One of the main responsibilities of TfL is to oversee various rail networks in Greater London, including the London Underground.
+We will use open data from TfL to practice how to construct HTTP (i.e.
+web link) requests via their API, and then parse the data which is returned.
+Much like how downloading data directly requires access to a data dictionary to make sense of the data (e.g.
+understanding its origins and flaws), we must find a similar document to understand how the API we plan on using works.
+Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.
+Usefully, TfL provide a documentation page with examples of how you can use the HTTP links to access make calls to the API (https://api-portal.tfl.gov.uk/docs).
+We will actually make these calls from within the R environment using a 'wrapper', detailed below, but it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
 
@@
 
-Often developers who work with APIs will share their code for doing so, and release them in the form of a package or module.
-This is called a **wrapper** because it's code that wraps the API in a nice package, making it easier for future users to access.
-The wrapper can take many forms, it could be a Python module, or an R package, or it can even be a web interface that provides a graphical user interface (GUI) for accessing the API in question.
+Often, developers who work with APIs will share their code, and release them in the form of a package or module.
+This is called a _wrapper_ because it uses code that 'wraps' around the API to make it a neater, more usable package.
+Wrappers remove (or at least lower) many of the obstacles to accessing open data noted earlier, particularly for people from a non-technical background.
+The wrapper can take many forms, such as a Python module, or an R package.
+It could even be a web interface that provides a graphical user interface (GUI) for accessing the API in question.
 
-To demonstrate this, we will be accessing data from Open Street Map (OSM).
-OpenStreetMap is a crowdsourced open data set of spatial information built by a community of mappers that contribute and maintain data about all sorts of environmental features, such as roads, trails, cafes, railway stations, and much more, all over the world.
-The result is a map rich in local knowledge, which prvides valuable information about places and their features.
-
-OSM has two types of wrappers available for its API, a web-based GUI called [Overpass Turbo](https://overpass-turbo.eu/), and an R package called `osmdata`. We will explore using both.
-
-<!-- ### Open Street Map -->
-
-<!-- #### What is Open Street Map? -->
-
-<!-- - Brief history. -->
-<!-- - Motivations for and advantages of an open web mapping platform. -->
-<!-- - Primary features (keys, values, elements). -->
-<!-- - Contributing to the data yourself. -->
-
-<!-- ### data.police.uk -->
-
-<!-- ### Transport for London Open Data Portal -->
-
-<!-- ### Downloading Open Street Map Data -->
-
-<!-- - Define and explain APIs. -->
-<!-- - For OSM, describe overpass queries and overpass-turbo. -->
-<!-- - Query example (brief). -->
-
-<!-- ### Using Open Street Map data in R -->
-
-<!-- - Made easy through the `osmdata` package in R. -->
-<!-- - More straightforward than overpass-turbo. -->
-<!-- - Compatible with both `sp` and `sf` classes of spatial data. -->
-<!-- - We focus on `sf` as its compatible with the `tidyverse`. -->
-<!-- - Outline the key documentation and references. -->
+To demonstrate this, we will be accessing data from Open Street Map, a database of geospatial information built by a community of mappers, enthusiasts and members of the public, who contribute and maintain data about all sorts of environmental features, such as roads, green spaces, restaurants and railway stations, amongst many other things, all over the world.
+As such, it is a prime example of 'crowdsourced' open data.
+You an view the information contributed to Open Street Map using their online mapping platform (https://www.openstreetmap.org/).
+The result of people's contributions is database of spatial information rich in local knowledge which provides invaluable information about places and their features, without being subject to strict terms on usage.
+Open Street Map has two types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
+We will explore the APIs using both, but focus on their usage from within R.
 
@@
 
-Let's get started! Getting back to our problem at hand, crime at bus stops, we want to know how crimes and bus stops interact...
-<!-- clear up what research q we can actually answer....!! -->
+So, let's get started.
+Recall that are interested in examining the extent to which crime concentrates in and around London Underground stations, but also the contrasts between the open data available via TfL and Open Street Map, as public sector and crowdsourced open data, respectively.
 
@@
 library(tidyr)
+library(purrr)
+library(patchwork)
 ```
-
-Now let's focus on acquiring our different data sets
-
-### Direct Dowload of crime data
-
-<!-- IDK how much detail we actually need here, probably not loads, about downloading extracting and importing police.uk data -->
 
@@
 
-Our first step is to read through the .
-The tells us what sort of data is possible to acquire.
-We know that currently we want to find the location of bus stops, and any information that is available from TfL about these bus stops.
-In the documentation is where we find specifically *how* we can request this.
+Reading through the TfL API documentation mentioned earlier (https://api-portal.tfl.gov.uk/docs), you will get a feel for the kind of data available.
+To draw conclusions about the degree to which crime occurs in and around London Underground stations, we first need spatial information so we can plot station locations on a map.
+In the documentation there in fact an 'API area', which offer some guidance on the types of data available, called _Stops_.
+We can see that this include information on bus stops by route (which are numbered e.g.
+24), and underground stations by line (which are named e.g.
+"Bakerloo").
+The example for the underground stations is a URL: https://api.tfl.gov.uk/line/bakerloo/stoppoints.
+This has quite an intuitive structure, since we first want to access the TfL API, look for data by line, select the name of the line we are interested in, and finally pull unformation about where stations stops.
+In fact, you can simply copy and paste this URL into your internet browser, and it should look something like Figure 1.
 
-Specifically, we can scroll down to where we find "".
-Next to this we see the example URL: `https://api.tfl.gov.uk/line/24/route/sequence/outbound` This example demonstrates how we can request this information, and gives us a template to alter, but also allows us to take a peak into the result this call will return.
-
-To see, try to take the above URL, and copy and paste it into your browser window. What do you see? It should be something like this:
-
-```{r tfl, out.width = "\\textwidth", fig.cap="Data from TfL"}
+```{r tfl, out.width = "\\textwidth", fig.cap="Figure 1: example underground stations data from Transport for London", echo = F}
 knitr::include_graphics("img/tfl_screenshot.png")
@@
 
-This contains the information we requested (bus route 24 stops outbound) in a format called **JSON**, which stands for JavaScript Object Notation.
+Although this contains the information we requested, it is in a format called _JSON_.
+This stands for JavaScript Object Notation.
 JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
+As you can see, it is intuitive to read in its current format.
+In R, we can use the `fromJSON()` function from the `rjson` package to parse all this information into a data frame (with rows and columns).
+It will then become more familiar and usable.
+All we need to do is input the URL from the TfL API into the `txt` argument of the function, with a bit of help from `readLines()` in base R which is used to read text from URLs.
 
-Great, but how do we get this into R?
-Well this is where the `rjson` package is handy, and specifically we can use the `fromJSON()` function to parse this JSON p=object, which we can read directly from the URL into R using the `readLines()` function (base R).
-So, if we wanted the results of the bus route 24, we would type:
+To keep things focused, let's request data about stop points on the Northern Line. Note how this is simply an amended version of the example provided by TfL in the API documentation.
 
-```{r getrt24, eval = F}
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/24/route/sequence/outbound"))
+```{r getapi, eval = F}
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
 ```
@@
 This gives us an object (`api_call`) which contains all the information returned by the TfL API.
-Now JSON is slightly different to dataframes which we are more used to dealing with, in that there is nested information
+JSON is slightly different to tradtional data frames with rows and columns, which we are probably quite familiar with, because the data is nested.
+For instance, `api_call` is classed as a data frame, but upon viewing the object using `View(api_call)` you will notice that there are lists nested within it.
+Another way of checking this is looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
+This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated.
 
-```{r conferttodf, eval = F}
-json_data_frame <- as.data.frame(api_call)
+Fortunately for us, R is more than capable of dealing with nested data.
+Moreover, we can see that the information we are interested in (namely, the location of underground stations) has already been successfully parsed into columns called **lon** (longitude) and **lat** (latitude), along with an identication column of the stop names called **commonNanme**.
+We can extract these columns, and create a spatial `sf` object using the coordinates, all in one chunk of code, using the piping operator `%>%` in the `dplyr` package.
+Note that we define the Coordinate Reference System (CRS) which the data has come in, and then instantly transform it into something more appropriate (a projected CRS known as the British National Grid).
+
+```{r apispatial, eval = F}
+tfl_north_sf <- api_call %>% 
+  select(commonName, lat, lon) %>% 
+  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326) %>% 
+  st_transform(27700)
 ```

### 2020-05-06 scripts/osm_explore_bike.R

@@ -4,75 +4,31 @@ library(sf)
 library(readr)
 library(dplyr)
 library(tidyr)
+library(jsonlite)
 
-# All queries begin with a bounding box specification i.e. the study region.
-# This can be obtained manually, which requires some existing knowledge about
-# an area using the latitude and longitude coordinates, but it is generally
-# easier to use a search term. For instance
+bb_sf <- getbb(place_name = "greater london", format_out = "sf_polygon") 
 
-bb <- getbb(place_name = "greater london", format_out = "sf_polygon") 
+osm_bike_sf <- opq(bbox = bb_sf) %>%                                  # select bounding box.
+  add_osm_feature(key = 'amenity', value = 'bicycle_rental') %>%      # select keys and values.
+  osmdata_sf() %>%                                                    # specify class (sf or sp).
+  trim_osmdata(bb_poly = bb_sf)                                       # trim by the study region.   
 
-p <- opq(bbox = bb) %>%                                           # select bounding box
-  add_osm_feature(key = 'amenity', value = 'bicycle_rental') %>%
-  osmdata_sf() %>%                                                    # specify class (sf or sp)
-  trim_osmdata(bb_poly = bb)   
+osm_bike_sf
+osm_bike_sf <- osm_bike_sf$osm_points # extract point only
 
-p
-p.sf <- p$osm_points # extract point only
+osm_bike_sf <- osm_bike_sf %>% 
+  filter(brand == "Santander Cycles")
 
-ggplot() +
-  geom_sf(data = bb) +
-  geom_sf(data = p.sf)
+# TfL API
+tfl_bike_sf <- api_call <- fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
 
-tfl_bike <- api_call <- fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
+# Make spatial.
+tfl_bike_sf <- tfl_bike_sf %>%
+  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326) 
 
-tfl_bike_sf <- tfl_bike %>% 
-  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326)
-
-
-
-p.sf <- st_transform(p.sf, 4326)
-
[…]

### 2020-05-06 scripts/osm_explore_tube.R

@@ -6,8 +6,7 @@ library(readr)
 library(dplyr)
 library(tidyr)
 library(purrr)
-library(cowplot) # maybe don't show?
-library(stringr)
+library(patchwork)
 
 # OSM query ==============================================================================================
 
@@ -31,50 +30,64 @@ osm_stat_sf <- osm_stat_sf$osm_points
 
 # Check the network tags.
 table(osm_stat_sf$network)
-table(osm_stat_sf$operator)
 
-# Filter any that include "London Underground".
-osm_tube_sf <- osm_stat_sf %>% 
-  filter(str_detect(network, "London Underground"))
+# Filter Northern line only/
+osm_north_sf <- osm_stat_sf %>% 
+  filter(line == "Northern")
 
-osm_tube_sf <- st_transform(osm_tube_sf, 27700)
+osm_north_sf <- st_transform(osm_north_sf, 27700)
 bb_sf      <- st_transform(bb_sf, 27700)
 
 ggplot() +
   geom_sf(data = bb_sf) +
-  geom_sf(data = osm_tube_sf, mapping = aes(colour = line))
+  geom_sf(data = osm_north_sf, size = 0.8)
 
 # TfL scrape ==============================================================================================
 
-# Scape bus stops on route 24 from TfL. Warning is fine.
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/Line/Meta/Modes"))
-
-
-
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/route/sequence/outbound"))
+# Scape northern line locations from TfL. Warning is fine.
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
 
 # Extract bus stop names and the lat-long coordinates, transform to BNG.
-tfl_bus91_df <- api_call %>% 
-  pluck("stations") %>% 
-  select(name, lat, lon) %>% 
+tfl_north_sf <- api_call %>% 
+  select(commonName, lat, lon) %>% 
   st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326) %>% 
   st_transform(27700)
 
 # Plot difference.
 ggplot() +
-  geom_sf(data = tfl_bus91_df, size = 2) +
[…]

### 2020-05-06 draft_chapter.Rmd
Finish tube example in chapter

@@
 library(tidyr)
+library(patchwork)
 ```
@@
 
-```{r getapi, eval = F}
+```{r getapi}
 api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
@@
 Fortunately for us, R is more than capable of dealing with nested data.
-Moreover, we can see that the information we are interested in (namely, the location of underground stations) has already been successfully parsed into columns called **lon** (longitude) and **lat** (latitude), along with an identication column of the stop names called **commonNanme**.
-We can extract these columns, and create a spatial `sf` object using the coordinates, all in one chunk of code, using the piping operator `%>%` in the `dplyr` package.
-Note that we define the Coordinate Reference System (CRS) which the data has come in, and then instantly transform it into something more appropriate (a projected CRS known as the British National Grid).
+Moreover, we can see that the information we are interested in (namely, the location of underground stations) has already been successfully parsed into columns called **lon** (longitude) and **lat** (latitude), along with an identification column of the station names called **commonName**.
+We can extract these columns, and create an `sf` point object using these coordinates, all in one chunk of code, using the piping operator `%>%` in the `dplyr` package.
+Note that we define the Coordinate Reference System (CRS) which the data has come in (WGS 84), and then transform it into something more appropriate (the British National Grid, a projected CRS used in the United Kingdom).
 
-```{r apispatial, eval = F}
+```{r apispatial}
 tfl_north_sf <- api_call %>% 
@@
 
-<!-- insert here tutorial on parsing the api data, this is a job for next week i'm tired -->
+We can now easily visualize this data using the `ggplot2` package, which is compatable with `sf` objects.
+
+```{r tflmap}
+ggplot(tfl_north_sf) +
+  geom_sf()
+```
+
+There we have it: with just a few lines of code in R, we have directly queried the Transport for London API, a public sector source of open data, and plotted a basic map of stations on the Northern underground line.
 
@@
 
-<!-- Maybe start with the GUI-based one, and then do the below in R? -->
+Now, let's explore the equivalent data using crowdsourced open data.
+This approach differs from the TfL request in that our data source, Open Street Map, has a wrapper for the API built into an R package called `osmdata`.
+As such, rather than the API query being made directly (like we did for TfL, using the URL), we can use R code.
 
-All queries begin with a bounding box specification to define the study region.
+Of course, unlike the TfL API, `osmdata` is an international database, so all queries begin with a bounding box specification to define the study region.
 This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, but it is generally easier to use a search term.
-Here, we select Greater London in the United Kingdom using the `getbb()` function, specifying that we want the content as a simple features (sf) polygon.
+Here, we select Greater London using the `getbb()` function, specifying that we want the content as an `sf` polygon object which we name `bb_sf`.
 
@@
 
-We now have our study region defined as the administrative boundaries of Greater London.
-This can be visualised using `ggplot2` and the _simple features_ geometry `geom_sf()` available with `sf`.
-Note that by default, the Coordinate Reference System (CRS) is the World Geodetic System 84.
+We now have our study region defined as the administrative boundaries of Greater London, which we can visualize quickly to check it has worked as expected.
 
@@
 
-Now we have our study region, we can scrape data from the OSM API using the `opq()` function, which is short for 'Overpass query'.
-This allows you to build an Overpass query, outlined in the previous section, from within the R environment.
-We specify the bounding box object which is our study area, and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define what we want to pull from the API.
-As we noted earlier, features in OSM have are defined through keys and values.
-Here, we specify that we want amenities (the key) defined as bicycle parking (the value).
-This query is then piped through to `osmdata_sf()` which ensures that the resulting object is a _simple features_ class for easy plotting with `ggplot2`.
-We trim the features pulled from the API using `tim_osmdata()` to ensure that everything stays within the boundaries of our study region.
+Now we have our study region, we can scrape data from the OSM API using the `opq()` function.
+The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
+A common way of doing this is using 'Overpass turbo' (https://overpass-turbo.eu/) which allows you to query the database alongside an interactive map which automatically updates based on your results.
+In the interests of reproducibility consistently, we will make these queries from within R using the `osmdata` wrapper.
+
+First, we specify the bounding box object which is our study area (`bb_sf`), and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define the feature which we want to pull from the API.
+Features in Open Street Map are defined through 'keys' and 'values'.
+Keys are used to describe a broad category of features (e.g.
+highway, amenity), and values are more specific descriptions of each key (e.g.
+cycleway, bar).
+These are tags which contributors to Open Street Map have defined.
+A useful way to explore these is by using the comprehensive Open Street Map Wiki page on map features (https://wiki.openstreetmap.org/wiki/Map_Features).
+Here, it is fairly intuitive, because our key is 'public transport' and our value is 'station'.
+This is not _exactly_ what we want yet, because we are only interested in underground stations, but again, this highlights an interesting obstacle when accessing open data which could serve any number of research purposes.
+Note that we specify that we want the output to be an `sf` object, and that we want to trim the queried features by our study region `bb_sf`.
 
 ```{r}
-bikes_sf <- opq(bbox = bb_sf) %>%                                 # select bounding box
-  add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
-  osmdata_sf() %>%                                                # specify class
-  trim_osmdata(bb_poly = bb_sf)                                   # trim to region
+osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
+  add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
+  osmdata_sf() %>%                                                 # specify class (sf or sp)
+  trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box 
 ```
 
-The resulting object `bikes_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the Console.
+The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the **Console**.
 
 ```{r}
-bikes_sf
+osm_stat_sf
 ```
 
-This confirms details like the bounding box, but also provides information on the simple features collected from the query.
-As one might expect, most information relating to bicycle parking has been recorded using points (i.e.
+This confirms details like the bounding box coordinates, but also provides information on the features collected from the query.
+As one might expect, most information relating to public transport station locations has been recorded using points (i.e.
 two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing.
 We also have around one hundred polygons.
-For now, let's extract the point information only and then transform the CRS to the BNG.
+For now, let's extract the point information.
 
 ```{r}
-bikes_points_sf <- bikes_sf$osm_points 
+osm_stat_sf <- osm_stat_sf$osm_points 
 ```
 
-We can then plot these points over our original boundaries of Greater London. We reduce the default size of the point to ensure that we avoid too much overlap between bicycle parking locations.
+We could visualize these points using `ggplot2`, just as we did earlier, but are currently working with any form of public transport labeled as a 'station' by contributors to Open Street Map.
+Exploring the `osm_stat_sf` object will reveal a number of additional variables which we can use to narrow down our selection, such as the _network_ and _line_ on which the station is located.
+You will notice from running the below code, which opens up a frequency table for each combination of network and line, that we have the information required to do this.
+
+```{r, results = F, eval = F}
+osm_stat_sf %>%
+  as_tibble() %>% 
+  group_by(network, line) %>% 
+  tally() %>% 
+  View()
+```
+
+In fact, because all Northern line stations are labeled as London Underground, we only need to filter by the line.
+
+```{r}
+osm_north_sf <- osm_stat_sf %>% 
+  filter(line == "Northern")
+```
+
+Niw, we can then plot these points over the station locations pulled from the TfL API. We ignore the study region boundary to get a more detailed view.
 
@@
 ggplot() +
-  geom_sf(data = bb_sf) +
-  geom_sf(data = bikes_points_sf, size = 0.3)
+  geom_sf(data = tfl_north_sf) +
+  geom_sf(data = osm_north_sf, color = "red", alpha = 0.5) 
 ```
 
-As we can see, most bicycle parking spaces are clustered around the city centre, especially just north of the river Thames.
-It is also possible to make out key roads flowing in and out of the city centre, which contain bicycle parking all along the street.
-
-Using open police recorded crime data we can then plot actual incidences of bicycle theft to explore whether there is a spatial relationship between bike theft and parking spots in Greater London.
-For this example, we just use crime recorded as occurring in January 2020.
-First, let's load in the data as it downloaded raw from https://data.police.uk/data/.
+As we can see, there is clearly some discrepency between the two data sources.
+This gives us a preliminary answer to our second research question already: data relating to London Underground stations on Open Street Map appears to be incomplete.
+As such, the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative.
+But what might the impact be on studying crime in and around London Underground stations?
+For that, we can load in our police recorded crime data which we downloaded directly from the web earlier.
 
 ```{r}
-crime.df <- read_csv("data/2020-01-metropolitan-street.csv")
+btp_df <- read_csv("data/2020-01-btp-street.csv")
 ```
 
-We then need to conduct a bit of preliminary data handling: filter crimes which were tagged as bicycle theft, convert the latitude and longitude columns to coordinates with _simple features_, state the WGS 84 CRS and then clip the points by our study region.
+In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions we used earlier.
+Note that we need to add `st_intersection()` to clip our points to our study region, because the British Transport Police are operational across the country, not just in London.
+To do this, we first project `bb_sf` to the British National Grid, so that we perform the intersection on data which has the same projection.
 
 ```{r}
-bike.crime.sf <- crime.df %>% 
-  filter(`Crime type` == "Bicycle theft") %>% 
+# First, project the study region boundaries.
+bb_sf <- st_transform(bb_sf, 27700)
+
+# Then, make the police data spatial, and clip.
+btp_sf <- btp_df %>%
   drop_na(Longitude, Latitude) %>%
   st_as_sf(coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>%
+  st_transform(27700) %>%
   st_intersection(bb_sf)
@@
 
-To demonstrate the data in its entirety, we can plot the Greater London boundaries, overlayed with the bicycle parking space locations, and open crime data about bicycle thefts.
-It is worth clarifying that open police recorded crime data in England and Wales is spatially anonymised by a process of snapping points to a pre-defined grid (see @tompson2015uk).
-For that reason, many of these points overlap, and thus a degree of transparency is used for the points.
+We now have the location of all crimes recorded by the British Transport Police in Greater London during January, 2020.
+For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 metre buffer around each station, for each source of data, and counting the number of points falling within each.
+It is worth noting that this definition is somewhat arbitary, and is subject to a degree of spatial inaccuracy is open police recorded crime data [see tompson2015uk].
 
 ```{r}
-ggplot() +
-  geom_sf(data = bb_sf) +
-  geom_sf(data = bikes_points_sf, aes(colour = "bike_park"), size = 0.4, alpha = 0.5) +
-  geom_sf(data = bike.crime.sf, aes(colour = "bike_crime"), size = 0.3, alpha = 0.5) +
-  scale_colour_manual(name = NULL, values = c(bike_park = "black", bike_crime = "red")) +
-  theme(legend.position = "bottom")
+# Assign the same CRS.
+osm_north_sf <- st_transform(osm_north_sf, 27700)
+tfl_north_sf <- st_transform(tfl_north_sf, 27700)
+
+# Create buffers to define 'in and around'.
+osm_buff_sf <- st_buffer(osm_north_sf, dist = 50)
+tfl_buff_sf <- st_buffer(tfl_north_sf, dist = 50)
+
+# Count number of crimes recorded within each buffer.
+osm_north_sf <- osm_buff_sf %>% 
+  mutate(crimes = lengths(st_intersects(osm_buff_sf, btp_sf)))
+
+tfl_north_sf <- tfl_buff_sf %>% 
+  mutate(crimes = lengths(st_intersects(tfl_buff_sf, btp_sf)))
 ```
 
-### Bus stop example
+We can then visualise these counts and compare the two sources of data by colouring in our point locations according to the crime count.
+Note that we arrange the plots side-by-side using syntax available using the `patchwork` package.
 
 ```{r}
-bus_sf <- opq(bbox = bb_sf) %>%                                 # select bounding box
-  add_osm_feature(key = 'highway', value = 'bus_stop') %>%      # select features
-  osmdata_sf() %>%                                              # specify class
-  trim_osmdata(bb_poly = bb_sf)  
+p1 <- ggplot(data = osm_north_sf) +
+  geom_sf(mapping = aes(colour = crimes), size = 2) +
+  labs(title = "Open Street Map") +
+  scale_colour_viridis_c()
+
+p2 <- ggplot(data = tfl_north_sf) +
+  geom_sf(mapping = aes(colour = crimes), size = 2) +
+  labs(title = "Transport for London") +
+  scale_colour_viridis_c()
+
+p1 + p2
 ```

### 2020-05-06 scripts/osm_explore_tube.R

@@ -26,8 +26,11 @@ osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select boun
 osm_stat_sf
 
 # Extract points only.
-osm_stat_sf <- osm_stat_sf$osm_points 
+temp <- osm_stat_sf$osm_polygons
+osm_stat_sf <- osm_stat_sf$osm_points
 
+  
+  
 # Check the network tags.
 table(osm_stat_sf$network)
 

### 2020-05-07 draft_chapter.Rmd
Discussion

@@
 Open data are a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements and negotiations [@manovich2011trending].
-Not only this, but open data can also lead to all sorts of novel insight within criminology and cime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional meaures [@solymosi2018role].
+Not only this, but open data can also lead to all sorts of novel insight within criminology and crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
 As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the social world around them have the skills and know-how to access, accurately interpret, assess and analyse open data sets.
@@
 
-In short, 'open data' should be data that are readily and publicly available, in a usable format, and without reistrictions on usage, modification and re-distribution.
+In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
 
@@
 As researchers studying crime and place, public sector data is particularly important.
-Many variables thought to explain the non-random distribution of crime, offenders and victims in space are coallated and published by governmental bodies.
+Many variables thought to explain the non-random distribution of crime, offenders and victims in space are collated and published by governmental bodies.
 For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighbourhoods' - a theoretically important scale in criminology.
-Data describing urban structure, such as street networks, are often generated and distributed by public bodies, and can be used in street syntax analysis to assess whether the configeration of streets dictates things like violent crime victimisation [@streetsyntax2017].
+Data describing urban structure, such as street networks, are often generated and distributed by public bodies, and can be used in street syntax analysis to assess whether the configuration of streets dictates things like violent crime victimisation [@streetsyntax2017].
 In many countries, victimisation surveys are openly available for download.
@@
 Both of these examples represent an interesting challenge to open data principles, since data tends to be anonymised to some degree before being made available to the public.
-Access to the raw records is dependent on a number of requrements (e.g.
+Access to the raw records is dependent on a number of requirements (e.g.
 background checks, contractual agreements, ICT security measures) which might exclude many individuals due to cost and practicality.
@@
 
-Opening up data generated by the private sector represents a significant challenge, largely due to the proprierary value to its creators [@kitchin2014data].
+Opening up data generated by the private sector represents a significant challenge, largely due to the proprietary value to its creators [@kitchin2014data].
 The aims and objectives of private companies, who have a duty to shareholders and operate in a competitive market environment, differ considerably from local and central government.
-That said, some datasets are released openly by private sector organisations, albiet often only a subset of what would be available as a paying customer.
+That said, some datasets are released openly by private sector organisations, albeit often only a subset of what would be available as a paying customer.
 The website for ArcGIS, a piece of software maintained by the Environmental Systems Research Institute (ESRI), a private company, host a number of open geographic datasets on their online cloud platform (https://hub.arcgis.com/search).
@@
 Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, and Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
-Often, the organisations who maintain and monitor these data collection activities are charities or non-govermental organisations that do not operate for profit, but instead provide some form of social good.
+Often, the organisations who maintain and monitor these data collection activities are charities or non-governmental organisations that do not operate for profit, but instead provide some form of social good.
 An example is the online reporting platform FixMyStreet (https://www.fixmystreet.com/), where people can report problems such as instances of graffiti, vandalism or environment issues.
@@
 applying a framework for its analysis), and computational skills in data scraping, wrangling and cleaning, in order to transform it into a usable format for research [@boyd2012critical].
-All sorts of open data remain unanalysed, and or inaccessible, to large portions of the population, simply because they lack the skills and know-how.
-Although this is certainly an obstacle for data usage more generally, it represents a key challenge to the governmental bodies, in particular, on which there is an onus to ensure transparency and faciliate scrutiny.
+All sorts of open data remain unanalyzed, and or inaccessible, to large portions of the population, simply because they lack the skills and know-how.
+Although this is certainly an obstacle for data usage more generally, it represents a key challenge to the governmental bodies, in particular, on which there is an onus to ensure transparency and facilitate scrutiny.
 Private companies, especially those who generate open data as a by-product of their primary activity (e.g.
@@
 For instance, whilst many online resources offer an 'honest' depiction of society, uncaptured by surveys, researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
-This could be a result of individuals willingly managing and 'curating' their online presence, or because of wider issues such as government consorship, or cultural norms around particular topics such as sexual assault.
-Relatedly, researchers should be aware of issues over sampling, and ultimately the generalisability of findings that emerge from the analysis of open data.
+This could be a result of individuals willingly managing and 'curating' their online presence, or because of wider issues such as government censorship, or cultural norms around particular topics such as sexual assault.
+Relatedly, researchers should be aware of issues over sampling, and ultimately the generalizability of findings that emerge from the analysis of open data.
 In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduced a degree of bias [@longley2012geodemographics].
 Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
-Contributions to resources such as Open Street Map, which we look at in the practical exercise later, are correlated with contextual characterisics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
+Contributions to resources such as Open Street Map, which we look at in the practical exercise later, are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
 These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
@@
 
-With this in mind, the following exercise utilises two different sources of open data on public transport and police recorded crime in London, England.
+With this in mind, the following exercise utilizes two different sources of open data on public transport and police recorded crime in London, England.
 
@@
 - Cleaning and wrangling data for analysis.
-- Visualising open spatial data.
+- Visualizing open spatial data.
 - Comparing different sources of open data.
@@
 Investigations have also taken place into crime concentrations at bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
-As auch, we can pose the following research question:
+As such, we can pose the following research question:
 
@@
 Another way of downloading open data is through an application programming interface (API).
-This is a tool which defines an interface for a programme to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
+This is a tool which defines an interface for a program to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
 Here, we are using the term 'API' to denote tools created by an open data provider to provide access to different subsets of their content.
@@
 
-```{r}
+```{r packages}
 library(jsonlite) 
@@
 The example for the underground stations is a URL: https://api.tfl.gov.uk/line/bakerloo/stoppoints.
-This has quite an intuitive structure, since we first want to access the TfL API, look for data by line, select the name of the line we are interested in, and finally pull unformation about where stations stops.
+This has quite an intuitive structure, since we first want to access the TfL API, look for data by line, select the name of the line we are interested in, and finally pull information about where stations stops.
 In fact, you can simply copy and paste this URL into your internet browser, and it should look something like Figure 1.
 
-```{r tfl, out.width = "\\textwidth", fig.cap="Figure 1: example underground stations data from Transport for London", echo = F}
+```{r tfl, out.width = "\\textwidth", fig.cap="example of data from Transport for London", echo = F}
 knitr::include_graphics("img/tfl_screenshot.png")
@@
 This gives us an object (`api_call`) which contains all the information returned by the TfL API.
-JSON is slightly different to tradtional data frames with rows and columns, which we are probably quite familiar with, because the data is nested.
+JSON is slightly different to traditional data frames with rows and columns, which we are probably quite familiar with, because the data is nested.
 For instance, `api_call` is classed as a data frame, but upon viewing the object using `View(api_call)` you will notice that there are lists nested within it.
@@
 
-We can now easily visualize this data using the `ggplot2` package, which is compatable with `sf` objects.
+We can now easily visualize this data using the `ggplot2` package, which is compatible with `sf` objects.
 
@@
 
-There we have it: with just a few lines of code in R, we have directly queried the Transport for London API, a public sector source of open data, and plotted a basic map of stations on the Northern underground line.
+There we have it: with just a few lines of code in R, we have queried the Transport for London API, a public sector source of open data, and plotted a basic map of stations on the Northern underground line.
 
@@
 
-```{r}
+```{r bb}
 bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon")
@@
 
-```{r}
+```{r bbmap}
 ggplot(data = bb_sf) +
@@
 A common way of doing this is using 'Overpass turbo' (https://overpass-turbo.eu/) which allows you to query the database alongside an interactive map which automatically updates based on your results.
-In the interests of reproducibility consistently, we will make these queries from within R using the `osmdata` wrapper.
+In the interests of consistency, we will make these queries from within R using the `osmdata` wrapper.
 
@@
 
-```{r}
+```{r opq}
 osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
   add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
-  osmdata_sf() %>%                                                 # specify class (sf or sp)
+  osmdata_sf() %>%                                                 # specify class
   trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box 
@@
 
-```{r}
+```{r printstations}
 osm_stat_sf
@@
 
-```{r}
+```{r stationpoints}
 osm_stat_sf <- osm_stat_sf$osm_points 
@@
 
-```{r, results = F, eval = F}
+```{r freqtab, results = F, eval = F}
 osm_stat_sf %>%
@@
 
-In fact, because all Northern line stations are labeled as London Underground, we only need to filter by the line.
+In fact, because all Northern line stations are labeled as "London Underground", we only need to filter by the line.
 
-```{r}
+```{r filterline}
 osm_north_sf <- osm_stat_sf %>% 
@@
 
-Niw, we can then plot these points over the station locations pulled from the TfL API. We ignore the study region boundary to get a more detailed view.
+Now, we can then plot these points over the station locations pulled from the TfL API. We do not plot the study region boundary in order to get a more detailed, local view.
 
-```{r}
+```{r mapline}
 ggplot() +
@@
 
-As we can see, there is clearly some discrepency between the two data sources.
-This gives us a preliminary answer to our second research question already: data relating to London Underground stations on Open Street Map appears to be incomplete.
-As such, the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative.
-But what might the impact be on studying crime in and around London Underground stations?
-For that, we can load in our police recorded crime data which we downloaded directly from the web earlier.
+As we can see, there is clearly some discrepancy between the two data sources.
+There are slight differences in the exact point locations of stations, but more saliently,  Open Street Map appears to be missing stations, especially in Central and South London.
+This could be due to a number of reasons, amongst them, variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting].
+Given that South London is historically less wealthy in comparison to the North, for instance, this would certainly be an interesting avenue to pursue.
+Interestingly, these missing stations _have_ actually been recorded by contributors in Open Street Map, but _without_ information about the line on which the station is located.
+We can see this by plotting out all Open Street Map stations using `osm_stat_sf`.
+That said, without local knowledge, an alternative open data source, or a critical approach to using open data, well-intentioned researchers might draw erroneous conclusions from filtering stations tagged as on the Northern line.
+This gives us a preliminary answer to our second research question already: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.
 
-```{r}
+But, what might the impact of this be on studying crime in and around London Underground stations, given the differences observed between the two open data sources?
+For that, we can load in our British Transport Police data which we downloaded directly from the web earlier.
+
+```{r crimeload}
 btp_df <- read_csv("data/2020-01-btp-street.csv")
@@
 Note that we need to add `st_intersection()` to clip our points to our study region, because the British Transport Police are operational across the country, not just in London.
-To do this, we first project `bb_sf` to the British National Grid, so that we perform the intersection on data which has the same projection.
+To do this, we first project `bb_sf` to the British National Grid, so that we perform the intersection on data which have the same projection.
 
-```{r}
+```{r crimespatial}
 # First, project the study region boundaries.
@@
 We now have the location of all crimes recorded by the British Transport Police in Greater London during January, 2020.
-For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 metre buffer around each station, for each source of data, and counting the number of points falling within each.
-It is worth noting that this definition is somewhat arbitary, and is subject to a degree of spatial inaccuracy is open police recorded crime data [see tompson2015uk].
+For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
+It is worth noting that this definition is somewhat arbitrary, and is subject to a degree of spatial inaccuracy is open police recorded crime data [see tompson2015uk], but serves to facilitate this demonstration.
 
-```{r}
+```{r buffer}
 # Assign the same CRS.
@@
   mutate(crimes = lengths(st_intersects(osm_buff_sf, btp_sf)))
-
 tfl_north_sf <- tfl_buff_sf %>% 
@@
 
-```{r}
+```{r crimemapcompare}
 p1 <- ggplot(data = osm_north_sf) +
@@
 
+A number of factors emerge from this visualization.
+First, issues arising disparity between the stations identified on the Northern line are compounded.
+Using the Open Street Map data in isolation, we are grossly underestimating the the number of crimes occurring in and around Northern line stations.
+This could have significant implications for how police allocate resource across the underground network.
+Secondly, the spatial patterning of crime on underground stations differs considerably between the two data sources.
+Using the Open Street Map data, the final station on the line, Morden, has the greatest concentration of crimes, whereas the TfL data suggests this is in the city centre, at Tottenham Court Road.
+The conclusions drawn are starkly different, and would determine the consistency of findings with existing research examining crime at underground stations [@ceccato2013security].
+
+That said, this is not to say that Open Street Map is an inadequate source of open data for crime and place research.
+Firstly, it is an ever-growing resource.
+It is plausible (indeed, likely) that in the near future, the missing information about London Underground lines will be completed.
+Secondly, as noted, systematic variation in the extent to which individuals have contributed to the database can generate interesting discussions and avenues of research in its own right.
+Thirdly, this is simply one of thousands of features contained in the Open Street Map database.
+If we were interested in examining concentrations of bicycle theft from bike rental docking stations (so-called "Boris bikes") in Greater London, we could query open data APIs from TfL and Open Street Map, just as we did earlier, but using slightly amended criteria.
+This is demonstrated below, all in one code chunk.
+
+```{r bikexample}
+bb_sf <- getbb(place_name = "greater london", format_out = "sf_polygon")
+
+osm_bike_sf <- opq(bbox = bb_sf) %>%                                 
+  add_osm_feature(key = 'amenity', value = 'bicycle_rental') %>%     
+  osmdata_sf() %>%                                                  
+  trim_osmdata(bb_poly = bb_sf)
+
+osm_bike_sf <- osm_bike_sf$osm_points
+
+osm_bike_sf <- osm_bike_sf %>% 
+  filter(brand == "Santander Cycles")
+
+tfl_bike_sf <- fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
+
+tfl_bike_sf <- tfl_bike_sf %>%
+  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326)
+
+ggplot() + 
+  geom_sf(data = osm_bike_sf) +
+  geom_sf(data = tfl_bike_sf, col = "red", alpha = 0.4)
+```
+
+As showcased, the resulting visualizations are highly similar (a Spatial Point Pattern Test will confirm this, see Section 4 Topic 4).
+Refer back to the example earlier if some of the code is unclear.
+Note that this is a fantastic benefit of using an API wrapper from within R, because the code used to query each open data source can easily be re-used, with minor amendments, to extract different features.
+
+# Discussion
+
+## Summary
+
+This chapter has sought to introduce open data as a novel, growing and invaluable tool in crime and place research through a review of key fundamentals and a substantive demonstration within R.
+We have defined open data, and outlined some key types of open information available, along with their respective strengths and weaknesses.
+An important component of this review has been encourage an engaged and critical research approach to using open data.
+Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources.
+By way of an example, we accessed different sources of open data in order to explore crime concentrations at London Underground stations: police recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
+We found key differences between the two sources of transport data.
+Specifically, missing information in Open Street Map resulted in a skewed picture of crime occurring in and around underground stations.
+This highlighted both the power of open data, but also its shortcomings.
+We concluded with an additional demonstration using bike rental docking stations in Greater London, which showcased a scenario in which both public sector and crowdsourced open data provide unique (and comparable) insight.
+
 ## Future of open data
@@
 
-## Conclusion
-
-- Re-cap on what we've covered.
-- Wrap-up the key points.
-
 # References

### 2020-05-07 draft_chapter.Rmd
Discussion begin

@@
 
+As it stands, open data represents a key resource in crime and place research, but as the discussions and demonstrations in this chapter have shown, there is still a long way to go.
+Firstly, as with any emerging domain, open data sources vary considerably in their degree of completeness and reliability.
+Whilst this can offer insight, for instance, in terms of public participation and sense of community, it can also represents a significant obstacle to empirical analysis.
+A key challenge will be to ensure that students and researcher practitioners ask critical questions of their data before blindly delving into analysis and public dissemination.
+Secondly, the longevity of open data licences remains an unknown parameter.
+With so much open data being distributed by the public sector, the accessibility of open information is subject to shifts to changes in government, along with changes in the amount of resource available to safe-guard continuity.
+Thirdly,
+
+Open data is a fundamental component in the ever-growing movement towards transparent and reproducible scientific research.
+
 - Threats to its sustainability (e.g. licence expiry).

### 2020-05-08 draft_chapter.Rmd
First draft done

--- 2020-05-07 draft_chapter.Rmd
@@
 ---
-title: "Topic 6.2: Open Data"
+title: "Open Data"
 author: "Samuel Langton (MMU) & Reka Solymosi (University of Manchester)"
@@
 
-Access to data in crime and place research, and even criminology more broadly, has traditionally been reserved to those who have the means to collect fresh data themselves, pay for access, or obtain data through formal data sharing agreements.
+Access to data in crime and place research, and even criminology more broadly, has traditionally been reserved for those who have the means to collect fresh data themselves, pay for access, or obtain data through formal data sharing agreements.
 Even when access is granted, the usage of these data often comes with conditions that circumscribe how the data can be used through licensing or policy [@kitchin2014data].
@@
 Open data are a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements and negotiations [@manovich2011trending].
-Not only this, but open data can also lead to all sorts of novel insight within criminology and crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
-As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the social world around them have the skills and know-how to access, accurately interpret, assess and analyse open data sets.
+Not only this, but open data can also lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
+As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the social world around them have the skills and know-how to access, interpret, critique and analyse open data sets.
 
-This chapter aims to lay the groundwork for developing such skills by providing a framework to approach and meaningfully interpret open data, in terms of understanding its origin and the implications this might have on the data itself.
-The chapter also offers a practical hands-on guide to accessing, wrangling, and analysing open data in order to draw conclusions about crime and place.
-We achieve this by first giving a background on open data, discussing the key types of data out there, critically assessing the strengths and limitations, and building a how-to guide checklist for researchers to follow when approaching an open data source.
+This chapter aims to lay the groundwork for developing such skills by providing a framework to approach and meaningfully interpret and engage with open data.
+The chapter also offers a practical hands-on guide to accessing, wrangling, and analysing different sources of open data in order to draw conclusions about crime and place.
+We achieve this by first giving a background on open data, discussing the key types of data out there, critically assessing the strengths and limitations of each, and building a how-to guide checklist for researchers to follow when approaching an open data source.
 Finally, we work through an example that shows how to access, wrangle, link, and interpret open data from various sources, providing a template which can be applied in other research areas.
@@
 Ideally, it should be machine-readable and modifiable.
-For example, it should be downloadable from a website as an unlocked spreadsheet, or json file, rather than presented as a summary table in a PDF document.
-- **Re-use and Redistribution:** the data must be provided under conditions that allow free use, such as modification, separation, or compilation of the data and permit re-use and redistribution including the intermixing (merging) with other datasets.
+For example, it should be downloadable from a website as an unlocked spreadsheet, or JSON file, rather than presented as a summary table in a PDF document.
+- **Re-use and Redistribution:** the data must be provided under conditions that allow free use, such as modification, separation, or compilation of the data, and permit re-use and redistribution including the intermixing (merging) with other datasets.
 - **Universal Participation:** everyone should be allowed to use, re-use and redistribute the data and its derivatives without restriction.
@@
 In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
-
 Open data practices, and the extent to which it is a reality, vary both within and between countries.
-For example, the United States have a history of making public sector data sets openly available.
+For example, the United States has a history of making public sector data sets openly available.
 The United Kingdom can be more strict, releasing data under a licence ('Open Government Licence'), with some data requiring a fee [@kitchin2014data].
-The _Global Open Data Index_ [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by countries.
-It providing a rank which indicates how well governments across the world follow open data practices on aggregate, but also according to specific domains.
+The _Global Open Data Index_ [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by country.
+It provides a rank which indicates how well governments across the world follow open data practices on aggregate, but also according to specific domains.
 Many of these have direct relevance to crime and place research, such as administrative boundaries and geographic locations (e.g.
@@
 
-### Public sector (including publicly funded research).
+### Public sector
 
@@
 In some countries, the national statistics body is independent from government, which serves to improve accountability and increase public trust in the data being collated and published.
-The _Global Open Data Index_ introduced earlier focuses on identifying the gaps in governmental organisations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
+The _Global Open Data Index_ introduced earlier focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
 
@@
 Many variables thought to explain the non-random distribution of crime, offenders and victims in space are collated and published by governmental bodies.
-For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighbourhoods' - a theoretically important scale in criminology.
+For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighborhoods' - a theoretically important scale in criminology.
 Data describing urban structure, such as street networks, are often generated and distributed by public bodies, and can be used in street syntax analysis to assess whether the configuration of streets dictates things like violent crime victimisation [@streetsyntax2017].
-In many countries, victimisation surveys are openly available for download.
+In many countries, representative victimisation surveys are openly available for download.
 This is not only useful for substantive research (e.g.
@@
 The aims and objectives of private companies, who have a duty to shareholders and operate in a competitive market environment, differ considerably from local and central government.
-That said, some datasets are released openly by private sector organisations, albeit often only a subset of what would be available as a paying customer.
+That said, some datasets are released openly by private sector organizations, albeit often only a subset of what would be available as a paying customer.
 The website for ArcGIS, a piece of software maintained by the Environmental Systems Research Institute (ESRI), a private company, host a number of open geographic datasets on their online cloud platform (https://hub.arcgis.com/search).
-Indeed, many papers have made use of data collected or distributed by private organisations to explore crime and place, such as Google Street View [@langton2017residential] and Twitter [@malleson2015impact].
-In this book you will learn to use the free Twitter API for such purposes in criminology (see Section 3, Topic 7).
+Indeed, many papers have made use of data collected or distributed by private organizations to explore crime and place, such as Google Street View images [@langton2017residential] and Twitter [@malleson2015impact].
+In fact, you will learn to use the free Twitter API for such purposes in this book (see Chapter 6, Topic 7).
 
@@
 
-Finally, there is a specific group of open data sources that collate data collectively, generated by large groups of individuals who do not specifically belong to any organisation, but instead work together on a collaborative project.
+Finally, there is a specific group of open data sources that collate data collectively, generated by large groups of individuals who do not specifically belong to any organization, but instead work together on a collaborative project.
 These are crowdsourced data sets.
 Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, and Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
-Often, the organisations who maintain and monitor these data collection activities are charities or non-governmental organisations that do not operate for profit, but instead provide some form of social good.
+Often, the organizations who maintain and monitor these data collection activities are charities or non-governmental organizations that do not operate for profit, but instead provide some form of social good.
 An example is the online reporting platform FixMyStreet (https://www.fixmystreet.com/), where people can report problems such as instances of graffiti, vandalism or environment issues.
 The data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
-Another chapter in this book (see Section 3, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analysing open data of this type.
+Another topic in this book (see Chapter 6, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analysing open data of this type.
 
@@
 All sorts of open data remain unanalyzed, and or inaccessible, to large portions of the population, simply because they lack the skills and know-how.
-Although this is certainly an obstacle for data usage more generally, it represents a key challenge to the governmental bodies, in particular, on which there is an onus to ensure transparency and facilitate scrutiny.
+Although this is certainly an obstacle for data usage more generally, it represents a key challenge to governmental bodies, in particular, on which there is an onus to ensure transparency and facilitate scrutiny.
 Private companies, especially those who generate open data as a by-product of their primary activity (e.g.
@@
 
-Secondly, despite the merits of open data in terms of remedying pitfalls in traditional survey methods, there are new threats to validity that may emerge.
+Secondly, despite the merits of open data in terms of remedying pitfalls in traditional survey methods, there are other threats to validity that may emerge.
 For instance, whilst many online resources offer an 'honest' depiction of society, uncaptured by surveys, researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
-This could be a result of individuals willingly managing and 'curating' their online presence, or because of wider issues such as government censorship, or cultural norms around particular topics such as sexual assault.
-Relatedly, researchers should be aware of issues over sampling, and ultimately the generalizability of findings that emerge from the analysis of open data.
-In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduced a degree of bias [@longley2012geodemographics].
+This could be a result of individuals willingly managing and 'curating' their online presence [@ellison2006managing], or because of wider issues such as government censorship, or cultural norms around particular topics such as sexual assault.
+Relatedly, researchers should be aware of issues over sampling, and ultimately, the generalizability of findings that emerge from the analysis of open data.
+In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduces a degree of bias [@longley2012geodemographics].
 Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
@@
 In crime and place research, both dependent variables (e.g.
-police recorded crime incidents, victimisation rates) and independent variables (e.g.
+police recorded crime incidents, victimization rates) and independent variables (e.g.
 demographic characteristics, ambient population estimates) can be sourced from open data, whether public, private or crowdsourced.
@@
 - **Who** is represented in the data, and who is excluded?
-- **What** concepts and constructs can and cannot be operationalised with this data?
+- **What** concepts and constructs can and cannot be operationalized with this data?
 - **When** did data collection take place, and how might have that influenced results?
 
-With this in mind, the following exercise utilizes two different sources of open data on public transport and police recorded crime in London, England.
+With this in mind, the following exercise utilizes mutliple sources of open data on public transport and police recorded crime in London, England, to explore crime and place.
 
@@
 The first is a substantive empirical investigation into crime in and around London Underground stations, in the spirit of existing research into transport crime (see newton2014above).
-Studies in Sweden, for instance, have shown the importance of environmental and neighbourhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
+Studies in Sweden, for instance, have shown the importance of environmental and neighborhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
 Investigations have also taken place into crime concentrations at bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
@@
 To access them we will use three different methods, namely, (1) direct download, (2) direct request to an API, 3) request to an API using a wrapper.
-This will give you few different ideas about how you might go about accessing other open data sets relevant to your research.
+This will give you a few different ideas about how you might go about accessing other open data sets relevant to your research.
 Locating, identifying and learning to access open data is a skill in itself.
@@
 This is a tool which defines an interface for a program to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
-Here, we are using the term 'API' to denote tools created by an open data provider to provide access to different subsets of their content.
+Here, we are using the term 'API' to denote tools created by an open data provider to give access to different subsets of their content.
 APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
@@
 
-Here, we will use the example of a local government organisation responsible for transport, Transport for London (TfL).
+Here, we will use the example of a local government organization responsible for transport, Transport for London (TfL).
 One of the main responsibilities of TfL is to oversee various rail networks in Greater London, including the London Underground.
@@
 Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.
-Usefully, TfL provide a documentation page with examples of how you can use the HTTP links to access make calls to the API (https://api-portal.tfl.gov.uk/docs).
-We will actually make these calls from within the R environment using a 'wrapper', detailed below, but it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
+Usefully, TfL provide a documentation page with examples of how you can use the HTTP links to make calls to the API (https://api-portal.tfl.gov.uk/docs).
+We will actually make these calls from within the R environment, but it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
 
@@
 This is called a _wrapper_ because it uses code that 'wraps' around the API to make it a neater, more usable package.
-Wrappers remove (or at least lower) many of the obstacles to accessing open data noted earlier, particularly for people from a non-technical background.
+Wrappers remove (or at least lower) many of the obstacles to accessing open data noted earlier.
 The wrapper can take many forms, such as a Python module, or an R package.
@@
 As such, it is a prime example of 'crowdsourced' open data.
-You an view the information contributed to Open Street Map using their online mapping platform (https://www.openstreetmap.org/).
-The result of people's contributions is database of spatial information rich in local knowledge which provides invaluable information about places and their features, without being subject to strict terms on usage.
+You can view the information contributed to Open Street Map using their online mapping platform (https://www.openstreetmap.org/).
+The result of people's contributions is a database of spatial information rich in local knowledge which provides invaluable information about places and their features, without being subject to strict terms on usage.
 Open Street Map has two types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
-We will explore the APIs using both, but focus on their usage from within R.
+We will focus on their usage from within R, as this is a tool used more widely to conduct reproducible research and statistics in criminology.
 
@@
 So, let's get started.
-Recall that are interested in examining the extent to which crime concentrates in and around London Underground stations, but also the contrasts between the open data available via TfL and Open Street Map, as public sector and crowdsourced open data, respectively.
+Recall that we are interested in examining the extent to which crime concentrates in and around London Underground stations, but also the contrasts between the open data available via TfL and Open Street Map, as public sector and crowdsourced open data, respectively.
 
 First, ensure that you have the relevant packages installed in R.
-Although the main packages used in this demonstration are `rjson` for parsing the TfL API results and `osmdata` for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation.
-If you don't have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
+Although the main packages used in this demonstration are `rjson` for parsing the TfL API results and `osmdata` for querying the Open Street Map API, we use a number of additional packages for data handling and visualization.
+If you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
 
@@
 library(tidyr)
-library(purrr)
 library(patchwork)
@@
 To draw conclusions about the degree to which crime occurs in and around London Underground stations, we first need spatial information so we can plot station locations on a map.
-In the documentation there in fact an 'API area', which offer some guidance on the types of data available, called _Stops_.
-We can see that this include information on bus stops by route (which are numbered e.g.
+In the documentation there an 'API area' (which offer some guidance on the types of data available) called _Stops_.
+We can see that this includes information on bus stops by route (which are numbered e.g.
 24), and underground stations by line (which are named e.g.
@@
 The example for the underground stations is a URL: https://api.tfl.gov.uk/line/bakerloo/stoppoints.
-This has quite an intuitive structure, since we first want to access the TfL API, look for data by line, select the name of the line we are interested in, and finally pull information about where stations stops.
+This has quite an intuitive structure, since we first want to access the TfL API, look for data by line, select the name of the line we are interested in, and finally pull information about where the train stops.
 In fact, you can simply copy and paste this URL into your internet browser, and it should look something like Figure 1.
@@
 JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
-As you can see, it is intuitive to read in its current format.
+As you can see, it is not intuitive to read in its current format.
 In R, we can use the `fromJSON()` function from the `rjson` package to parse all this information into a data frame (with rows and columns).
@@
 
-To keep things focused, let's request data about stop points on the Northern Line. Note how this is simply an amended version of the example provided by TfL in the API documentation.
+To keep things focused, let's request data about stop points on the Northern line, one of the largest and busiest on the network.
+Note how this is simply an amended version of the example provided by TfL in the API documentation.
 
 ```{r getapi}
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
+api_call <- fromJSON(txt = readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
 ```
@@
 Moreover, we can see that the information we are interested in (namely, the location of underground stations) has already been successfully parsed into columns called **lon** (longitude) and **lat** (latitude), along with an identification column of the station names called **commonName**.
-We can extract these columns, and create an `sf` point object using these coordinates, all in one chunk of code, using the piping operator `%>%` in the `dplyr` package.
+We can extract these columns, and create an `sf` (spatial) point object using these coordinates, all in one chunk of code, using the piping operator (`%>%`) in the `dplyr` package.
 Note that we define the Coordinate Reference System (CRS) which the data has come in (WGS 84), and then transform it into something more appropriate (the British National Grid, a projected CRS used in the United Kingdom).
@@
 
-There we have it: with just a few lines of code in R, we have queried the Transport for London API, a public sector source of open data, and plotted a basic map of stations on the Northern underground line.
+There we have it: with just a few lines of code in R, we have queried the TfL API, a public sector source of open data, and plotted a basic map of stations on the Northern underground line.
 
@@
 This approach differs from the TfL request in that our data source, Open Street Map, has a wrapper for the API built into an R package called `osmdata`.
-As such, rather than the API query being made directly (like we did for TfL, using the URL), we can use R code.
+As such, rather than the API query being made directly (like we did for TfL, using the URL), we can use R functions available from a bespoke package.
 
@@
 
-Now we have our study region, we can scrape data from the OSM API using the `opq()` function.
+Now we have our study region, we can scrape data from the Open Street Map API using the `opq()` function.
 The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
-A common way of doing this is using 'Overpass turbo' (https://overpass-turbo.eu/) which allows you to query the database alongside an interactive map which automatically updates based on your results.
+A common way of doing this outside of R is by using 'Overpass turbo' (https://overpass-turbo.eu/) which allows you to query the database alongside an interactive map which automatically updates based on your results.
 In the interests of consistency, we will make these queries from within R using the `osmdata` wrapper.
@@
 Keys are used to describe a broad category of features (e.g.
-highway, amenity), and values are more specific descriptions of each key (e.g.
+highway, amenity), and values are more specific descriptions (e.g.
 cycleway, bar).
@@
 
-We could visualize these points using `ggplot2`, just as we did earlier, but are currently working with any form of public transport labeled as a 'station' by contributors to Open Street Map.
+We could visualize these points using `ggplot2`, just as we did earlier, but we are currently working with any form of public transport labeled as a 'station' by contributors to Open Street Map.
 Exploring the `osm_stat_sf` object will reveal a number of additional variables which we can use to narrow down our selection, such as the _network_ and _line_ on which the station is located.
@@
 For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
-It is worth noting that this definition is somewhat arbitrary, and is subject to a degree of spatial inaccuracy is open police recorded crime data [see tompson2015uk], but serves to facilitate this demonstration.
+It is worth noting that this definition is somewhat arbitrary, and is subject to a degree of spatial inaccuracy is open police recorded crime data [see tompson2015uk], but it does serve to facilitate this demonstration.
 
@@
 
-We can then visualise these counts and compare the two sources of data by colouring in our point locations according to the crime count.
+We can then visualize these counts and compare the two sources of data by coloring in our point locations according to the crime count.
 Note that we arrange the plots side-by-side using syntax available using the `patchwork` package.
@@
 p1 <- ggplot(data = osm_north_sf) +
-  geom_sf(mapping = aes(colour = crimes), size = 2) +
+  geom_sf(mapping = aes(color = crimes), size = 2) +
   labs(title = "Open Street Map") +
-  scale_colour_viridis_c()
+  scale_color_viridis_c()
 
 p2 <- ggplot(data = tfl_north_sf) +
-  geom_sf(mapping = aes(colour = crimes), size = 2) +
+  geom_sf(mapping = aes(color = crimes), size = 2) +
   labs(title = "Transport for London") +
-  scale_colour_viridis_c()
+  scale_color_viridis_c()
 
@@
 A number of factors emerge from this visualization.
-First, issues arising disparity between the stations identified on the Northern line are compounded.
-Using the Open Street Map data in isolation, we are grossly underestimating the the number of crimes occurring in and around Northern line stations.
+First, issues arising from the disparity between the stations identified on the Northern line are compounded.
+Using the Open Street Map data in isolation, we are grossly underestimating the number of crimes occurring in and around Northern line stations.
 This could have significant implications for how police allocate resource across the underground network.
-Secondly, the spatial patterning of crime on underground stations differs considerably between the two data sources.
-Using the Open Street Map data, the final station on the line, Morden, has the greatest concentration of crimes, whereas the TfL data suggests this is in the city centre, at Tottenham Court Road.
+Secondly, as a result of this, the spatial patterning of crime on underground stations differs considerably between the two data sources.
+Using the Open Street Map data, the final station on the line, Morden, has the greatest concentration of crimes, whereas the TfL data suggests that the most problematic station is Tottenham Court Road, in the city centre.
 The conclusions drawn are starkly different, and would determine the consistency of findings with existing research examining crime at underground stations [@ceccato2013security].
@@
 Secondly, as noted, systematic variation in the extent to which individuals have contributed to the database can generate interesting discussions and avenues of research in its own right.
-Thirdly, this is simply one of thousands of features contained in the Open Street Map database.
-If we were interested in examining concentrations of bicycle theft from bike rental docking stations (so-called "Boris bikes") in Greater London, we could query open data APIs from TfL and Open Street Map, just as we did earlier, but using slightly amended criteria.
+Thirdly, this is simply one of thousands of features contained in the Open Street Map database, many of which are more complete.
+If we were interested in examining concentrations of bicycle theft from bike rental docking stations (so-called "Boris bikes") in Greater London, we could query the APIs from TfL and Open Street Map, just as we did earlier, but using slightly amended criteria.
 This is demonstrated below, all in one code chunk.
@@
 
-As showcased, the resulting visualizations are highly similar (a Spatial Point Pattern Test will confirm this, see Section 4 Topic 4).
+As showcased, the resulting visualizations are highly similar (a Spatial Point Pattern Test will confirm this, see Chapter 8, Topic 4).
 Refer back to the example earlier if some of the code is unclear.
-Note that this is a fantastic benefit of using an API wrapper from within R, because the code used to query each open data source can easily be re-used, with minor amendments, to extract different features.
+Note that this is a major benefit of using an API wrapper from within R, because the code used to query each open data source can easily be re-used, with minor amendments, to extract different features.
 
@@
 We have defined open data, and outlined some key types of open information available, along with their respective strengths and weaknesses.
-An important component of this review has been encourage an engaged and critical research approach to using open data.
+An important component of this review has been to encourage an engaged and critical research approach to using open data.
 Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources.
@@
 
-- Threats to its sustainability (e.g. licence expiry).
-- Prospects in crime of place research.
-- Examples of cool projects (e.g. Colouring London).
-- Suggestions for new avenues which can expand the field.
+As it stands, open data represents a key resource in crime and place research.
+It is a fundamental component in the movement towards transparent and reproducible scientific research.
+The skills required to effectively access and use open data, such as those demonstrated in this chapter, are increasingly being taught and deployed in social science.
+The number and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology.
+For instance, information which traces building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colourin], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
+But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go.
+As with any emerging domain, open data sources vary considerably in their degree of completeness and reliability.
+Whilst this can offer insight, for instance, in terms of public participation and sense of community, it can also represent a significant obstacle to empirical analysis.
+A key challenge for the future will be to ensure that students and research practitioners ask critical questions of their data before blindly delving into analysis and public dissemination.
+Moreover, the longevity of open data licences remains an unknown parameter.
+With so much open data being distributed by the public sector, the accessibility of open information is subject to fluctuations in the amount of resource (and goodwill) available to safe-guard continuity.
+One way of justifying the continued collection and distribution of open public sector data is simply to use the data for public good, such as informing interventions which help reduce crime victimization.
+
+\newpage
 

### 2020-05-24 draft_chapter.Rmd
another read through some errors tho :(

@@
 
-Access to data in crime and place research, and even criminology more broadly, has traditionally been reserved for those who have the means to collect fresh data themselves, pay for access, or obtain data through formal data sharing agreements.
+Access to data in crime and place research has traditionally been reserved for those who have the means to collect fresh data themselves, pay for access, or obtain data through formal data sharing agreements.
 Even when access is granted, the usage of these data often comes with conditions that circumscribe how the data can be used through licensing or policy [@kitchin2014data].
-The public dissemination of findings which emerge from analysis might even be subject to restrictions.
+Even the public dissemination of findings which emerge from analysis might be subject to restrictions.
 This can lead to unequal access, controlled usage and curb the diffusion of findings, severely limiting the insight that can be attained from data.
 
-Open data are a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements and negotiations [@manovich2011trending].
-Not only this, but open data can also lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
-As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the social world around them have the skills and know-how to access, interpret, critique and analyse open data sets.
+Open data initiatives provide a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements, and negotiations [@manovich2011trending].
+Open data can lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
+As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyse open data sets.
 
-This chapter aims to lay the groundwork for developing such skills by providing a framework to approach and meaningfully interpret and engage with open data.
-The chapter also offers a practical hands-on guide to accessing, wrangling, and analysing different sources of open data in order to draw conclusions about crime and place.
-We achieve this by first giving a background on open data, discussing the key types of data out there, critically assessing the strengths and limitations of each, and building a how-to guide checklist for researchers to follow when approaching an open data source.
+This chapter will outline how to develop such skills by providing a framework to approach and meaningfully interpret open data.
+The chapter also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.
+
+<!-- We achieve this by first giving a background on open data, discussing the key types of data out there, critically assessing the strengths and limitations of each, and building a how-to guide checklist for researchers to follow when approaching an open data source.
 Finally, we work through an example that shows how to access, wrangle, link, and interpret open data from various sources, providing a template which can be applied in other research areas.
+-->
 
@@
 - **Universal Participation:** everyone should be allowed to use, re-use and redistribute the data and its derivatives without restriction.
-For instance, restrictions that only allow non-commercial or educational usage would not constitute open data.
+For instance, restrictions that only allow non-commercial or educational usage would *not* constitute open data.
 
-In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
+>In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
+
 Open data practices, and the extent to which it is a reality, vary both within and between countries.
@@
 The _Global Open Data Index_ [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by country.
-It provides a rank which indicates how well governments across the world follow open data practices on aggregate, but also according to specific domains.
+It provides a rank which indicates how well governments across the world follow open data practices across varoius domains.
 Many of these have direct relevance to crime and place research, such as administrative boundaries and geographic locations (e.g.
 coordinates).
-
-## What are types of open data?
-
-Now we have a reasonable idea about what open data is in a broad sense, we can consider the different types of open data which you might come across.
-Here, we formulate a typology of open data sets based on their origin to help provide an overview.
-
-### Public sector
-
-Much of the open data movement has focused on information generated by local and national state agencies, or publicly funded research.
-Open data has become a key component of transparency and accountability in government, given that much of the data collected has been funded by the public purse [@kitchin2014data].
-In some countries, the national statistics body is independent from government, which serves to improve accountability and increase public trust in the data being collated and published.
-The _Global Open Data Index_ introduced earlier focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
-
-As researchers studying crime and place, public sector data is particularly important.
-Many variables thought to explain the non-random distribution of crime, offenders and victims in space are collated and published by governmental bodies.
-For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighborhoods' - a theoretically important scale in criminology.
-Data describing urban structure, such as street networks, are often generated and distributed by public bodies, and can be used in street syntax analysis to assess whether the configuration of streets dictates things like violent crime victimisation [@streetsyntax2017].
-In many countries, representative victimisation surveys are openly available for download.
-This is not only useful for substantive research (e.g.
-quantifying long-term crime trends), but also scrutiny over reliability and accuracy in other forms of open data, such as police recorded crime records.
-Both of these examples represent an interesting challenge to open data principles, since data tends to be anonymised to some degree before being made available to the public.
-Access to the raw records is dependent on a number of requirements (e.g.
-background checks, contractual agreements, ICT security measures) which might exclude many individuals due to cost and practicality.
 
@@
 
+## What are types of open data?
+
+Now we have a reasonable idea about what open data is in a broad sense, we can consider the different types of open data which you might come across.
+Here, we formulate a typology of open data sets based on their origin.
+
+### Public sector
+
+The Public Sector refers to organizations owned and operated by local or federal government with the core aim to provide services to the public.
+Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data].
+These can include national statistics, or administrative data.
+Another set of data included in this group is data collected by publicly funded research, for example victimisation surveys, housing surveys, and others.
+
+The _Global Open Data Index_ introduced earlier focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
+
+Specific to crime and place research, data sets of interest in this particular domain might include
+<!-- As researchers studying crime and place, public sector data is particularly important.
+Many variables thought to explain the non-random distribution of crime, offenders and victims in space are collated and published by governmental bodies.
+For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighborhoods' - a theoretically important scale in criminology.
+-->
+data describing urban structure, such as street networks to assess whether the configuration of streets dictates things like violent crime victimisation [@streetsyntax2017], or victimisation surveys mentioned earlier to quantify citizens' perceived safety and security.
+
+<!-- In many countries, representative victimisation surveys are openly available for download.
+This is not only useful for substantive research (e.g.
+quantifying long-term crime trends), but also scrutiny over reliability and accuracy in other forms of open data, such as police recorded crime records.
+Both of these examples represent an interesting challenge to open data principles, since data tends to be anonymised to some degree before being made available to the public.
+Access to the raw records is dependent on a number of requirements (e.g.
+background checks, contractual agreements, ICT security measures) which might exclude many individuals due to cost and practicality.
+-->
+
 ### Private sector
 
+Contrary to public sector, private sector refers to the rest of the economy whihc are not under direct state control.
 Opening up data generated by the private sector represents a significant challenge, largely due to the proprietary value to its creators [@kitchin2014data].
 The aims and objectives of private companies, who have a duty to shareholders and operate in a competitive market environment, differ considerably from local and central government.
+
 That said, some datasets are released openly by private sector organizations, albeit often only a subset of what would be available as a paying customer.
@@
 These are crowdsourced data sets.
-Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, and Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
+Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, or Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
 Often, the organizations who maintain and monitor these data collection activities are charities or non-governmental organizations that do not operate for profit, but instead provide some form of social good.
@@
 applying a framework for its analysis), and computational skills in data scraping, wrangling and cleaning, in order to transform it into a usable format for research [@boyd2012critical].
-All sorts of open data remain unanalyzed, and or inaccessible, to large portions of the population, simply because they lack the skills and know-how.
+All sorts of open data remain inaccessible to people who may lack the skills and know-how to acquire them.
 Although this is certainly an obstacle for data usage more generally, it represents a key challenge to governmental bodies, in particular, on which there is an onus to ensure transparency and facilitate scrutiny.
@@
 Moreover, in both public and private spheres, licences and conditions on open data are not necessarily concrete, and might be subject to change with little or no notice.
-This is an important consideration when planning and running long-term research projects.
+This is an important consideration when planning and running long-term research projects which involve open data.
 
@@
 For instance, whilst many online resources offer an 'honest' depiction of society, uncaptured by surveys, researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
-This could be a result of individuals willingly managing and 'curating' their online presence [@ellison2006managing], or because of wider issues such as government censorship, or cultural norms around particular topics such as sexual assault.
+This could be a result of individuals willingly managing and 'curating' their online presence [@ellison2006managing], or because of wider issues such as government censorship, or cultural norms around particular topics, particularly sensitive topics of interest to researchers of crime, such as sexual assault or drug use.
+
 Relatedly, researchers should be aware of issues over sampling, and ultimately, the generalizability of findings that emerge from the analysis of open data.
@@
 Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
-Contributions to resources such as Open Street Map, which we look at in the practical exercise later, are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
+Contributions to resources such as Open Street Map (which we look at in the practical exercise later) are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
 These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
+
+### What can be done?
 
@@
 demographic characteristics, ambient population estimates) can be sourced from open data, whether public, private or crowdsourced.
+
+So what can we do to make sure we make good use of these data?
 With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science.
@@
 
-With this in mind, the following exercise utilizes mutliple sources of open data on public transport and police recorded crime in London, England, to explore crime and place.
+The answers to these questions will put the data sources in a context of understanding, and ensure that the researchers use them appropriately and with care.
+
+With this in mind, we now move on to a practical exercise, where we will utilize mutliple sources of open data, and explore police recorded crime on and around public transport in London, England.
 
@@
 
-In this exercise we will demonstrate a number of different skills, including:
+In this exercise you will acquire a number of different skills, including:
 
-- Accessing open data using direct downloads and Application Programming Interfaces (APIs).
-- Cleaning and wrangling data for analysis.
-- Visualizing open spatial data.
+- Accessing open data using 3 different methods: 
+    + direct downloads 
+    + direct calls to an Application Programming Interface (API)
+    + using an API wrapper.
+- Cleaning, wrangling, and visualizing open spatial data.
 - Comparing different sources of open data.
-- Engaging with the critical questions outline above to better understand open data.
+- How to engage with the critical 'Where', 'Why', 'Who', 'What', 'When' questions to better understand open data.
 
@@
 
-In demonstrating these skills, we will be answering two key research questions.
-The first is a substantive empirical investigation into crime in and around London Underground stations, in the spirit of existing research into transport crime (see newton2014above).
+In this exercise we will explore crime in and around London Underground stations.
+We know that there are important environmental features when it comes to public transport areas being more or less criminogenic environments.
 Studies in Sweden, for instance, have shown the importance of environmental and neighborhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
-Investigations have also taken place into crime concentrations at bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
-As such, we can pose the following research question:
+Similar work has been done exploring bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
 
-- RQ1: To what extent does crime cluster in and around London Underground stations?
+Now we will consider the case of London. Specifically, we will examine to what extent does crime cluster in and around London Underground stations?
 
-The second is more methodological, and relates to our considered and critical usage of open data, by exploring the validity of different open data sources. As such, we can pose:
-
-- RQ2: How comparable are public sector and crowdsourced open data on London underground stations?
-
-First, we will go through how to access the open data used in the exercises.
+We will use various sources of open data to answer this question, and this will allow us to explore the strengths and limitations across public sector and crowdsourced open data sources.
 
@@
 The simplest way that open data can be made available is through direct download from a website.
-This is the case for open police recorded crime data in England and Wales.
-It can be accessed and downloaded using an online web portal (https://data.police.uk/) under Open Government Licence.
-Using the "Data" tab, anyone from researchers to members of the public can select the time period and police force of interest, and the type of information required (e.g.
-crimes, stop and search, outcomes), and then "Generate file" to create a download link.
-Data is downloaded in  comma-separated values format (.csv file extension).
-The website also contains detailed information about data quality, downloads and tools under the "About" tab, which is particularly useful when considering the critical questions posed earlier (where, why, who, what, when).
+In such a case, you can visit a website, select some parameters, and save a file containing the data you requested locally on your computer.
+
+In the United Kingdom, police recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence.
+
+Visit the website above, and you will see a welcome message, and 6 tabs across the top of it, which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
+
+Before we download any data, we can learn more about it by clicking on the "About" tab.
+This brings up a lot of information that it is important to carefully read through to be able to answer the where, why, who, what, when questions posed earlier.
+Take a moment to read through, and try to take notes on what you think might be relevant for your analysis.
+For example, if we want to map crimes, we might want to explore if there is any type of "anonymisation" that might take place before the data are released (to protect the privcy of the victims for example).
+If you read the about page here, you might find the following note:
+
+> Location Anonymisation
+The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.
+
+This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate, and this may have implications for our findings later!
+
+To then actually download some data, move on to clicking on the "Data" tab.
+This should open for you a page titled "Data downloads" where you can see another 5 tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".
+Staying on this "Custom download" page you can make some selections on what sort of data you wanted to download.
+
+We can select the time period and police force of interest, and the type of information required (e.g.
+crimes, stop and search, outcomes).
+We are also informed that the data are downloaded in comma-separated values format (.csv file extension), which is in line with machine-readable, easy to manipulate data format requirements.
+
 For this exercise, we are going to use British Transport Police data, a force which operates on railways and light-rail systems across the country, for the month of January in 2020.
 Select the time period using the dropdown menus, the force using the tickboxes, and then generate and download the file.
-We save this file locally to a subfolder named "data", and during the exercise, the data is loaded into R from this folder.
+
+Save this file locally (in your working directory) to a subfolder named "data", and then you can load it in with the `read_csv()` command from the `readr` package (remember to load it with `library(readr)`).
+
+```{r crimeload1}
+library(readr)
+btp_df <- read_csv("data/2020-01-btp-street.csv")
+```
+
+Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our github page:
+
+```{r crimeload_gh, eval = FALSE}
+btp_df <- read_csv("https://raw.githubusercontent.com/langtonhugh/osm_crim/master/data/2019-01-greater-manchester-street.csv")
+```
+
+You should have now a dataframe with `r nrow(btp_df)` crimes in them ready for you to explore!
 
@@
 Another way of downloading open data is through an application programming interface (API).
-This is a tool which defines an interface for a program to interact with a software component or system, for example, what sort of requests or calls can be made to it, and how these can be made.
+This is a tool which defines an interface for a program to interact with a software component, for example, what sort of requests or calls can be made to it, and how these can be made.
 Here, we are using the term 'API' to denote tools created by an open data provider to give access to different subsets of their content.
-APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
+Such APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
 APIs can take many different forms and be of varying quality and usefulness [@foster2016big].
@@
 
-Here, we will use the example of a local government organization responsible for transport, Transport for London (TfL).
-One of the main responsibilities of TfL is to oversee various rail networks in Greater London, including the London Underground.
-We will use open data from TfL to practice how to construct HTTP (i.e.
-web link) requests via their API, and then parse the data which is returned.
-Much like how downloading data directly requires access to a data dictionary to make sense of the data (e.g.
-understanding its origins and flaws), we must find a similar document to understand how the API we plan on using works.
+Here we will make a direct request to the API created by London's local government organization responsible for transport: Transport for London (TfL).
+TfL oversee the London Underground, and provide access to their open data through their unified API.
+Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What', 'When' questions, we must find a similar document here.
 Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.
-Usefully, TfL provide a documentation page with examples of how you can use the HTTP links to make calls to the API (https://api-portal.tfl.gov.uk/docs).
-We will actually make these calls from within the R environment, but it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
+
+They further provide a documentation page with examples of how you can use HTTP (i.e.
+web link) requests to make calls to the API (see: https://api-portal.tfl.gov.uk/docs).
+If you follow this link you should see a page titled "Our Unified API" and some examples of calls.
+
+To draw conclusions about crimes at London Underground stations, we will need some data specifically about these underground stations.
+
+In the documentation there 'API area' (which offer some guidance on the types of data available).
+Scroll down and find the example calle for stops _Stops_.
+In the description you will  see that this call returns information on stops for busses or London Underground lines, and there are two examples (bus route 24 and the Bakerloo line).
+Take their demonstration url you see for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser.
+When you visit this page you should see something like Figure 1:
+
+```{r tfl, out.width = "\\textwidth", fig.cap="example of data from Transport for London", echo = F}
+knitr::include_graphics("img/tfl_screenshot.png")
+```
+
+Although this contains the information we requested, it is not very legible for our human eyes.
+It in a format called _JSON_.
+This stands for JavaScript Object Notation.
+JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
+As you can see, it is not intuitive to read in its current format but it is very machine readable friendly, once again in line with this requirement for open data.
+
+Now you won't usually make these calls by pasting them into a web browser, but we wanted for you to see how it works.
+In readlity, we will actually make these calls from within the R environment, but it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
+For example, we now know that to request the stops from the "Bakerloo" line, we need the url:
+
+`"https://api.tfl.gov.uk/line/bakerloo/stoppoints"`
+
+Let's say instead we want data for the Northern Line. What do you think that URL will look like ?
+
+Indeed, all you need to do is replace "bakerloo" with "northern", and you got the stops for this route instead. Now let's see how we would make this call in R.
+
+In R, we can use the `fromJSON()` function from the `rjson` package to parse all this information into a data frame (with rows and columns).
+It will then become more familiar and usable.
+All we need to do is input the URL from the TfL API into the function, with a bit of help from `readLines()`, another function in base R which is used to read text from URLs.
+
+To keep things focused, let's request data about stop points on the Northern line, one of the largest and busiest on the network.
+Note how this is simply an amended version of the example provided by TfL in the API documentation.
+
+```{r getapi}
+library(rjson)
+
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
+```
+
+This gives us an object (`api_call`) which contains all the information returned by the TfL API.
+JSON is slightly different to traditional data frames with rows and columns, which we are probably more familiar with, because the data is nested.
+For instance, `api_call` is classed as a data frame, but just try  viewing the object using `View(api_call)`.
+
+You will notice that you see a list of lists, which is what we call nested lists.
+Another way of checking what sort of object this `api_call` object is is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
+
+If you run this, you will see that each element in `api_call` is a "list".
+
+This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated, not always a neatly formatted dataframe like the csv from the police.uk open data portal.
+But we will learn how to manage this data into a more familiar format later.
 
@@
 
-Often, developers who work with APIs will share their code, and release them in the form of a package or module.
+Finally, the last way we will look at how open data is often shared is by using API wrappers.
+Often, developers who work with APIs will share their code, and release them in the form of a package or module, so that other people can use it too.
 This is called a _wrapper_ because it uses code that 'wraps' around the API to make it a neater, more usable package.
@@
 The result of people's contributions is a database of spatial information rich in local knowledge which provides invaluable information about places and their features, without being subject to strict terms on usage.
+
 Open Street Map has two types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
-We will focus on their usage from within R, as this is a tool used more widely to conduct reproducible research and statistics in criminology.
 
-## Walk-through
+If we load and use the package `osmdata` we can use its funcitons to query the Open Street Map API, rather than the API query being made directly (like we did for TfL, using the URL).
+Once again, we want to have some documentation to be able to refer to about our data and the ways to query it.
+When it comes to OSM, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about).
+Regarding the wrapper, as it is an R package, it will come with a handy vignette, which in this case we can find at [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).
 
-So, let's get started.
-Recall that we are interested in examining the extent to which crime concentrates in and around London Underground stations, but also the contrasts between the open data available via TfL and Open Street Map, as public sector and crowdsourced open data, respectively.
+Unlike the TfL API, `osmdata` is an international database, and has lots of data that we might not necessarily need.
+The first thing we need to specify therefore, is where we want to get data from.
+To do this, we use the `getbb()` function from the `osmdata` package, which stands for "get bounding box".
+You can think of the bounding box as a box drawn around an area that we are interested in (in this case, London, UK) which tells the OSM API that we want everything *inside* the box, but nothing *outside* the box.
 
-First, ensure that you have the relevant packages installed in R.
-Although the main packages used in this demonstration are `rjson` for parsing the TfL API results and `osmdata` for querying the Open Street Map API, we use a number of additional packages for data handling and visualization.
-If you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
+So how can we name a bounding box specification to define the study region?
+This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, or you can use a search term.
+Here, we want to select Greater London, in the UK so we can use the search term "greater london united kingdom".
+Besides specifying where we want data from, we can also tell the `getbb()` function what format we want the data in.
+Let's say in this case we want a spatial object, specifically an `sf` polygon object which we name `bb_sf`:
+
+```{r bb}
+library(osmdata)
+
+bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon")
+```
+
+We now have our study region defined as the administrative boundaries of Greater London, which we can visualize quickly to check it has worked as expected.
+To visualise we can use the `sf` and `ggplot2` packages.
+
+```{r bbmap}
+library(ggplot2)
+library(sf)
+
+ggplot(data = bb_sf) +
+  geom_sf()
+```
+
+If you are familiar with the shape of London then this image should be familiar as well!
+Pleased with our bounding box, we can now move on to scrape data from the Open Street Map API using the `opq()` function.
+The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
+
+Besides specifying what area we want to query with our bounding box object (`bb_sf`) in the `opq()` function, we must also define the feature which we want to pull from the API.
+Features in Open Street Map are defined through 'keys' and 'values'.
+Keys are used to describe a broad category of features (e.g.
+highway, amenity), and values are more specific descriptions (e.g.
+cycleway, bar).
+These are tags which contributors to Open Street Map have defined.
+A useful way to explore these is by using the comprehensive Open Street Map Wiki page on map features (https://wiki.openstreetmap.org/wiki/Map_Features).
+
+We do this using the `add_osm_feature()` function, specifying our key as 'public transport' and our value as 'station'.
+
+We also want to specify what sort of object (what class) to get our data into, and as we're still working with spatial data we stick to the `sf` format, for which the function is `osmdata_sf()`.
+
+And finally, we trim everything down into our bounding box parameters with the `trim_osmdata()` function. We add all these together using a pipe (`%>%`) and make the call like so:
+
+```{r opq}
+osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
+  add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
+  osmdata_sf() %>%                                                 # specify class
+  trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box 
+```
+
+The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the **Console**.
+
+```{r printstations}
+osm_stat_sf
+```
+
+This confirms details like the bounding box coordinates, but also provides information on the features collected from the query.
+As one might expect, most information relating to public transport station locations has been recorded using points (i.e.
+two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing.
+We also have around one hundred polygons.
+
+For now, let's extract the point information.
+
+```{r stationpoints}
+osm_stat_sf <- osm_stat_sf$osm_points 
+```
+
+We now have a simple features object with all the public transport stations in London, which the volunteers of OSM have mapped, and also 130 variables of auxiliary data they collected, such as the "fare_zone" the station is in, what "amenity" it may have, whether it has "toilets", and many many more.
+
+Of course, this is up to the volunteers whether they collect all these data, and in many cases they just might not. For example, if we look at fare zone, we might see the following:
+
+```{r, echo=FALSE}
+library(dplyr)
+
+osm_stat_sf %>% group_by(fare_zone) %>% count() %>% st_drop_geometry() %>% arrange(-n)
+
+```
+
+You can see for most of these stations (`r osm_stat_sf %>% group_by(fare_zone) %>% count() %>% st_drop_geometry() %>% filter(is.na(fare_zone)) %>% pull(n)`) we have missing data (denoted by NA).
+Nevertheless, when the details are recorded, they provide rich insight and local knowledge we may otherwise not attain.
+
+One last thing to do is to select the stations which are relevant to us (ie: they fall along the Northern Line).
+The variable `line` can help us with this.
+First let's look at the values in this variable by using the `unique()` function:
+
+```{r valuesofline}
+
+unique(osm_stat_sf$line)
+
+```
+
+You might see that Northern appears a few times, once as "Northern" but also once as "Northern City".
+As we know there is no line called "Northern City" and we can see many other combinations in these values (eg Circle;District, indicating the station serves both Circle and District lines) we can assume that this is a station that serves both Northern line and one other.
+
+To select all Northern line stations we can use the `grepl()` function to collect everything with "Northern" in the name:
+
+```{r filterline}
+osm_north_sf <- osm_stat_sf %>% 
+  filter(grepl("Northern", line))
+```
+
+And now we have all our open data sets together in one place!
+
+## Cleaning, wrangling, and visualising our data
+
+Just because we have our data doesn't necesarily mean it's in the shape we want yet.
+The first thing to do once we have acquired our open data is investigate it, using data cleaning, wrangling, and visualisation.
+
+We've already loaded the `dplyr` and `sf` and `ggplot2` packages, but here we will also make use of `tidyr` and `patchwork`, so load these (remember, if you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.)
 
 ```{r packages}
-library(jsonlite) 
-library(osmdata)
-library(ggplot2)
-library(sf)
-library(readr)
-library(dplyr)
 library(tidyr)
@@
 
-### Direct request to TfL API
+### From JSON to data frame
 
-Reading through the TfL API documentation mentioned earlier (https://api-portal.tfl.gov.uk/docs), you will get a feel for the kind of data available.
-To draw conclusions about the degree to which crime occurs in and around London Underground stations, we first need spatial information so we can plot station locations on a map.
-In the documentation there an 'API area' (which offer some guidance on the types of data available) called _Stops_.
-We can see that this includes information on bus stops by route (which are numbered e.g.
-24), and underground stations by line (which are named e.g.
-"Bakerloo").
-The example for the underground stations is a URL: https://api.tfl.gov.uk/line/bakerloo/stoppoints.
-This has quite an intuitive structure, since we first want to access the TfL API, look for data by line, select the name of the line we are interested in, and finally pull information about where the train stops.
-In fact, you can simply copy and paste this URL into your internet browser, and it should look something like Figure 1.
-
-```{r tfl, out.width = "\\textwidth", fig.cap="example of data from Transport for London", echo = F}
-knitr::include_graphics("img/tfl_screenshot.png")
-```
-
-Although this contains the information we requested, it is in a format called _JSON_.
-This stands for JavaScript Object Notation.
-JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
-As you can see, it is not intuitive to read in its current format.
-In R, we can use the `fromJSON()` function from the `rjson` package to parse all this information into a data frame (with rows and columns).
-It will then become more familiar and usable.
-All we need to do is input the URL from the TfL API into the `txt` argument of the function, with a bit of help from `readLines()` in base R which is used to read text from URLs.
-
-To keep things focused, let's request data about stop points on the Northern line, one of the largest and busiest on the network.
-Note how this is simply an amended version of the example provided by TfL in the API documentation.
-
-```{r getapi}
-api_call <- fromJSON(txt = readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
-```
-
-This gives us an object (`api_call`) which contains all the information returned by the TfL API.
-JSON is slightly different to traditional data frames with rows and columns, which we are probably quite familiar with, because the data is nested.
-For instance, `api_call` is classed as a data frame, but upon viewing the object using `View(api_call)` you will notice that there are lists nested within it.
-Another way of checking this is looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
-This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated.
+We usually want to work with data that is in the dataframe format, that is all your observations are rows and all your variables are columns.
+While this is the case for our police data and our OSM data, recall that the TfL data came in the nested JSON format.
 
@@
 Moreover, we can see that the information we are interested in (namely, the location of underground stations) has already been successfully parsed into columns called **lon** (longitude) and **lat** (latitude), along with an identification column of the station names called **commonName**.
-We can extract these columns, and create an `sf` (spatial) point object using these coordinates, all in one chunk of code, using the piping operator (`%>%`) in the `dplyr` package.
+We can extract these columns, and create an `sf` (spatial) point object using these coordinates, all in one chunk of code, using the piping operator (`%>%`).
 Note that we define the Coordinate Reference System (CRS) which the data has come in (WGS 84), and then transform it into something more appropriate (the British National Grid, a projected CRS used in the United Kingdom).
@@
 
-### Using wrappers to access Open Street Map data
+### Visualise to check and compare our data sets
 
-Now, let's explore the equivalent data using crowdsourced open data.
-This approach differs from the TfL request in that our data source, Open Street Map, has a wrapper for the API built into an R package called `osmdata`.
-As such, rather than the API query being made directly (like we did for TfL, using the URL), we can use R functions available from a bespoke package.
-
-Of course, unlike the TfL API, `osmdata` is an international database, so all queries begin with a bounding box specification to define the study region.
-This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, but it is generally easier to use a search term.
-Here, we select Greater London using the `getbb()` function, specifying that we want the content as an `sf` polygon object which we name `bb_sf`.
-
-```{r bb}
-bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon")
-```
-
-We now have our study region defined as the administrative boundaries of Greater London, which we can visualize quickly to check it has worked as expected.
-
-```{r bbmap}
-ggplot(data = bb_sf) +
-  geom_sf()
-```
-
-Now we have our study region, we can scrape data from the Open Street Map API using the `opq()` function.
-The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
-A common way of doing this outside of R is by using 'Overpass turbo' (https://overpass-turbo.eu/) which allows you to query the database alongside an interactive map which automatically updates based on your results.
-In the interests of consistency, we will make these queries from within R using the `osmdata` wrapper.
-
-First, we specify the bounding box object which is our study area (`bb_sf`), and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define the feature which we want to pull from the API.
-Features in Open Street Map are defined through 'keys' and 'values'.
-Keys are used to describe a broad category of features (e.g.
-highway, amenity), and values are more specific descriptions (e.g.
-cycleway, bar).
-These are tags which contributors to Open Street Map have defined.
-A useful way to explore these is by using the comprehensive Open Street Map Wiki page on map features (https://wiki.openstreetmap.org/wiki/Map_Features).
-Here, it is fairly intuitive, because our key is 'public transport' and our value is 'station'.
-This is not _exactly_ what we want yet, because we are only interested in underground stations, but again, this highlights an interesting obstacle when accessing open data which could serve any number of research purposes.
-Note that we specify that we want the output to be an `sf` object, and that we want to trim the queried features by our study region `bb_sf`.
-
-```{r opq}
-osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
-  add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
-  osmdata_sf() %>%                                                 # specify class
-  trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box 
-```
-
-The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the **Console**.
-
-```{r printstations}
-osm_stat_sf
-```
-
-This confirms details like the bounding box coordinates, but also provides information on the features collected from the query.
-As one might expect, most information relating to public transport station locations has been recorded using points (i.e.
-two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing.
-We also have around one hundred polygons.
-For now, let's extract the point information.
-
-```{r stationpoints}
-osm_stat_sf <- osm_stat_sf$osm_points 
-```
-
-We could visualize these points using `ggplot2`, just as we did earlier, but we are currently working with any form of public transport labeled as a 'station' by contributors to Open Street Map.
-Exploring the `osm_stat_sf` object will reveal a number of additional variables which we can use to narrow down our selection, such as the _network_ and _line_ on which the station is located.
-You will notice from running the below code, which opens up a frequency table for each combination of network and line, that we have the information required to do this.
-
-```{r freqtab, results = F, eval = F}
-osm_stat_sf %>%
-  as_tibble() %>% 
-  group_by(network, line) %>% 
-  tally() %>% 
-  View()
-```
-
-In fact, because all Northern line stations are labeled as "London Underground", we only need to filter by the line.
-
-```{r filterline}
-osm_north_sf <- osm_stat_sf %>% 
-  filter(line == "Northern")
-```
-
-Now, we can then plot these points over the station locations pulled from the TfL API. We do not plot the study region boundary in order to get a more detailed, local view.
+We can then plot these OSM station points over the station locations pulled from the TfL API, to see how the two data sets compare.
+We do not plot the study region boundary in order to get a more detailed, local view.
 
@@
 There are slight differences in the exact point locations of stations, but more saliently,  Open Street Map appears to be missing stations, especially in Central and South London.
-This could be due to a number of reasons, amongst them, variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting].
+This could be due to a number of reasons, among them: variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting].
 Given that South London is historically less wealthy in comparison to the North, for instance, this would certainly be an interesting avenue to pursue.
@@
 But, what might the impact of this be on studying crime in and around London Underground stations, given the differences observed between the two open data sources?
-For that, we can load in our British Transport Police data which we downloaded directly from the web earlier.
-
-```{r crimeload}
-btp_df <- read_csv("data/2020-01-btp-street.csv")
-```
+For that, we can turn to our British Transport Police data.
 

### 2020-05-25 draft_chapter.Rmd
jsonlite edit (now knits)

@@
 library(readr)
-library(dplyr)
 library(tidyr)
 library(patchwork)
+library(dplyr)
 ```
@@
 
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
+api_call <- jsonlite::fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
 ```
@@
 
-tfl_bike_sf <- fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
+tfl_bike_sf <- jsonlite::fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
 

### 2020-05-29 draft_chapter.Rmd
read thru

--- 2020-05-24 draft_chapter.Rmd
@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, eval = F}
 knitr::opts_chunk$set(echo = T, message = F, warning = F)
-library(jsonlite)
-library(osmdata)
-library(ggplot2)
-library(sf)
-library(readr)
-library(dplyr)
-library(tidyr)
-library(patchwork)
+# library(jsonlite)
+# library(osmdata)
+# library(ggplot2)
+# library(sf)
+# library(readr)
+# library(tidyr)
+# library(patchwork)
+# library(dplyr)
 ```
@@
 Even the public dissemination of findings which emerge from analysis might be subject to restrictions.
-This can lead to unequal access, controlled usage and curb the diffusion of findings, severely limiting the insight that can be attained from data.
+This can lead to unequal access, controlled usage and curb the diffusion of findings, severely limiting the insight that can be obtained from data.
 
@@
 
->In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
+In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
 
@@
 The _Global Open Data Index_ [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by country.
-It provides a rank which indicates how well governments across the world follow open data practices across varoius domains.
+It provides a rank which indicates how well governments across the world follow open data practices across various domains.
 Many of these have direct relevance to crime and place research, such as administrative boundaries and geographic locations (e.g.
@@
 
-The Public Sector refers to organizations owned and operated by local or federal government with the core aim to provide services to the public.
+The 'public sector' refers to organizations owned and operated by local or federal government with the core aim to provide services to the public.
 Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data].
-These can include national statistics, or administrative data.
-Another set of data included in this group is data collected by publicly funded research, for example victimisation surveys, housing surveys, and others.
-
-The _Global Open Data Index_ introduced earlier focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
+These can include national statistics or administrative data, but also data collected by publicly funded research projects such as victimization and household surveys, amongst others.
+The _Global Open Data Index_, introduced earlier, focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
 
@@
 -->
-data describing urban structure, such as street networks to assess whether the configuration of streets dictates things like violent crime victimisation [@streetsyntax2017], or victimisation surveys mentioned earlier to quantify citizens' perceived safety and security.
+data describing urban structure, such as street networks to assess whether the configuration of roads dictates things like violent crime victimisation [@streetsyntax2017], or victimisation surveys to quantify citizens' perceived safety and security.
 
@@
 
-Contrary to public sector, private sector refers to the rest of the economy whihc are not under direct state control.
+In contrast to the public sector, 'private sector' tends to refer to parts of the economy which are not under direct state control, and often operate for profit.
 Opening up data generated by the private sector represents a significant challenge, largely due to the proprietary value to its creators [@kitchin2014data].
@@
 Indeed, many papers have made use of data collected or distributed by private organizations to explore crime and place, such as Google Street View images [@langton2017residential] and Twitter [@malleson2015impact].
-In fact, you will learn to use the free Twitter API for such purposes in this book (see Chapter 6, Topic 7).
+In fact, you will learn to how to obtain free Twitter data for such purposes in this book (see Chapter 6, Topic 7).
 
@@
 Often, the organizations who maintain and monitor these data collection activities are charities or non-governmental organizations that do not operate for profit, but instead provide some form of social good.
-An example is the online reporting platform FixMyStreet (https://www.fixmystreet.com/), where people can report problems such as instances of graffiti, vandalism or environment issues.
-The data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
+An example is the online reporting platform FixMyStreet (https://www.fixmystreet.com/), where people can report problems such as instances of graffiti, vandalism or environmental issues.
+This data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
 Another topic in this book (see Chapter 6, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analysing open data of this type.
@@
 
-With this in mind, we now move on to a practical exercise, where we will utilize mutliple sources of open data, and explore police recorded crime on and around public transport in London, England.
+With this in mind, we now move on to a practical exercise in which we will utilize multiple sources of open data to explore police recorded crime on and around public transport in London, England.
 
@@
 
-- Accessing open data using 3 different methods: 
-    + direct downloads 
-    + direct calls to an Application Programming Interface (API)
-    + using an API wrapper.
+- Accessing open data using three different methods: 
+    + Direct download.
+    + Direct calls to an Application Programming Interface (API).
+    + Calls to an API using a wrapper.
 - Cleaning, wrangling, and visualizing open spatial data.
 - Comparing different sources of open data.
-- How to engage with the critical 'Where', 'Why', 'Who', 'What', 'When' questions to better understand open data.
+- How to engage with the critical 'Where', 'Why', 'Who', 'What' and 'When' questions to better understand open data.
 
@@
 In this exercise we will explore crime in and around London Underground stations.
-We know that there are important environmental features when it comes to public transport areas being more or less criminogenic environments.
+We know that environmental features are important when it comes to public transport areas being more or less criminogenic.
 Studies in Sweden, for instance, have shown the importance of environmental and neighborhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
-Similar work has been done exploring bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
+Similar work has been carried out exploring bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
 
-Now we will consider the case of London. Specifically, we will examine to what extent does crime cluster in and around London Underground stations?
-
-We will use various sources of open data to answer this question, and this will allow us to explore the strengths and limitations across public sector and crowdsourced open data sources.
+Here, we will consider the case of London.
+Specifically, we will examine the question: to what extent does crime cluster in and around London Underground stations?
+We will use various sources of open data to answer this question.
+This will allow us to explore the strengths and limitations of public sector and crowdsourced open data sources.
 
@@
 We will be using two different types of open data in this exercise: public sector data (police recorded crime data and local transport authority data) and crowdsourced data (from Open Street Map).
-To access them we will use three different methods, namely, (1) direct download, (2) direct request to an API, 3) request to an API using a wrapper.
+To access them, we will use three different methods, namely, (1) direct download, (2) direct request to an API, 3) request to an API using a wrapper.
 This will give you a few different ideas about how you might go about accessing other open data sets relevant to your research.
@@
 In the United Kingdom, police recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence.
-
-Visit the website above, and you will see a welcome message, and 6 tabs across the top of it, which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
+Visit this website, and you will see a welcome message, and six tabs across the top of it, which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
 
 Before we download any data, we can learn more about it by clicking on the "About" tab.
-This brings up a lot of information that it is important to carefully read through to be able to answer the where, why, who, what, when questions posed earlier.
-Take a moment to read through, and try to take notes on what you think might be relevant for your analysis.
-For example, if we want to map crimes, we might want to explore if there is any type of "anonymisation" that might take place before the data are released (to protect the privcy of the victims for example).
-If you read the about page here, you might find the following note:
+This brings up a lot of information that it is important to review carefully in order to answer the where, why, who, what and when questions posed earlier.
+Take a moment to read through this information, and try to take notes on what you think might be relevant for your analysis.
+For example, if we want to map crimes, we might want to explore if there is any type of "anonymization" that might take place before the data are released (to protect the privacy of the victims).
+If you read the "About" page, you might find the following note:
 
-> Location Anonymisation
+> Location anonymization
 The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.
 
-This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate, and this may have implications for our findings later!
+This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate/ This may have implications for our findings later!
 
 To then actually download some data, move on to clicking on the "Data" tab.
-This should open for you a page titled "Data downloads" where you can see another 5 tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".
-Staying on this "Custom download" page you can make some selections on what sort of data you wanted to download.
-
+This should open a page entitled "Data downloads" under which you can see another five tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".
+By staying on this "Custom download" page you can select what sort of data you want to download.
 We can select the time period and police force of interest, and the type of information required (e.g.
 crimes, stop and search, outcomes).
-We are also informed that the data are downloaded in comma-separated values format (.csv file extension), which is in line with machine-readable, easy to manipulate data format requirements.
+We are also informed that the data are downloaded in comma-separated values format (.csv file extension), which meets our machine-readable, easy-to-manipulate data format requirements.
 
@@
 
-Save this file locally (in your working directory) to a subfolder named "data", and then you can load it in with the `read_csv()` command from the `readr` package (remember to load it with `library(readr)`).
+Save this file locally (in your working directory) to a subfolder named "data".
+You can load it in with the `read_csv()` command from the `readr` package.
+Remember that you will need to load it with `library(readr)`.
 
@@
 
-Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our github page:
+Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our github page.
+Note that the line break below is purely for easy formatting - this will need removing in your own script so that the URL runs across one line.
 
 ```{r crimeload_gh, eval = FALSE}
-btp_df <- read_csv("https://raw.githubusercontent.com/langtonhugh/osm_crim/master/data/2019-01-greater-manchester-street.csv")
+btp_df <- read_csv("https://raw.githubusercontent.com/langtonhugh/osm_crim/master/data/
+                   2019-01-greater-manchester-street.csv")
 ```
 
-You should have now a dataframe with `r nrow(btp_df)` crimes in them ready for you to explore!
+You should now have a data frame with `r nrow(btp_df)` crimes in, ready for you to explore!
 
@@
 
-Another way of downloading open data is through an application programming interface (API).
-This is a tool which defines an interface for a program to interact with a software component, for example, what sort of requests or calls can be made to it, and how these can be made.
+Another way of downloading open data is through an Application Programming Interface (API).
+This is a tool which defines an interface for a programme to interact with a software component.
+For example, it defines the sort of requests or calls which can be made, and how these calls and requests can be carried out.
 Here, we are using the term 'API' to denote tools created by an open data provider to give access to different subsets of their content.
@@
 
-Here we will make a direct request to the API created by London's local government organization responsible for transport: Transport for London (TfL).
-TfL oversee the London Underground, and provide access to their open data through their unified API.
-Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What', 'When' questions, we must find a similar document here.
+Here, we will make a direct request to the API created by London's local government organization responsible for transport: Transport for London (TfL).
+TfL oversee the London Underground.
+They provide access to their open data through a unified API.
+Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What', 'When' questions, we must find a similar document for TfL.
 Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.
@@
 
-To draw conclusions about crimes at London Underground stations, we will need some data specifically about these underground stations.
+To draw conclusions about crimes in and around London Underground stations, we will need some spatial data about the stations themselves.
 
-In the documentation there 'API area' (which offer some guidance on the types of data available).
-Scroll down and find the example calle for stops _Stops_.
-In the description you will  see that this call returns information on stops for busses or London Underground lines, and there are two examples (bus route 24 and the Bakerloo line).
-Take their demonstration url you see for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser.
-When you visit this page you should see something like Figure 1:
+In the documentation, there is a section called 'API area' which offer some guidance on the types of data available.
+Scroll down and find the example called _Stops_.
+You will see that this call returns information on stops for buses or London Underground lines, and there are two examples (bus route 24 and the Bakerloo line).
+Copy their demonstration URL for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser.
+When you visit this page you should see something like Figure 1.
 
@@
 Although this contains the information we requested, it is not very legible for our human eyes.
-It in a format called _JSON_.
+It is in a format called _JSON_.
 This stands for JavaScript Object Notation.
 JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
-As you can see, it is not intuitive to read in its current format but it is very machine readable friendly, once again in line with this requirement for open data.
+As you can see, it is not intuitive to read in its current format _but_ it is machine-readable friendly, which again aligns well with our requirements for open data.
 
-Now you won't usually make these calls by pasting them into a web browser, but we wanted for you to see how it works.
-In readlity, we will actually make these calls from within the R environment, but it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
-For example, we now know that to request the stops from the "Bakerloo" line, we need the url:
+In reality, you won't usually make these calls by pasting them into a web browser, but it is useful to see how it works.
+More likely, we will actually make these calls from within the R environment.
+That said, it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
+For example, we now know that to request the stops from the "Bakerloo" line, we need the URL:
 
@@
 
-Let's say instead we want data for the Northern Line. What do you think that URL will look like ?
+Let's say that instead, we want data for the Northern Line. What do you think that URL will look like? Indeed, all you need to do is replace "bakerloo" with "northern".
 
-Indeed, all you need to do is replace "bakerloo" with "northern", and you got the stops for this route instead. Now let's see how we would make this call in R.
-
-In R, we can use the `fromJSON()` function from the `rjson` package to parse all this information into a data frame (with rows and columns).
+In R, we can use the `fromJSON()` function from the `jsonlite` package to parse all this information into a data frame (with rows and columns).
 It will then become more familiar and usable.
-All we need to do is input the URL from the TfL API into the function, with a bit of help from `readLines()`, another function in base R which is used to read text from URLs.
+All we need to do is input the URL from the TfL API into the function with a bit of help from `readLines()`, a function in base R for reading text from URLs.
 
@@
 ```{r getapi}
-library(rjson)
+library(jsonlite)
 
@@
 JSON is slightly different to traditional data frames with rows and columns, which we are probably more familiar with, because the data is nested.
-For instance, `api_call` is classed as a data frame, but just try  viewing the object using `View(api_call)`.
+For instance, `api_call` is classed as a data frame, but now try viewing the object using `View(api_call)`.
+You will notice that some of the columns are actually lists, rather than character or factor vectors, which what we might usually expect.
+Another way of exploring the `api_call` object is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
 
-You will notice that you see a list of lists, which is what we call nested lists.
-Another way of checking what sort of object this `api_call` object is is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
-
-If you run this, you will see that each element in `api_call` is a "list".
-
-This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated, not always a neatly formatted dataframe like the csv from the police.uk open data portal.
-But we will learn how to manage this data into a more familiar format later.
+This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated.
+It is not always a neatly formatted data frame like the csv from the open police data portal.
+That said, we can transform this data into something more familiar from within R, as we will see later on.
 
@@
 
-Finally, the last way we will look at how open data is often shared is by using API wrappers.
-Often, developers who work with APIs will share their code, and release them in the form of a package or module, so that other people can use it too.
+Finally, we will look at how open data can be obtained using API wrappers.
+Often, developers who work with APIs will share their code, and release them in the form of a package or module, so that other people can use it.
 This is called a _wrapper_ because it uses code that 'wraps' around the API to make it a neater, more usable package.
@@
 
-If we load and use the package `osmdata` we can use its funcitons to query the Open Street Map API, rather than the API query being made directly (like we did for TfL, using the URL).
-Once again, we want to have some documentation to be able to refer to about our data and the ways to query it.
+If we load the package `osmdata` we can use its functions to query the Open Street Map API, rather than the API query being made directly (like we did for TfL, using the URL).
+Once again, we will want to identify documentation which can help us understand and critique our data, and learn how to query it.
 When it comes to OSM, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about).
-Regarding the wrapper, as it is an R package, it will come with a handy vignette, which in this case we can find at [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).
+Using the wrapper in R, we can refer to the package documentation and the associated vignette which we can find at [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).
 
 Unlike the TfL API, `osmdata` is an international database, and has lots of data that we might not necessarily need.
-The first thing we need to specify therefore, is where we want to get data from.
+As such, the first thing we need to specify is our study region.
 To do this, we use the `getbb()` function from the `osmdata` package, which stands for "get bounding box".
-You can think of the bounding box as a box drawn around an area that we are interested in (in this case, London, UK) which tells the OSM API that we want everything *inside* the box, but nothing *outside* the box.
+You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the OSM API that we want everything *inside* the box, but nothing *outside* the box.
 
-So how can we name a bounding box specification to define the study region?
+So, how can we name a bounding box specification to define the study region?
 This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, or you can use a search term.
-Here, we want to select Greater London, in the UK so we can use the search term "greater london united kingdom".
-Besides specifying where we want data from, we can also tell the `getbb()` function what format we want the data in.
-Let's say in this case we want a spatial object, specifically an `sf` polygon object which we name `bb_sf`:
+Here, we want to select Greater London, so we can use the search term "greater london united kingdom".
+Besides specifying the study region, we can also tell the `getbb()` function what format we want the data to be in.
+In this case we want a spatial object, specifically an `sf` polygon object, which we name `bb_sf`:
 
@@
 
-We now have our study region defined as the administrative boundaries of Greater London, which we can visualize quickly to check it has worked as expected.
-To visualise we can use the `sf` and `ggplot2` packages.
+We now have our study region defined as the administrative boundaries of Greater London. We can visualize this quickly to check that it has worked as expected using the `sf` and `ggplot2` packages.
 
@@
 
-If you are familiar with the shape of London then this image should be familiar as well!
-Pleased with our bounding box, we can now move on to scrape data from the Open Street Map API using the `opq()` function.
+If you are familiar with the shape of London then this image should be quite recognizable!
+We can now move on to scrape data from the Open Street Map API using the `opq()` function.
 The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
@@
 
-We do this using the `add_osm_feature()` function, specifying our key as 'public transport' and our value as 'station'.
+We can select what features we want using the `add_osm_feature()` function, specifying our key as 'public transport' and our value as 'station'.
+We also want to specify what sort of object (what class) to get our data into, and as we're still working with spatial data, we stick to the `sf` format, for which the function is `osmdata_sf()`.
 
-We also want to specify what sort of object (what class) to get our data into, and as we're still working with spatial data we stick to the `sf` format, for which the function is `osmdata_sf()`.
-
-And finally, we trim everything down into our bounding box parameters with the `trim_osmdata()` function. We add all these together using a pipe (`%>%`) and make the call like so:
-
+Finally, we trim everything down to our study region with the `trim_osmdata()` function.
+Without this addition, we would get points for everywhere within our bounding box, including some areas outside of Greater London.
+We add all these together using a pipe (`%>%`) and make the call like so:
 ```{r opq}
@@
 We also have around one hundred polygons.
-
 For now, let's extract the point information.
@@
 
-We now have a simple features object with all the public transport stations in London, which the volunteers of OSM have mapped, and also 130 variables of auxiliary data they collected, such as the "fare_zone" the station is in, what "amenity" it may have, whether it has "toilets", and many many more.
+We now have an `sf` object with all the public transport stations in London mapped by OSM volunteers, along with 130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others.
+Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not have added information.
+Nevertheless, when the details are recorded, they provide rich insight and local knowledge we may otherwise not be able to obtain.
 
-Of course, this is up to the volunteers whether they collect all these data, and in many cases they just might not. For example, if we look at fare zone, we might see the following:
-
-```{r, echo=FALSE}
-library(dplyr)
-
-osm_stat_sf %>% group_by(fare_zone) %>% count() %>% st_drop_geometry() %>% arrange(-n)
-
-```
-
-You can see for most of these stations (`r osm_stat_sf %>% group_by(fare_zone) %>% count() %>% st_drop_geometry() %>% filter(is.na(fare_zone)) %>% pull(n)`) we have missing data (denoted by NA).
-Nevertheless, when the details are recorded, they provide rich insight and local knowledge we may otherwise not attain.
-
-One last thing to do is to select the stations which are relevant to us (ie: they fall along the Northern Line).
+One additional step needed is to select the stations which are relevant to us (i.e.
+they fall along the Northern Line).
 The variable `line` can help us with this.
-First let's look at the values in this variable by using the `unique()` function:
+First, let's look at the values in this variable by using the `unique()` function:
 
@@
 
-You might see that Northern appears a few times, once as "Northern" but also once as "Northern City".
-As we know there is no line called "Northern City" and we can see many other combinations in these values (eg Circle;District, indicating the station serves both Circle and District lines) we can assume that this is a station that serves both Northern line and one other.
-
-To select all Northern line stations we can use the `grepl()` function to collect everything with "Northern" in the name:
+You might see that the Northern line appears a few times, once as "Northern" but also once as "Northern City".
+We know that there is no line called "Northern City", and we can see that many other combinations of these values exist (e.g.
+"Circle;District" indicates that the station serves both Circle and District lines).
+Therefore, we can assume that stations like "Northern City" serve both the Northern line and one other.
+To get round this, and select all stations with "Northern" somewhere in the name, we can use the `grepl()` function from base R.
+Note that we also make use of `filter()` from within `dplyr`, so we first need to load that package.
 
 ```{r filterline}
+library(dplyr)
+
 osm_north_sf <- osm_stat_sf %>% 
@@
 
-And now we have all our open data sets together in one place!
+Now we have all our open data sets loaded into the R environment!
 
@@
 
-Just because we have our data doesn't necesarily mean it's in the shape we want yet.
-The first thing to do once we have acquired our open data is investigate it, using data cleaning, wrangling, and visualisation.
-
-We've already loaded the `dplyr` and `sf` and `ggplot2` packages, but here we will also make use of `tidyr` and `patchwork`, so load these (remember, if you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.)
+Obtaining open data from a direct download or API is often only half the battle.
+It doesn't necessarily mean that the data is in the shape needed to conduct analysis.
+The first thing to do once we have acquired open data is investigate it using data cleaning, wrangling, and visualisation.
+We've already loaded the `dplyr`, `sf` and `ggplot2` packages, but here, we will also make use of `tidyr` and `patchwork`, so load these now.
+Remember, if you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
 
@@
 
-We usually want to work with data that is in the dataframe format, that is all your observations are rows and all your variables are columns.
+We usually want to work with data that is in the format of a data frame, that is, all your observations are rows and all your variables are columns.
 While this is the case for our police data and our OSM data, recall that the TfL data came in the nested JSON format.
@@
 
-We can then plot these OSM station points over the station locations pulled from the TfL API, to see how the two data sets compare.
+We can now plot the OSM station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare.
 We do not plot the study region boundary in order to get a more detailed, local view.
@@
 That said, without local knowledge, an alternative open data source, or a critical approach to using open data, well-intentioned researchers might draw erroneous conclusions from filtering stations tagged as on the Northern line.
-This gives us a preliminary answer to our second research question already: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.
+This gives us a preliminary answer to our earlier questions: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.
 
@@
 For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
-It is worth noting that this definition is somewhat arbitrary, and is subject to a degree of spatial inaccuracy is open police recorded crime data [see tompson2015uk], but it does serve to facilitate this demonstration.
+It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see tompson2015uk], but it does serve to facilitate this demonstration.
 
@@
 ```{r crimemapcompare}
+library(patchwork)
+
 p1 <- ggplot(data = osm_north_sf) +
@@
 The number and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology.
-For instance, information which traces building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colourin], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
+For instance, information which traces building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
 But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go.

### 2020-05-29 draft_chapter.Rmd
Final edits

@@
 
-```{r setup, include=FALSE, eval = F}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = T, message = F, warning = F)
@@
 
-The 'public sector' refers to organizations owned and operated by local or federal government with the core aim to provide services to the public.
+The 'public sector' refers to organizations owned and operated by local or central government with the core aim to provide services to the public.
 Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data].
@@
 
-So what can we do to make sure we make good use of these data?
+So, what can we do to make sure we make good use of these data?
 With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science.
@@
 In the United Kingdom, police recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence.
-Visit this website, and you will see a welcome message, and six tabs across the top of it, which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
+Visiting this website, you will see a welcome message, and six tabs across the top which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
 
 Before we download any data, we can learn more about it by clicking on the "About" tab.
-This brings up a lot of information that it is important to review carefully in order to answer the where, why, who, what and when questions posed earlier.
-Take a moment to read through this information, and try to take notes on what you think might be relevant for your analysis.
-For example, if we want to map crimes, we might want to explore if there is any type of "anonymization" that might take place before the data are released (to protect the privacy of the victims).
+This brings up information that is important to review carefully in order to answer the where, why, who, what and when questions posed earlier.
+Take a moment to read through this information, and take notes on what you think might be relevant for your analysis.
+For example, if we want to map crimes, we will want to explore if there is any type of "anonymization" that might take place before the data are released (to protect the privacy of the victims).
 If you read the "About" page, you might find the following note:
 
-> Location anonymization
+> Location anonymization.
 The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.
 
-This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate/ This may have implications for our findings later!
+This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate. This may have implications for our findings!
 
-To then actually download some data, move on to clicking on the "Data" tab.
+To then download some data, move on to clicking on the "Data" tab.
 This should open a page entitled "Data downloads" under which you can see another five tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".
@@
 You can load it in with the `read_csv()` command from the `readr` package.
-Remember that you will need to load it with `library(readr)`.
+Remember that you will need to load it first using `library()`.
 
@@
 library(readr)
+
 btp_df <- read_csv("data/2020-01-btp-street.csv")
@@
 
-Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our github page.
+Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our GitHub page.
 Note that the line break below is purely for easy formatting - this will need removing in your own script so that the URL runs across one line.
@@
 They provide access to their open data through a unified API.
-Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What', 'When' questions, we must find a similar document for TfL.
+Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What' and 'When' questions, we must find a similar document for TfL.
 Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.
@@
 If you follow this link you should see a page titled "Our Unified API" and some examples of calls.
-
 To draw conclusions about crimes in and around London Underground stations, we will need some spatial data about the stations themselves.
 
-In the documentation, there is a section called 'API area' which offer some guidance on the types of data available.
+In the documentation, there is a section called 'API area' which offers some guidance on the types of data available.
 Scroll down and find the example called _Stops_.
@@
 Copy their demonstration URL for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser.
-When you visit this page you should see something like Figure 1.
+When you visit this page, you should see something like Figure 1.
 
@@
 For instance, `api_call` is classed as a data frame, but now try viewing the object using `View(api_call)`.
-You will notice that some of the columns are actually lists, rather than character or factor vectors, which what we might usually expect.
+You will notice that some of the columns are actually lists, rather than character or factor vectors, which is what we might usually expect.
 Another way of exploring the `api_call` object is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
@@
 This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated.
-It is not always a neatly formatted data frame like the csv from the open police data portal.
+It is not always a neatly formatted data frame like the CSV file from the open police data portal.
 That said, we can transform this data into something more familiar from within R, as we will see later on.
@@
 Once again, we will want to identify documentation which can help us understand and critique our data, and learn how to query it.
-When it comes to OSM, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about).
-Using the wrapper in R, we can refer to the package documentation and the associated vignette which we can find at [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).
+When it comes to Open Street Map, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about).
+Using the wrapper in R, we can refer to the package documentation and the associated vignette online: [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).
 
@@
 To do this, we use the `getbb()` function from the `osmdata` package, which stands for "get bounding box".
-You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the OSM API that we want everything *inside* the box, but nothing *outside* the box.
+You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the Open Street Map API that we want everything *inside* the box, but nothing *outside* the box.
 
@@
 Besides specifying the study region, we can also tell the `getbb()` function what format we want the data to be in.
-In this case we want a spatial object, specifically an `sf` polygon object, which we name `bb_sf`:
+In this case we want a spatial object, specifically an `sf` polygon object, which we name `bb_sf`.
 
@@
 
-We now have an `sf` object with all the public transport stations in London mapped by OSM volunteers, along with 130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others.
-Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not have added information.
-Nevertheless, when the details are recorded, they provide rich insight and local knowledge we may otherwise not be able to obtain.
+We now have an `sf` object with all the public transport stations in London mapped by Open Street Map volunteers, along with 130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others.
+Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not added information.
+Nevertheless, when the details are recorded, they provide rich insight and local knowledge that we may otherwise be unable to obtain.
 
@@
 We usually want to work with data that is in the format of a data frame, that is, all your observations are rows and all your variables are columns.
-While this is the case for our police data and our OSM data, recall that the TfL data came in the nested JSON format.
+While this is the case for our police data and our Open Street Map data, recall that the TfL data came in the nested JSON format.
 
@@
 
-We can now plot the OSM station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare.
+We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare.
 We do not plot the study region boundary in order to get a more detailed, local view.
@@
 This could be due to a number of reasons, among them: variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting].
-Given that South London is historically less wealthy in comparison to the North, for instance, this would certainly be an interesting avenue to pursue.
 Interestingly, these missing stations _have_ actually been recorded by contributors in Open Street Map, but _without_ information about the line on which the station is located.
@@
 That said, without local knowledge, an alternative open data source, or a critical approach to using open data, well-intentioned researchers might draw erroneous conclusions from filtering stations tagged as on the Northern line.
-This gives us a preliminary answer to our earlier questions: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.
+This gives us a preliminary answer to our earlier question: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.
 
-But, what might the impact of this be on studying crime in and around London Underground stations, given the differences observed between the two open data sources?
+But, what might the impact of this be when studying crime in and around London Underground stations, given the differences observed between the two open data sources?
 For that, we can turn to our British Transport Police data.
@@
 
-We now have the location of all crimes recorded by the British Transport Police in Greater London during January, 2020.
+We now have the recorded location of all crimes recorded by the British Transport Police in Greater London during January, 2020.
 For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
-It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see tompson2015uk], but it does serve to facilitate this demonstration.
+It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see @tompson2015uk], but it does serve to facilitate this demonstration.
 
@@
 ```{r crimemapcompare}
-library(patchwork)
-
 p1 <- ggplot(data = osm_north_sf) +
@@
 The conclusions drawn are starkly different, and would determine the consistency of findings with existing research examining crime at underground stations [@ceccato2013security].
+
+## Alternative example
 
@@
 Refer back to the example earlier if some of the code is unclear.
-Note that this is a major benefit of using an API wrapper from within R, because the code used to query each open data source can easily be re-used, with minor amendments, to extract different features.
+Note that this is a major benefit of using an API wrapper from within R, because the code used to query each feature can easily be re-used, with minor amendments, to extract different information.
 

### 2020-05-30 draft_chapter.Rmd
refs update done

@@
 First, it is useful to clarify what we mean when we refer to 'open data'.
-The _Open Data Handbook_, compiled by the Open Knowledge Foundation, states that open data are "data that can be freely used, re-used and redistributed by anyone - subject only, at most, to the requirement to attribute and sharealike" [@dietrich2009open].
+The _Open Data Handbook_, compiled by the Open Knowledge Foundation, states that open data are "data that can be freely used, re-used and redistributed by anyone - subject only, at most, to the requirement to attribute and sharealike" [@dietrich2009open, para.
+3].
 Specifically, open data can be defined along three key domains [@knowledge2016open].

### 2020-06-11 draft_chapter.Rmd
anon & Word version

@@
 title: "Open Data"
-author: "Samuel Langton (MMU) & Reka Solymosi (University of Manchester)"
 date: "June 2020"
+editor_options:
+  chunk_output_type: console
 output:
+  word_document: default
   pdf_document:
     latex_engine: xelatex
-editor_options: 
-  chunk_output_type: console
 bibliography: refs.bib
@@
 ```{r crimeload_gh, eval = FALSE}
-btp_df <- read_csv("https://raw.githubusercontent.com/langtonhugh/osm_crim/master/data/
+btp_df <- read_csv("https://raw.githubusercontent.com/anonymous/osm_crim/master/data/
                    2019-01-greater-manchester-street.csv")

### 2020-07-14 draft_chapter.Rmd
updated to nrow for points retrieved

@@
 If you are familiar with the shape of London then this image should be quite recognizable!
-We can now move on to scrape data from the Open Street Map API using the `opq()` function.
+We can now move on to query data from the Open Street Map API using the `opq()` function.
 The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
@@
 As one might expect, most information relating to public transport station locations has been recorded using points (i.e.
-two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing.
+two-dimensional vertices, coordinates) of which we have `r nrow(osm_stat_sf$osm_points)` at the time of writing.
 We also have around one hundred polygons.

### 2020-09-25 draft_chapter.Rmd
re-knit pdf and word

@@
 output:
-  word_document: default
   pdf_document:
     latex_engine: xelatex
+  word_document: default
 bibliography: refs.bib

### 2020-09-28 scripts/osm_explore_tube.R
jubilee line + bbox using lat long

@@ -7,6 +7,7 @@ library(dplyr)
 library(tidyr)
 library(purrr)
 library(patchwork)
+library(stringr)
 
 # OSM query ==============================================================================================
 
@@ -15,12 +16,13 @@ library(patchwork)
 # an area using the latitude and longitude coordinates, but it is generally
 # easier to use a search term. For instance
 
-bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon") 
+# no longer works.
+# bb_sf <- getbb(place_name = "Greater London United Kingdom", format_out = "sf_polygon")
 
-osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
-  add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
-  osmdata_sf() %>%                                                 # specify class (sf or sp)
-  trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box  
+osm_stat_sf <- opq(bbox = c(-0.5103751, 51.2867602, 0.3340155, 51.6918741)) %>% # select bounding box
+  add_osm_feature(key = 'public_transport', value = 'station') %>%           # select features
+  osmdata_sf()                                                               # specify class (sf or sp)
+  # trim_osmdata(bb_poly = bb_sf)                                           # trim by bounding box  
 
 # View contents.
 osm_stat_sf
@@ -29,37 +31,42 @@ osm_stat_sf
 temp <- osm_stat_sf$osm_polygons
 osm_stat_sf <- osm_stat_sf$osm_points
 
-  
-  
 # Check the network tags.
-table(osm_stat_sf$network)
+table(osm_stat_sf$line)
 
-# Filter Northern line only/
-osm_north_sf <- osm_stat_sf %>% 
-  filter(line == "Northern")
+# Filter jubern line only/
+osm_jub_sf <- osm_stat_sf %>% 
+  mutate(tube_line = line) %>% 
+  filter(str_detect(tube_line, "Jubilee"))
 
-osm_north_sf <- st_transform(osm_north_sf, 27700)
[…]

### 2020-09-30 draft_chapter.Rmd
Jubilee change rmd

@@
 ```{r crimeload_gh, eval = FALSE}
-btp_df <- read_csv("https://raw.githubusercontent.com/anonymous/osm_crim/master/data/
-                   2019-01-greater-manchester-street.csv")
+btp_df <- read_csv("https://raw.githubusercontent.com/langtonhugh/osm_crim/
+                   master/data/2020-01-metropolitan-street.csv")
 ```
@@
 
-To keep things focused, let's request data about stop points on the Northern line, one of the largest and busiest on the network.
+To keep things focused, let's request data about stop points on the Jubilee line, one of the busiest on the network.
 Note how this is simply an amended version of the example provided by TfL in the API documentation.
@@
 
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/northern/stoppoints"))
+api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/jubilee/stoppoints"))
 ```
@@
 As such, the first thing we need to specify is our study region.
-To do this, we use the `getbb()` function from the `osmdata` package, which stands for "get bounding box".
+To do this, need to specify a bounding box.
 You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the Open Street Map API that we want everything *inside* the box, but nothing *outside* the box.
@@
 So, how can we name a bounding box specification to define the study region?
-This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, or you can use a search term.
+One way to do this is through a search term.
 Here, we want to select Greater London, so we can use the search term "greater london united kingdom".
 Besides specifying the study region, we can also tell the `getbb()` function what format we want the data to be in.
-In this case we want a spatial object, specifically an `sf` polygon object, which we name `bb_sf`.
+In this case we want a spatial object, specifically an `sf` polygon object (from the `sf` package) which we name `bb_sf`.
 
-```{r bb}
+```{r bb, message=F, warning=F}
 library(osmdata)
+library(sf)
 
@@
 
-We now have our study region defined as the administrative boundaries of Greater London. We can visualize this quickly to check that it has worked as expected using the `sf` and `ggplot2` packages.
+One downside of using search terms to define the bounding box, as with any search term, is that inconsistencies in terminology, wordings and spelling might lead to unexpected outputs (or none at all).
+Another way to obtain the bounding box is to manually specify the latitude and longitude coordinates.
+This requires a bit of pre-existing knowledge or research about your study region.
+A quick internet search, or exploration of the interactive OSM online map (https://www.openstreetmap.org/), will give you an idea of the coordinates for your study region.
+For Greater London, we can specify our bounding box as follows, saving the coordinates to the object `bb_gl` for later use.
 
-```{r bbmap}
-library(ggplot2)
-library(sf)
-
-ggplot(data = bb_sf) +
-  geom_sf()
+```{r bbmanual}
+bb_gl <- c(-0.51037, 51.28676, 0.33401, 51.69187) # xmin, ymin, xmax, ymax
 ```
 
-If you are familiar with the shape of London then this image should be quite recognizable!
 We can now move on to query data from the Open Street Map API using the `opq()` function.
@@
 
-Besides specifying what area we want to query with our bounding box object (`bb_sf`) in the `opq()` function, we must also define the feature which we want to pull from the API.
+Besides specifying what area we want to query with our bounding box object(s) in the `opq()` function, we must also define the feature which we want to pull from the API.
 Features in Open Street Map are defined through 'keys' and 'values'.
@@
 We also want to specify what sort of object (what class) to get our data into, and as we're still working with spatial data, we stick to the `sf` format, for which the function is `osmdata_sf()`.
+Here, we specify our bounding box as the manual coordinates contained in `bb_gl`.
+^[If you use the bounding box obtained through `getbb()` one can subsequently trim down the outputs from `add_osm_feature()` using `trim_osmdata()`.
+For instance, we could add `trim_osmdata(bb_poly = bb_sf)` to our initial query.]
 
-Finally, we trim everything down to our study region with the `trim_osmdata()` function.
-Without this addition, we would get points for everywhere within our bounding box, including some areas outside of Greater London.
-We add all these together using a pipe (`%>%`) and make the call like so:
 ```{r opq}
-osm_stat_sf <- opq(bbox = bb_sf) %>%                               # select bounding box
+osm_stat_sf <- opq(bbox = bb_gl) %>%                               # select bounding box
   add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
-  osmdata_sf() %>%                                                 # specify class
-  trim_osmdata(bb_poly = bb_sf)                                    # trim by bounding box 
+  osmdata_sf()                                                     # specify class
 ```
@@
 two-dimensional vertices, coordinates) of which we have `r nrow(osm_stat_sf$osm_points)` at the time of writing.
-We also have around one hundred polygons.
+We also have around fifty polygons.
 For now, let's extract the point information.
@@
 
-We now have an `sf` object with all the public transport stations in London mapped by Open Street Map volunteers, along with 130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others.
+We now have an `sf` object with all the public transport stations in our study region mapped by Open Street Map volunteers, along with  ~130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others.
 Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not added information.
@@
 One additional step needed is to select the stations which are relevant to us (i.e.
-they fall along the Northern Line).
+they fall along the Jubilee Line).
 The variable `line` can help us with this.
-First, let's look at the values in this variable by using the `unique()` function:
+First, let's look at the values in this variable by using the `table()` function:
 
-```{r valuesofline}
-
-unique(osm_stat_sf$line)
-
+```{r valuesofline, eval = F}
+table(osm_stat_sf$line)
 ```
 
-You might see that the Northern line appears a few times, once as "Northern" but also once as "Northern City".
-We know that there is no line called "Northern City", and we can see that many other combinations of these values exist (e.g.
-"Circle;District" indicates that the station serves both Circle and District lines).
-Therefore, we can assume that stations like "Northern City" serve both the Northern line and one other.
-To get round this, and select all stations with "Northern" somewhere in the name, we can use the `grepl()` function from base R.
+You will see that "Jubilee" appears several times under different line names.
+The main category ("Jubilee") is the most common, but many other combinations exist (e.g.
+Central;Jubilee) for stations which sit on multiple lines.
+To get round this, and select all stations with "Jubilee" somewhere in the name, we can use the `grepl()` function from base R.
 Note that we also make use of `filter()` from within `dplyr`, so we first need to load that package.
@@
 
-osm_north_sf <- osm_stat_sf %>% 
-  filter(grepl("Northern", line))
+osm_jub_sf <- osm_stat_sf %>% 
+  filter(grepl("Jubilee", line))
 ```
@@
 The first thing to do once we have acquired open data is investigate it using data cleaning, wrangling, and visualisation.
-We've already loaded the `dplyr`, `sf` and `ggplot2` packages, but here, we will also make use of `tidyr` and `patchwork`, so load these now.
+We've already loaded the `osmdata` and `dplyr` packages, but here, we will also make use of `ggplot2`, `sf`, `tidyr` and `patchwork`, so load these now.
 Remember, if you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
@@
 ```{r packages}
+library(ggplot2)
+library(sf)
 library(tidyr)
@@
 ```{r apispatial}
-tfl_north_sf <- api_call %>% 
+tfl_jub_sf <- api_call %>% 
   select(commonName, lat, lon) %>% 
@@
 ```{r tflmap}
-ggplot(tfl_north_sf) +
+ggplot(tfl_jub_sf) +
   geom_sf()
@@
 
-There we have it: with just a few lines of code in R, we have queried the TfL API, a public sector source of open data, and plotted a basic map of stations on the Northern underground line.
+There we have it: with just a few lines of code in R, we have queried the TfL API, a public sector source of open data, and plotted a basic map of stations on the Jubilee underground line.
 
@@
 We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare.
-We do not plot the study region boundary in order to get a more detailed, local view.
 
@@
 ggplot() +
-  geom_sf(data = tfl_north_sf) +
-  geom_sf(data = osm_north_sf, color = "red", alpha = 0.5) 
+  geom_sf(data = tfl_jub_sf) +
+  geom_sf(data = osm_jub_sf, color = "red", alpha = 0.5) 
 ```
 
-As we can see, there is clearly some discrepancy between the two data sources.
-There are slight differences in the exact point locations of stations, but more saliently,  Open Street Map appears to be missing stations, especially in Central and South London.
-This could be due to a number of reasons, among them: variation in residents' willingness to contribute to information about their local area, and the demographic characteristics known to influence the volume of contributions [@mashhadi2013putting].
-Interestingly, these missing stations _have_ actually been recorded by contributors in Open Street Map, but _without_ information about the line on which the station is located.
-We can see this by plotting out all Open Street Map stations using `osm_stat_sf`.
-That said, without local knowledge, an alternative open data source, or a critical approach to using open data, well-intentioned researchers might draw erroneous conclusions from filtering stations tagged as on the Northern line.
-This gives us a preliminary answer to our earlier question: the public sector open data from TfL appears to hold some advantage over its crowdsourced alternative, Open Street Map, in terms of London Underground lines.
+As we can see, the two data sources appear to be largely comparable in terms of their spatial patterning.
+However, the exact location of stations differs between OSM and TfL.
+Determining a single location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels.
+For operational purposes, TfL might have a specific definition for the location of a station, such as its centroid, which has been systematically applied to the dataset.
+The data pulled from OSM might have been compiled by a volunteer contributor with their own perception of how to define the location of a station, such as the ticket booth or the main entrance.
+This definition _might_ be more theoretically meaningful than that used by TfL, since it might reflect the public's perception of the station's built environment, and thus opportunities for crime.
+However, it is also plausible that the locations from OSM have been collated by multiple different contributors, each with their own idea of where each station lies on the line.
+This demonstrates why critically engaging with open data sources is so crucial, but also challenging, because it might generate lots of new questions which need to be explored.
 
-But, what might the impact of this be when studying crime in and around London Underground stations, given the differences observed between the two open data sources?
+Given the differences observed between the two open data sources, what might the impact be when studying crime in and around London Underground stations?
 For that, we can turn to our British Transport Police data.
 
-In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions we used earlier.
-Note that we need to add `st_intersection()` to clip our points to our study region, because the British Transport Police are operational across the country, not just in London.
-To do this, we first project `bb_sf` to the British National Grid, so that we perform the intersection on data which have the same projection.
+In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions from the `sf` package that we used earlier.
 
-```{r crimespatial}
-# First, project the study region boundaries.
-bb_sf <- st_transform(bb_sf, 27700)
-
-# Then, make the police data spatial, and clip.
+```{r crimespatial, echo=F}
 btp_sf <- btp_df %>%
@@
   st_as_sf(coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
-  st_transform(27700) %>%
-  st_intersection(bb_sf)
+  st_transform(27700)
 ```
 
-We now have the recorded location of all crimes recorded by the British Transport Police in Greater London during January, 2020.
-For the purposes of this demonstration, we can define crimes occurring as 'in and around' Northern line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
+We now have the recorded location of all crimes recorded by the British Transport Police during January, 2020.
+For the purposes of this demonstration, we can define crimes occurring as 'in and around' Jubilee line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
 It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see @tompson2015uk], but it does serve to facilitate this demonstration.
@@
 ```{r buffer}
-# Assign the same CRS.
-osm_north_sf <- st_transform(osm_north_sf, 27700)
-tfl_north_sf <- st_transform(tfl_north_sf, 27700)
+# Ensure the same CRS (BNG, 27700).
+osm_jub_sf <- st_transform(osm_jub_sf, st_crs(tfl_jub_sf))
 
 # Create buffers to define 'in and around'.
-osm_buff_sf <- st_buffer(osm_north_sf, dist = 50)
-tfl_buff_sf <- st_buffer(tfl_north_sf, dist = 50)
+osm_buff_sf <- st_buffer(osm_jub_sf, dist = 50)
+tfl_buff_sf <- st_buffer(tfl_jub_sf, dist = 50)
 
 # Count number of crimes recorded within each buffer.
-osm_north_sf <- osm_buff_sf %>% 
+osm_jub_sf <- osm_buff_sf %>% 
   mutate(crimes = lengths(st_intersects(osm_buff_sf, btp_sf)))
-tfl_north_sf <- tfl_buff_sf %>% 
+tfl_jub_sf <- tfl_buff_sf %>% 
   mutate(crimes = lengths(st_intersects(tfl_buff_sf, btp_sf)))
@@
 We can then visualize these counts and compare the two sources of data by coloring in our point locations according to the crime count.
-Note that we arrange the plots side-by-side using syntax available using the `patchwork` package.
+Note that we arrange the plots using syntax available using the `patchwork` package, and set the scales to be comparable using the `lim` argument within `scale_color_viridis_c()`.
 
 ```{r crimemapcompare}
-p1 <- ggplot(data = osm_north_sf) +
+p1 <- ggplot(data = osm_jub_sf) +
   geom_sf(mapping = aes(color = crimes), size = 2) +
   labs(title = "Open Street Map") +
-  scale_color_viridis_c()
+  scale_color_viridis_c(lim = c(0,42))
 
-p2 <- ggplot(data = tfl_north_sf) +
+p2 <- ggplot(data = tfl_jub_sf) +
   geom_sf(mapping = aes(color = crimes), size = 2) +
   labs(title = "Transport for London") +
-  scale_color_viridis_c()
+  scale_color_viridis_c(lim = c(0,42))
 
-p1 + p2
+p1 / p2
 ```
@@
 
-```{r bikexample}
+```{r bikexample, eval = F}
 bb_sf <- getbb(place_name = "greater london", format_out = "sf_polygon")

### 2020-09-30 draft_chapter.Rmd
jubilee redraft done

@@
 ---
-title: "Open Data"
-date: "June 2020"
+title: "Open Data for _Crime and Place_ Research: A Practical Guide in R."
+author:
+  - Samuel Langton\dag
+  - Reka Solymosi\ddag
 editor_options:
@@
 bibliography: refs.bib
+csl: apa.csl
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = T, message = F, warning = F)
+```{r setup, include = F, echo = F}
+knitr::opts_chunk$set(message = F, warning = F, eval = T, echo = T)
 # library(jsonlite)
@@
 
+\enspace
+
+This is an unedited draft, not for citation, of a book chapter that has been accepted for publication in E.
+Groff and C.
+Haberman (Eds.), The study of crime and place: A methods handbook (Temple University Press, 2021).
+Temple University Press retains all rights and subrights of this document.
+This pre-print is not published under a CC license.
+
+\enspace
+
+**Keywords:** crime, open data, Open Street Map, transport, geography.
+
+**Corresponding author:** Samuel Langton, Manchester Metropolitan University, Crime and Well-Being
+
+Big Data Centre, Geoffrey Manton Building, Manchester, M16 6EB, United Kingdom.
+
+Telephone: +44 (0)161 247 2000. Email: s.langton@mmu.ac.uk.
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+\enspace
+
+***
+
+\dag Department of Sociology, Manchester Metropolitan University, Manchester, United Kingdom. ORCID: 0000-0002-1322-1553.
+
+\ddag School of Social Sciences, University of Manchester, Manchester, United Kingdom. ORCID: 0000-0001-8689-1526.
+
+\newpage
+
 # Introduction
@@
 Open data can lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
-As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyse open data sets.
+As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyse open datasets.
 
@@
 The chapter also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.
-
-<!-- We achieve this by first giving a background on open data, discussing the key types of data out there, critically assessing the strengths and limitations of each, and building a how-to guide checklist for researchers to follow when approaching an open data source.
-Finally, we work through an example that shows how to access, wrangle, link, and interpret open data from various sources, providing a template which can be applied in other research areas.
--->
 
@@
 Open data practices, and the extent to which it is a reality, vary both within and between countries.
-For example, the United States has a history of making public sector data sets openly available.
+For example, the United States has a history of making public sector datasets openly available.
 The United Kingdom can be more strict, releasing data under a licence ('Open Government Licence'), with some data requiring a fee [@kitchin2014data].
@@
 The drive towards transparency and scrutiny of public information is further enhanced by an increasing need for replication - a hallmark for open science research practice.
-Opening up data sets used in criminology publications and wider social sciences fosters and facilitates a culture of replication [@pridemore2018replication].
+Opening up datasets used in criminology publications and wider social sciences fosters and facilitates a culture of replication [@pridemore2018replication].
 As such, open data fits into the wider discourse around transparency, open review and open scrutiny, which (in time) may even become a policy requirement irrespective of the research field or the source of funding [@vuong2017open].
@@
 Now we have a reasonable idea about what open data is in a broad sense, we can consider the different types of open data which you might come across.
-Here, we formulate a typology of open data sets based on their origin.
+Here, we formulate a typology of open datasets based on their origin.
 
@@
 
-The 'public sector' refers to organizations owned and operated by local or central government with the core aim to provide services to the public.
+The 'public sector' refers to organizations owned and operated by local or central government with the core aim of providing services to the public.
 Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data].
 These can include national statistics or administrative data, but also data collected by publicly funded research projects such as victimization and household surveys, amongst others.
-The _Global Open Data Index_, introduced earlier, focuses on identifying the gaps in governmental organizations, encouraging them think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
+The _Global Open Data Index_, introduced earlier, focuses on identifying the gaps in governmental organizations, encouraging them to think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
 
-Specific to crime and place research, data sets of interest in this particular domain might include
-<!-- As researchers studying crime and place, public sector data is particularly important.
-Many variables thought to explain the non-random distribution of crime, offenders and victims in space are collated and published by governmental bodies.
-For instance, demographic characteristics about resident populations can be collected via national censuses, with data usually released at aggregations which constitute 'neighborhoods' - a theoretically important scale in criminology.
--->
-data describing urban structure, such as street networks to assess whether the configuration of roads dictates things like violent crime victimisation [@streetsyntax2017], or victimisation surveys to quantify citizens' perceived safety and security.
-
-<!-- In many countries, representative victimisation surveys are openly available for download.
-This is not only useful for substantive research (e.g.
-quantifying long-term crime trends), but also scrutiny over reliability and accuracy in other forms of open data, such as police recorded crime records.
-Both of these examples represent an interesting challenge to open data principles, since data tends to be anonymised to some degree before being made available to the public.
-Access to the raw records is dependent on a number of requirements (e.g.
-background checks, contractual agreements, ICT security measures) which might exclude many individuals due to cost and practicality.
--->
+Specific to crime and place research, datasets of interest in this particular domain might include data describing urban structure, such as street networks, to assess whether the configuration of roads dictates things like violent crime victimization [@streetsyntax2017], or victimisation surveys to quantify citizens' perceived safety and security.
 
@@
 Indeed, many papers have made use of data collected or distributed by private organizations to explore crime and place, such as Google Street View images [@langton2017residential] and Twitter [@malleson2015impact].
-In fact, you will learn to how to obtain free Twitter data for such purposes in this book (see Chapter 6, Topic 7).
+In fact, you will learn how to obtain free Twitter data for such purposes in this book (see Chapter 6, Topic 7).
 
@@
 Finally, there is a specific group of open data sources that collate data collectively, generated by large groups of individuals who do not specifically belong to any organization, but instead work together on a collaborative project.
-These are crowdsourced data sets.
+These are crowdsourced datasets.
 Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, or Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
@@
 This data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
-Another topic in this book (see Chapter 6, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analysing open data of this type.
+Another topic in this book (see Chapter 6, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analyzing open data of this type.
 
@@
 In crime and place research, both dependent variables (e.g.
-police recorded crime incidents, victimization rates) and independent variables (e.g.
+police-recorded crime incidents, victimization rates) and independent variables (e.g.
 demographic characteristics, ambient population estimates) can be sourced from open data, whether public, private or crowdsourced.
@@
 With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science.
-In sum, researchers and analysts using open data must first ask critical questions:
+In sum, researchers and analysts using open data must ask critical questions:
 
@@
 
-The answers to these questions will put the data sources in a context of understanding, and ensure that the researchers use them appropriately and with care.
+The answers to these questions will put the data sources in context, and ensure that researchers use them appropriately and with due care.
 
-With this in mind, we now move on to a practical exercise in which we will utilize multiple sources of open data to explore police recorded crime on and around public transport in London, England.
+With this in mind, we now move on to a practical exercise in which we will utilize multiple sources of open data to explore police-recorded crime on and around public transport in London, England.
 
@@
 
-We will be using two different types of open data in this exercise: public sector data (police recorded crime data and local transport authority data) and crowdsourced data (from Open Street Map).
-To access them, we will use three different methods, namely, (1) direct download, (2) direct request to an API, 3) request to an API using a wrapper.
-This will give you a few different ideas about how you might go about accessing other open data sets relevant to your research.
+We will be using two different types of open data in this exercise: public sector data (police-recorded crime data and local transport authority data) and crowdsourced data (from Open Street Map).
+To access them, we will use three different methods, namely, (1) direct download, (2) direct request to an API, and (3) request to an API using a wrapper.
+This will give you a few different ideas about how you might go about accessing other open datasets relevant to your research.
 Locating, identifying and learning to access open data is a skill in itself.
@@
 The simplest way that open data can be made available is through direct download from a website.
-In such a case, you can visit a website, select some parameters, and save a file containing the data you requested locally on your computer.
+In such a case, you can visit a website, select some parameters, and save a file containing the data you requested locally to your computer.
 
-In the United Kingdom, police recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence.
+In the United Kingdom, police-recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence.
 Visiting this website, you will see a welcome message, and six tabs across the top which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
@@
 Take a moment to read through this information, and take notes on what you think might be relevant for your analysis.
-For example, if we want to map crimes, we will want to explore if there is any type of "anonymization" that might take place before the data are released (to protect the privacy of the victims).
+For example, if we want to map crimes, we will want to explore if there is any type of anonymization that might take place before the data are released (to protect the privacy of the victims).
 If you read the "About" page, you might find the following note:
@@
 This is a tool which defines an interface for a programme to interact with a software component.
-For example, it defines the sort of requests or calls which can be made, and how these calls and requests can be carried out.
+For example, it defines the kind of requests or calls which can be made, and how these calls and requests can be carried out.
 Here, we are using the term 'API' to denote tools created by an open data provider to give access to different subsets of their content.
@@
 
-In reality, you won't usually make these calls by pasting them into a web browser, but it is useful to see how it works.
+In reality, you will not usually make these calls by pasting them into a web browser, but it is useful to see how it works.
 More likely, we will actually make these calls from within the R environment.
@@
 
-Let's say that instead, we want data for the Northern Line. What do you think that URL will look like? Indeed, all you need to do is replace "bakerloo" with "northern".
+Let's say that instead, we want data for the Jubilee Line. What do you think that URL will look like? Indeed, all you need to do is replace "bakerloo" with "jubilee".
 
@@
 You will notice that some of the columns are actually lists, rather than character or factor vectors, which is what we might usually expect.
-Another way of exploring the `api_call` object is by looping the `class()` function through all the columns in the data frame using `lapply(api_call, class)`.
-
 This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated.
@@
 Wrappers remove (or at least lower) many of the obstacles to accessing open data noted earlier.
-The wrapper can take many forms, such as a Python module, or an R package.
+The wrapper can take many forms, such as a Python module or an R package.
 It could even be a web interface that provides a graphical user interface (GUI) for accessing the API in question.
@@
 
-Open Street Map has two types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
+Open Street Map has two main types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
 
@@
 As such, the first thing we need to specify is our study region.
-To do this, need to specify a bounding box.
+To do this, we need to specify a bounding box.
 You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the Open Street Map API that we want everything *inside* the box, but nothing *outside* the box.
@@
 One way to do this is through a search term.
-Here, we want to select Greater London, so we can use the search term "greater london united kingdom".
-Besides specifying the study region, we can also tell the `getbb()` function what format we want the data to be in.
-In this case we want a spatial object, specifically an `sf` polygon object (from the `sf` package) which we name `bb_sf`.
+Here, we want to select Greater London, so we can use the search term "greater london united kingdom" within the `getbb()` function.
+Using this function we can also specify what format we want the data to be in.
+In this case, we want a spatial object, specifically an `sf` polygon object (from the `sf` package), which we name `bb_sf`.
 
@@
 Another way to obtain the bounding box is to manually specify the latitude and longitude coordinates.
-This requires a bit of pre-existing knowledge or research about your study region.
-A quick internet search, or exploration of the interactive OSM online map (https://www.openstreetmap.org/), will give you an idea of the coordinates for your study region.
+This requires a bit of pre-existing knowledge about your study region.
+A quick internet search, or exploration of the interactive Open Street Map platform (https://www.openstreetmap.org/), will give you an idea of the coordinates for your study region.
 For Greater London, we can specify our bounding box as follows, saving the coordinates to the object `bb_gl` for later use.
@@
 We can select what features we want using the `add_osm_feature()` function, specifying our key as 'public transport' and our value as 'station'.
-We also want to specify what sort of object (what class) to get our data into, and as we're still working with spatial data, we stick to the `sf` format, for which the function is `osmdata_sf()`.
+We also want to specify what sort of object (what class) to get our data into, and as we are still working with spatial data, we stick to the `sf` format, for which the function is `osmdata_sf()`.
 Here, we specify our bounding box as the manual coordinates contained in `bb_gl`.
@@
 
-We now have an `sf` object with all the public transport stations in our study region mapped by Open Street Map volunteers, along with  ~130 variables of auxiliary data, such as the "fare_zone" the station is in, what "amenity" it may have and whether it has "toilets", amongst many others.
+We now have an `sf` object with all the public transport stations in our study region mapped by Open Street Map volunteers, along with  ~130 variables of auxiliary data, such as the `fare_zone` the station is in, what `amenity` it may have and whether it has `toilets`, amongst many others.
 Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not added information.
@@
 The variable `line` can help us with this.
-First, let's look at the values in this variable by using the `table()` function:
+First, let's explore this variable by using the `table()` function.
 
@@
 The main category ("Jubilee") is the most common, but many other combinations exist (e.g.
-Central;Jubilee) for stations which sit on multiple lines.
+"Central;Jubilee") for stations which sit on multiple lines.
 To get round this, and select all stations with "Jubilee" somewhere in the name, we can use the `grepl()` function from base R.
@@
 
-Now we have all our open data sets loaded into the R environment!
+This gives us Open Street Map data on all 27 stations on the Jubilee underground line. Now we have all our open datasets loaded into the R environment!
 
@@
 It doesn't necessarily mean that the data is in the shape needed to conduct analysis.
-The first thing to do once we have acquired open data is investigate it using data cleaning, wrangling, and visualisation.
+The first thing to do once we have acquired open data is investigate it using data cleaning, wrangling, and visualization.
 We've already loaded the `osmdata` and `dplyr` packages, but here, we will also make use of `ggplot2`, `sf`, `tidyr` and `patchwork`, so load these now.
@@
 
-### Visualise to check and compare our data sets
+```{r savedata, echo = F, message = F, warning = F, results = F, eval = F}
+# Select important columns only, as there are non-unique field names which stops save.
+temp_osm_jub_sf <- osm_jub_sf %>% 
+  select(name, geometry)
 
-We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two data sets compare.
+st_write(obj = temp_osm_jub_sf, dsn = "data/jubilee_osm.shp")
+st_write(obj = tfl_jub_sf, dsn = "data/jubilee_tfl.shp")
+```
+
+### Visualise to check and compare our datasets
+
+We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two datasets compare.
 
@@
 
-As we can see, the two data sources appear to be largely comparable in terms of their spatial patterning.
-However, the exact location of stations differs between OSM and TfL.
-Determining a single location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels.
-For operational purposes, TfL might have a specific definition for the location of a station, such as its centroid, which has been systematically applied to the dataset.
-The data pulled from OSM might have been compiled by a volunteer contributor with their own perception of how to define the location of a station, such as the ticket booth or the main entrance.
-This definition _might_ be more theoretically meaningful than that used by TfL, since it might reflect the public's perception of the station's built environment, and thus opportunities for crime.
-However, it is also plausible that the locations from OSM have been collated by multiple different contributors, each with their own idea of where each station lies on the line.
-This demonstrates why critically engaging with open data sources is so crucial, but also challenging, because it might generate lots of new questions which need to be explored.
+As we can see, the two data sources appear to be broadly comparable in terms of their spatial patterning.
+However, the exact location of stations differs between Open Street Map and TfL.
+Determining a single pinpoint location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels.
+For operational purposes, TfL might have a specific definition for the location of a station, such as its centroid, which has been systematically applied to create the coordinates.
+The data pulled from Open Street Map might have been compiled by a volunteer contributor with their own perception of how to define the location of a station, such as the ticket booth or the main entrance.
+This definition _might_ be more theoretically meaningful than that used by TfL, since it could reflect a realistic 'on the ground' perception of the station's built environment, and thus opportunities for crime.
+On the other hand, if the locations from Open Street Map were collated by multiple different contributors, each with their own idea of where each station lies on the line, we might conclude that the measure is unreliable.
+These kinds of questions demonstrate why critically engaging with open data sources is so crucial, but also challenging.
 
@@
 For that, we can turn to our British Transport Police data.
-
 In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions from the `sf` package that we used earlier.
@@
 For the purposes of this demonstration, we can define crimes occurring as 'in and around' Jubilee line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
-It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police recorded crime data [see @tompson2015uk], but it does serve to facilitate this demonstration.
+It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police-recorded crime data [see @tompson2015uk], but it does serve to facilitate this demonstration.
 
 ```{r buffer}
-# Ensure the same CRS (BNG, 27700).
+# Ensure each has the same CRS (BNG, 27700).
 osm_jub_sf <- st_transform(osm_jub_sf, st_crs(tfl_jub_sf))
@@
 
-```{r crimemapcompare}
+```{r crimemapcomparesave, echo = F}
+p1 <- ggplot(data = osm_jub_sf) +
+  geom_sf(mapping = aes(color = crimes), size = 2) +
+  labs(title = "Open Street Map") +
+  scale_color_viridis_c(lim = c(0,42))
+
+p2 <- ggplot(data = tfl_jub_sf) +
+  geom_sf(mapping = aes(color = crimes), size = 2) +
+  labs(title = "Transport for London") +
+  scale_color_viridis_c(lim = c(0,42))
+
+p12 <- p1 / p2
+
+ggsave(filename = "img/maps.png", height = 16, width = 24, unit = "cm")
+```
+
+```{r crimemapcompare, eval = F}
 p1 <- ggplot(data = osm_jub_sf) +
@@
 
-A number of factors emerge from this visualization.
-First, issues arising from the disparity between the stations identified on the Northern line are compounded.
-Using the Open Street Map data in isolation, we are grossly underestimating the number of crimes occurring in and around Northern line stations.
-This could have significant implications for how police allocate resource across the underground network.
-Secondly, as a result of this, the spatial patterning of crime on underground stations differs considerably between the two data sources.
-Using the Open Street Map data, the final station on the line, Morden, has the greatest concentration of crimes, whereas the TfL data suggests that the most problematic station is Tottenham Court Road, in the city centre.
-The conclusions drawn are starkly different, and would determine the consistency of findings with existing research examining crime at underground stations [@ceccato2013security].
-
-## Alternative example
-
-That said, this is not to say that Open Street Map is an inadequate source of open data for crime and place research.
-Firstly, it is an ever-growing resource.
-It is plausible (indeed, likely) that in the near future, the missing information about London Underground lines will be completed.
-Secondly, as noted, systematic variation in the extent to which individuals have contributed to the database can generate interesting discussions and avenues of research in its own right.
-Thirdly, this is simply one of thousands of features contained in the Open Street Map database, many of which are more complete.
-If we were interested in examining concentrations of bicycle theft from bike rental docking stations (so-called "Boris bikes") in Greater London, we could query the APIs from TfL and Open Street Map, just as we did earlier, but using slightly amended criteria.
-This is demonstrated below, all in one code chunk.
-
-```{r bikexample, eval = F}
-bb_sf <- getbb(place_name = "greater london", format_out = "sf_polygon")
-
-osm_bike_sf <- opq(bbox = bb_sf) %>%                                 
-  add_osm_feature(key = 'amenity', value = 'bicycle_rental') %>%     
-  osmdata_sf() %>%                                                  
-  trim_osmdata(bb_poly = bb_sf)
-
-osm_bike_sf <- osm_bike_sf$osm_points
-
-osm_bike_sf <- osm_bike_sf %>% 
-  filter(brand == "Santander Cycles")
-
-tfl_bike_sf <- fromJSON(readLines("https://api.tfl.gov.uk/bikepoint"))
-
-tfl_bike_sf <- tfl_bike_sf %>%
-  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326)
-
-ggplot() + 
-  geom_sf(data = osm_bike_sf) +
-  geom_sf(data = tfl_bike_sf, col = "red", alpha = 0.4)
+```{r mapfig, out.width = "\\textwidth", echo = F}
+knitr::include_graphics("img/maps.png")
 ```
 
-As showcased, the resulting visualizations are highly similar (a Spatial Point Pattern Test will confirm this, see Chapter 8, Topic 4).
-Refer back to the example earlier if some of the code is unclear.
-Note that this is a major benefit of using an API wrapper from within R, because the code used to query each feature can easily be re-used, with minor amendments, to extract different information.
+A number of factors emerge from this investigation.
+First, there is a disparity between the total number of crimes captured by the buffers in each dataset.
+Using the Open Street Map data, a total of `r sum(osm_jub_sf$crimes)` crimes were recorded across all stations on the line.
+Using the TfL data, only `r sum(tfl_jub_sf$crimes)` were recorded.
+Thus, the volume of crime observed in and around Jubilee stations during January will differ depending on the open data source being used.
+This could have important implications for how community and policing resources are distributed between underground lines.
+Second, the spatial patterning of crime hotspots on the line differs considerably between the two data sources.
+Using the Open Street Map data, the final station on the line, Stratford, has the greatest concentration of crimes, whereas the TfL data suggests that the most problematic station is Green Park, in the city centre.
+Despite the minor differences in the locations of each station, the conclusions drawn are starkly different, and would determine the consistency of findings with existing research examining crime at underground stations [@ceccato2013security].
+
+In this example, there is no obvious 'right or wrong' answer to determine which open data source is most appropriate.
+Engaging with the creators of the data would be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
+For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations.
+For Open Street Map, its online transparency means that we could find out which user(s) recorded the Jubilee station locations, and directly contact this community of volunteers to find out how station locations were recorded and why.
+Based on this information, an informed decision could be made on which data source is most appropriate, given the research question and theoretical framework being used.
 
@@
 Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources.
-By way of an example, we accessed different sources of open data in order to explore crime concentrations at London Underground stations: police recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
-We found key differences between the two sources of transport data.
-Specifically, missing information in Open Street Map resulted in a skewed picture of crime occurring in and around underground stations.
-This highlighted both the power of open data, but also its shortcomings.
-We concluded with an additional demonstration using bike rental docking stations in Greater London, which showcased a scenario in which both public sector and crowdsourced open data provide unique (and comparable) insight.
+By way of an example, we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
+Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line.
+The demonstration highlighted both the power of open data, but the need to critically engage with sources.
 

### 2020-09-30 scripts/osm_explore_tube.R

@@ -19,7 +19,10 @@ library(stringr)
 # no longer works.
 # bb_sf <- getbb(place_name = "Greater London United Kingdom", format_out = "sf_polygon")
 
-osm_stat_sf <- opq(bbox = c(-0.5103751, 51.2867602, 0.3340155, 51.6918741)) %>% # select bounding box
+
+bb_gl <- c(-0.51037, 51.28676, 0.33401, 51.69187)
+
+osm_stat_sf <- opq(bbox = bb_gl) %>% # select bounding box
   add_osm_feature(key = 'public_transport', value = 'station') %>%           # select features
   osmdata_sf()                                                               # specify class (sf or sp)
   # trim_osmdata(bb_poly = bb_sf)                                           # trim by bounding box  

### 2020-11-13 draft_chapter.Rmd
straford map

@@
 
-**Corresponding author:** Samuel Langton, Manchester Metropolitan University, Crime and Well-Being
-
-Big Data Centre, Geoffrey Manton Building, Manchester, M16 6EB, United Kingdom.
-
-Telephone: +44 (0)161 247 2000. Email: s.langton@mmu.ac.uk.
+**Corresponding author:** Samuel Langton, University of Leeds, School of Law, Leeds, United Kingdom. Email: s.langton@leeds.ac.uk.
 
@@
 
-\dag Department of Sociology, Manchester Metropolitan University, Manchester, United Kingdom. ORCID: 0000-0002-1322-1553.
+\dag School of Law, University of Leeds, Leeds, United Kingdom. ORCID: 0000-0002-1322-1553.
 
@@
 Increasingly, researchers are making use of online repositories for their papers (e.g.
-https://osf.io/), data (e.g.
-https://www.ukdataservice.ac.uk/deposit-data) and code (e.g.
-https://github.com/), all in the name of transparency and replication.
+https://arxiv.org/), data (e.g.
+https://www.ukdataservice.ac.uk/deposit-data), code (e.g.
+https://github.com/), or whole projects (e.g.
+https://osf.io/) all in the name of transparency and replication.
 These are guidelines and tools worth considering, not just when conducting analysis on secondary data sources, but also collecting and sharing your own data.
@@
 This will allow us to explore the strengths and limitations of public sector and crowdsourced open data sources.
+Should you struggle to download any of the data during the exercise, or if the data sources have been updated since publishing, each one has been archived and is available on the corresponding GitHub page for this topic (https://github.com/langtonhugh/osm_crim).
 
@@
 ```{r crimeload_gh, eval = FALSE}
-btp_df <- read_csv("https://raw.githubusercontent.com/langtonhugh/osm_crim/
-                   master/data/2020-01-metropolitan-street.csv")
+btp_df <- read_csv("https://github.com/langtonhugh/osm_crim/raw/master/
+                   data/2020-01-btp-street.csv")
 ```
@@
 These kinds of questions demonstrate why critically engaging with open data sources is so crucial, but also challenging.
+We return to this at the end of the exercise.
 
@@
 For that, we can turn to our British Transport Police data.
-In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions from the `sf` package that we used earlier.
+In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions from the `sf` package that we used earlier, after first dropping crimes that have incomplete coordinates.
 
-```{r crimespatial, echo=F}
+```{r crimespatial}
 btp_sf <- btp_df %>%
@@
 
-We can then visualize these counts and compare the two sources of data by coloring in our point locations according to the crime count.
+We can then visualize these counts and compare the two sources of data by coloring in our buffers according to the crime count.
 Note that we arrange the plots using syntax available using the `patchwork` package, and set the scales to be comparable using the `lim` argument within `scale_color_viridis_c()`.
@@
 
-In this example, there is no obvious 'right or wrong' answer to determine which open data source is most appropriate.
-Engaging with the creators of the data would be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
+We can explore Stratford station as a potential crime hotspot using some of the skills we have learnt so far.
+One useful first step might to visualize the built environment around the station, along with the crime buffers from Open Street Map and TfL respectively, and the crime locations themselves.
+This way, we might find out why the TfL buffer is failing to capture any crimes, while the Open Street Map buffer is capturing an apparent crime hotspot.
+In doing so, it is worth remembering that crime locations in open police data are approximation to ensure anonymity.
+
+```{r, eval = F}
+# Retrieve Stratford buffer from Open Street Map.
+stratford_osm_sf <- osm_jub_sf %>% 
+  filter(name == "Stratford")
+
+# Do the same using TfL.
+stratford_tfl_sf <- tfl_jub_sf %>% 
+  filter(commonName == "Stratford Underground Station")
+
+# Create a mini study area for our visualization, by getting the bounding box for the
+# Open Street Map (or TfL) buffer. Note that we transform to WGS 84 first in preperation
+# for the Open Street Map API.
+bb_sf <- stratford_osm_sf %>%
+  st_transform(crs = 4326) %>% 
+  st_bbox()
+
+# Retrieve the coordinates of the bounding box, and create a numeric list for use with
+# osmdata package.
+strat_bb <- c(bb_sf[[1]], bb_sf[[2]], bb_sf[[3]], bb_sf[[4]])
+
+# Pull building footprints in this bounding box from the Open Street Map API.
+strat_fp_sf <- opq(strat_bb) %>%
+    add_osm_feature(key = 'building') %>% 
+    osmdata_sf()
+
+# Get the polygons only.
+strat_fp_poly_sf <- strat_fp_sf$osm_polygons
+
+# Transform back to British National Grid to match our other data.
+strat_fp_poly_sf <- st_transform(strat_fp_poly_sf, 27700)
+
+# Clip the crime data to the Open Street Map buffer. Remember that the TfL buffer
+# did not capture any crimes.
+strat_osm_btp_sf <- st_intersection(stratford_osm_sf, btp_sf)
+
+# Visualize the building footprints, Open Street Map buffer, TfL buffer, and crime
+# locations. Note that the single point for the crimes is due to the approximation!
+# There are actually multiple crimes.
+ggplot() +
+  theme_bw() +
+  geom_sf(data = strat_fp_poly_sf, fill = "black") +
+  geom_sf(data = stratford_osm_sf, fill = "pink", alpha = 0.3) +
+  geom_sf(data = stratford_tfl_sf, fill = "blue", alpha = 0.3) +
+  geom_sf(data = strat_osm_btp_sf, col = "red") 
+```
+
+```{r strat, echo = F}
+knitr::include_graphics("img/stratford.png")
+```
+
+This quick check has shed some light on what is happening.
+The buffer generated from the TfL station location (in blue) covers a different part of the station to the Open Street Map buffer (in pink).
+The latter covers most of the main station building, while the former centers around the platforms to the north.
+Because the British Transport Police crime locations have been geocoded to a location in the main station building (in red), the TfL buffer has missed them.
+If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station.
+There is no obvious 'right or wrong' answer to determine which open data source is most appropriate.
+Here, we might consider investigating other Jubilee stations before deciding which data to use, and ensure that findings are reported transparently with reference to these discussions.
+
+In many cases, engaging with the creators of the data would be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
 For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations.
 For Open Street Map, its online transparency means that we could find out which user(s) recorded the Jubilee station locations, and directly contact this community of volunteers to find out how station locations were recorded and why.
-Based on this information, an informed decision could be made on which data source is most appropriate, given the research question and theoretical framework being used.
 
@@
 Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line.
-The demonstration highlighted both the power of open data, but the need to critically engage with sources.
+The demonstration highlighted both the power of open data and the need to critically engage with sources.
 
@@
 The number and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology.
-For instance, information which traces building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
+For instance, information about building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
 But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go.

### 2020-11-13 scripts/osm_explore_tube.R

@@ -64,12 +64,12 @@ ggplot() +
   geom_sf(data = osm_jub_sf) +
   geom_sf(data = tfl_jub_sf, color = "red", alpha = 0.5) 
 
-# Save before some nerd updates the line like the jubern line did.
-temp_osm_jub_sf <- osm_jub_sf %>% 
-  select(name, geometry)
-
-st_write(obj = temp_osm_jub_sf, dsn = "data/jubilee_osm.shp")
-st_write(obj = tfl_jub_sf, dsn = "data/jubilee_tfl.shp")
+# Save before someone changes the lines.
+# temp_osm_jub_sf <- osm_jub_sf %>% 
+#   select(name, geometry)
+# 
+# st_write(obj = temp_osm_jub_sf, dsn = "data/jubilee_osm.shp")
+# st_write(obj = tfl_jub_sf, dsn = "data/jubilee_tfl.shp")
 
 # Assessing the impact in terms of crime. BTP data in January 2020.
 btp_df <- read_csv("data/2020-01-btp-street.csv")
@@ -109,16 +109,90 @@ p1 / p2
 sum(osm_jub_sf$crimes) # 168
 sum(tfl_jub_sf$crimes) # 136
 
+# Which is the crime hotspot at the end of the line?
+osm_jub_sf %>% 
+  arrange(desc(crimes)) %>% 
+  slice(1) %>% 
+  pull(name) 
+# Straford.
+
+# let's plot a comparison map of this station location. We can get building footprints.
+stratford_osm_sf <- osm_jub_sf %>% 
+  arrange(desc(crimes)) %>% 
+  slice(1)
+
+stratford_osm_sf <- osm_jub_sf %>% 
+  filter(name == "Stratford")
+
+# Get the TFL Stratford buffer (no crimes captured).
+stratford_tfl_sf <- tfl_jub_sf %>% 
+  filter(commonName == "Stratford Underground Station")
+
+# extract bb of stratford to create a mini study region.
+bb_sf <- stratford_osm_sf %>%
+  st_transform(crs = 4326) %>% 
+  st_bbox() 
+
+# Retrieve coordinates into list.
+strat_bb <- c(bb_sf[[1]], bb_sf[[2]], bb_sf[[3]], bb_sf[[4]])
+
+# Get building footprints.
+strat_fp_sf <- opq(strat_bb) %>%
+    add_osm_feature(key = 'building') %>% 
+    osmdata_sf()
+
[…]

### 2020-11-15 draft_chapter.Rmd
stratford text update

@@
 This way, we might find out why the TfL buffer is failing to capture any crimes, while the Open Street Map buffer is capturing an apparent crime hotspot.
-In doing so, it is worth remembering that crime locations in open police data are approximation to ensure anonymity.
+In doing so, it is worth remembering that crime locations in open police data are approximations of the real geography in order to ensure anonymity.
 
@@
 If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station.
-There is no obvious 'right or wrong' answer to determine which open data source is most appropriate.
-Here, we might consider investigating other Jubilee stations before deciding which data to use, and ensure that findings are reported transparently with reference to these discussions.
+In many cases, there will not be an obvious 'right or wrong' answer to determine which open data source is most appropriate, but investigations like this one will help inform the decision.
+We might consider investigating other stations before deciding on which data to use, and ensure that findings are reported transparently with reference to these discussions.
 
-In many cases, engaging with the creators of the data would be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
+Engaging with the creators of the data would also be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
 For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations.

### 2020-11-16 draft_chapter.Rmd
update stratford example & pre print version

@@
 
-## Cleaning, wrangling, and visualising our data
+## Cleaning, wrangling, and visualizing our data
 
@@
   geom_sf(data = tfl_jub_sf) +
-  geom_sf(data = osm_jub_sf, color = "red", alpha = 0.5) 
+  geom_sf(data = osm_jub_sf, col = "red", alpha = 0.5) 
 ```
@@
 As we can see, the two data sources appear to be broadly comparable in terms of their spatial patterning.
-However, the exact location of stations differs between Open Street Map and TfL.
+However, the exact location of stations sometimes differs between Open Street Map and TfL.
 Determining a single pinpoint location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels.
@@
 # locations. Note that the single point for the crimes is due to the approximation!
-# There are actually multiple crimes.
+# There are actually multiple crimes. Colors have been manually selected
+# from the viridis package to be easier to read by those with colorblindness.
 ggplot() +
-  theme_bw() +
   geom_sf(data = strat_fp_poly_sf, fill = "black") +
-  geom_sf(data = stratford_osm_sf, fill = "pink", alpha = 0.3) +
-  geom_sf(data = stratford_tfl_sf, fill = "blue", alpha = 0.3) +
-  geom_sf(data = strat_osm_btp_sf, col = "red") 
+  geom_sf(data = stratford_osm_sf, fill = "#440154FF", alpha = 0.3) +
+  geom_sf(data = stratford_tfl_sf, fill = "#FDE725FF", alpha = 0.3) +
+  geom_sf(data = strat_osm_btp_sf, col = "#3CBC75FF", fill = "white") +
+  theme(axis.text = element_text(size = 6))
 ```
@@
 This quick check has shed some light on what is happening.
-The buffer generated from the TfL station location (in blue) covers a different part of the station to the Open Street Map buffer (in pink).
+The buffer generated from the TfL station location (in yellow) covers a different part of the station to the Open Street Map buffer (in blue).
 The latter covers most of the main station building, while the former centers around the platforms to the north.
-Because the British Transport Police crime locations have been geocoded to a location in the main station building (in red), the TfL buffer has missed them.
+Because the British Transport Police crime locations have been geocoded to a location in the main station building (in green) the TfL buffer has missed them.
 If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station.
@@
 
+# Software
+
+As noted in the tutorial, a number of R packages were used to create the material in this chapter.
+Packages _readr_ [@readrpack] and _jsonlite_ [@jsonlitepack] were used for reading and parsing data, _osmdata_ for querying the Open Street Map API [@osmdatapack], _sf_ for handling and visualizing spatial data [@sfpack], _dplyr_ and _tidyr_ for wrangling data [@dplyrpack; @tidyrpack], _ggplot2_ for visualization [@ggplotpack] and _patchwork_ for arranging plots [@patchworkpack].
+The document itself was prepared using _rmarkdown_ [@rmarkdownpack].
+
+\newpage
+
 # References

### 2020-11-16 scripts/osm_explore_tube.R

@@ -155,12 +155,12 @@ strat_fp_poly_sf <- st_transform(strat_fp_poly_sf, 27700)
 
 # Quick plot.
 strat_gg <- ggplot() +
-  # theme_bw() +
   geom_sf(data = strat_fp_poly_sf, fill = "black") +
-  geom_sf(data = stratford_osm_sf, fill = "pink", alpha = 0.3) +
-  geom_sf(data = stratford_tfl_sf, fill = "blue", alpha = 0.3) +
-  geom_sf(data = strat_osm_btp_sf, col = "Red") +
+  geom_sf(data = stratford_osm_sf, fill = "#440154FF", alpha = 0.3) +
+  geom_sf(data = stratford_tfl_sf, fill = "#FDE725FF", alpha = 0.3) +
+  geom_sf(data = strat_osm_btp_sf, col = "#3CBC75FF") +
   theme(axis.text = element_text(size = 6))
+strat_gg
 
 ggsave(plot = strat_gg, file = "img/stratford.png", width = 12, height = 14, unit = "cm")  
 

### 2020-11-18 drafts/preprint.Rmd
revisions v1 no tiffs

--- 2020-11-16 draft_chapter.Rmd
@@
   - Reka Solymosi\ddag
+abstract: "This chapter provides a framework to approach and meaningfully interpret open data.
+The chapter also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in R in order to draw conclusions about crime and place."
 editor_options:

### 2020-11-18 manuscript.Rmd
revisions v1 no tiffs

--- 2020-11-15 draft_chapter.Rmd
@@
 author:
-  - Samuel Langton\dag
-  - Reka Solymosi\ddag
+  - Samuel Langton (University of Leeds)
+  - Reka Solymosi (University of Manchester)
 editor_options:
@@
 output:
-  pdf_document:
-    latex_engine: xelatex
-  word_document: default
+  word_document:
+    reference_docx: reference_doc.docx
 bibliography: refs.bib
@@
 knitr::opts_chunk$set(message = F, warning = F, eval = T, echo = T)
-# library(jsonlite)
-# library(osmdata)
-# library(ggplot2)
-# library(sf)
-# library(readr)
-# library(tidyr)
-# library(patchwork)
-# library(dplyr)
 ```
-
-\enspace
-
-This is an unedited draft, not for citation, of a book chapter that has been accepted for publication in E.
-Groff and C.
-Haberman (Eds.), The study of crime and place: A methods handbook (Temple University Press, 2021).
-Temple University Press retains all rights and subrights of this document.
-This pre-print is not published under a CC license.
-
-\enspace
 
@@
 **Corresponding author:** Samuel Langton, University of Leeds, School of Law, Leeds, United Kingdom. Email: s.langton@leeds.ac.uk.
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-***
-
-\dag School of Law, University of Leeds, Leeds, United Kingdom. ORCID: 0000-0002-1322-1553.
-
-\ddag School of Social Sciences, University of Manchester, Manchester, United Kingdom. ORCID: 0000-0001-8689-1526.
-
-\newpage
 
@@
 This chapter will outline how to develop such skills by providing a framework to approach and meaningfully interpret open data.
-The chapter also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.
+The supplementary material also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.
 
@@
 Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
-Contributions to resources such as Open Street Map (which we look at in the practical exercise later) are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
+Contributions to resources such as Open Street Map (which we look at in the practical exercise) are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
 These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
@@
 
-With this in mind, we now move on to a practical exercise in which we will utilize multiple sources of open data to explore police-recorded crime on and around public transport in London, England.
-
-# Practical exercise
-
-In this exercise you will acquire a number of different skills, including:
-
-- Accessing open data using three different methods: 
-    + Direct download.
-    + Direct calls to an Application Programming Interface (API).
-    + Calls to an API using a wrapper.
-- Cleaning, wrangling, and visualizing open spatial data.
-- Comparing different sources of open data.
-- How to engage with the critical 'Where', 'Why', 'Who', 'What' and 'When' questions to better understand open data.
-
-## Our aim
-
-In this exercise we will explore crime in and around London Underground stations.
-We know that environmental features are important when it comes to public transport areas being more or less criminogenic.
-Studies in Sweden, for instance, have shown the importance of environmental and neighborhood characteristics in determining crime concentrations at underground stations in Stockholm, along with the positioning of stations on the line [@ceccato2013security].
-Similar work has been carried out exploring bus stops in Los Angeles [@loukaitou1999hot] and the impact of intensive policing along bus corridors in Merseyside, England [@newton2004crime], amongst others.
-
-Here, we will consider the case of London.
-Specifically, we will examine the question: to what extent does crime cluster in and around London Underground stations?
-We will use various sources of open data to answer this question.
-This will allow us to explore the strengths and limitations of public sector and crowdsourced open data sources.
-Should you struggle to download any of the data during the exercise, or if the data sources have been updated since publishing, each one has been archived and is available on the corresponding GitHub page for this topic (https://github.com/langtonhugh/osm_crim).
-
-## Accessing data
-
-We will be using two different types of open data in this exercise: public sector data (police-recorded crime data and local transport authority data) and crowdsourced data (from Open Street Map).
-To access them, we will use three different methods, namely, (1) direct download, (2) direct request to an API, and (3) request to an API using a wrapper.
-This will give you a few different ideas about how you might go about accessing other open datasets relevant to your research.
-Locating, identifying and learning to access open data is a skill in itself.
-
-### Direct download
-
-The simplest way that open data can be made available is through direct download from a website.
-In such a case, you can visit a website, select some parameters, and save a file containing the data you requested locally to your computer.
-
-In the United Kingdom, police-recorded crime data in England and Wales can be accessed this way using an online web portal (https://data.police.uk/) under Open Government Licence.
-Visiting this website, you will see a welcome message, and six tabs across the top which should read "Home", "Data", "API", "Changelog", "Contact", and "About".
-
-Before we download any data, we can learn more about it by clicking on the "About" tab.
-This brings up information that is important to review carefully in order to answer the where, why, who, what and when questions posed earlier.
-Take a moment to read through this information, and take notes on what you think might be relevant for your analysis.
-For example, if we want to map crimes, we will want to explore if there is any type of anonymization that might take place before the data are released (to protect the privacy of the victims).
-If you read the "About" page, you might find the following note:
-
-> Location anonymization.
-The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.
-
-This indicates that although we get a latitude and longitude coordinate with each crime event, it may only be approximate. This may have implications for our findings!
-
-To then download some data, move on to clicking on the "Data" tab.
-This should open a page entitled "Data downloads" under which you can see another five tabs: "Custom download", "Archive", "Boundaries", "Open data", and "Statistical data".
-By staying on this "Custom download" page you can select what sort of data you want to download.
-We can select the time period and police force of interest, and the type of information required (e.g.
-crimes, stop and search, outcomes).
-We are also informed that the data are downloaded in comma-separated values format (.csv file extension), which meets our machine-readable, easy-to-manipulate data format requirements.
-
-For this exercise, we are going to use British Transport Police data, a force which operates on railways and light-rail systems across the country, for the month of January in 2020.
-Select the time period using the dropdown menus, the force using the tickboxes, and then generate and download the file.
-
-Save this file locally (in your working directory) to a subfolder named "data".
-You can load it in with the `read_csv()` command from the `readr` package.
-Remember that you will need to load it first using `library()`.
-
-```{r crimeload1}
-library(readr)
-
-btp_df <- read_csv("data/2020-01-btp-street.csv")
-```
-
-Note: If you didn't manage to follow along with the download instructions, you can also access this dataset from our GitHub page.
-Note that the line break below is purely for easy formatting - this will need removing in your own script so that the URL runs across one line.
-
-```{r crimeload_gh, eval = FALSE}
-btp_df <- read_csv("https://github.com/langtonhugh/osm_crim/raw/master/
-                   data/2020-01-btp-street.csv")
-```
-
-You should now have a data frame with `r nrow(btp_df)` crimes in, ready for you to explore!
-
-### Direct request to an API
-
-Another way of downloading open data is through an Application Programming Interface (API).
-This is a tool which defines an interface for a programme to interact with a software component.
-For example, it defines the kind of requests or calls which can be made, and how these calls and requests can be carried out.
-Here, we are using the term 'API' to denote tools created by an open data provider to give access to different subsets of their content.
-Such APIs facilitate scripted and programmatic extraction of content, as permitted by the API provider [@olmedilla2016harvesting].
-APIs can take many different forms and be of varying quality and usefulness [@foster2016big].
-For the purposes of accessing open data from the web, we are specifically talking about _RESTful_ APIs.
-The 'REST' stands for Representational State Transfer.
-These APIs work directly over the web, which means users can play with the API with relative ease in order to understand how it works [@foster2016big].
-
-Here, we will make a direct request to the API created by London's local government organization responsible for transport: Transport for London (TfL).
-TfL oversee the London Underground.
-They provide access to their open data through a unified API.
-Much like how we found the "About" page when downloading police data directly, to answer our 'Where', 'Why', 'Who', 'What' and 'When' questions, we must find a similar document for TfL.
-Details about the TfL API are provided through their open data page (https://tfl.gov.uk/info-for/open-data-users/unified-api), which explains what data is available, and how the API is designed.
-
-They further provide a documentation page with examples of how you can use HTTP (i.e.
-web link) requests to make calls to the API (see: https://api-portal.tfl.gov.uk/docs).
-If you follow this link you should see a page titled "Our Unified API" and some examples of calls.
-To draw conclusions about crimes in and around London Underground stations, we will need some spatial data about the stations themselves.
-
-In the documentation, there is a section called 'API area' which offers some guidance on the types of data available.
-Scroll down and find the example called _Stops_.
-You will see that this call returns information on stops for buses or London Underground lines, and there are two examples (bus route 24 and the Bakerloo line).
-Copy their demonstration URL for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser.
-When you visit this page, you should see something like Figure 1.
-
-```{r tfl, out.width = "\\textwidth", fig.cap="example of data from Transport for London", echo = F}
-knitr::include_graphics("img/tfl_screenshot.png")
-```
-
-Although this contains the information we requested, it is not very legible for our human eyes.
-It is in a format called _JSON_.
-This stands for JavaScript Object Notation.
-JSON is an open standard way of storing and exchanging data, and will most likely be the format in which data are returned from most API calls.
-As you can see, it is not intuitive to read in its current format _but_ it is machine-readable friendly, which again aligns well with our requirements for open data.
-
-In reality, you will not usually make these calls by pasting them into a web browser, but it is useful to see how it works.
-More likely, we will actually make these calls from within the R environment.
-That said, it might be useful (and interesting) to examine how basic queries are constructed using this documentation.
-For example, we now know that to request the stops from the "Bakerloo" line, we need the URL:
-
-`"https://api.tfl.gov.uk/line/bakerloo/stoppoints"`
-
-Let's say that instead, we want data for the Jubilee Line. What do you think that URL will look like? Indeed, all you need to do is replace "bakerloo" with "jubilee".
-
-In R, we can use the `fromJSON()` function from the `jsonlite` package to parse all this information into a data frame (with rows and columns).
-It will then become more familiar and usable.
-All we need to do is input the URL from the TfL API into the function with a bit of help from `readLines()`, a function in base R for reading text from URLs.
-
-To keep things focused, let's request data about stop points on the Jubilee line, one of the busiest on the network.
-Note how this is simply an amended version of the example provided by TfL in the API documentation.
-
-```{r getapi}
-library(jsonlite)
-
-api_call <- fromJSON(readLines("https://api.tfl.gov.uk/line/jubilee/stoppoints"))
-```
-
-This gives us an object (`api_call`) which contains all the information returned by the TfL API.
-JSON is slightly different to traditional data frames with rows and columns, which we are probably more familiar with, because the data is nested.
-For instance, `api_call` is classed as a data frame, but now try viewing the object using `View(api_call)`.
-You will notice that some of the columns are actually lists, rather than character or factor vectors, which is what we might usually expect.
-This demonstrates an important challenge faced by researchers when using open data, because dealing with data in this format can be messy and complicated.
-It is not always a neatly formatted data frame like the CSV file from the open police data portal.
-That said, we can transform this data into something more familiar from within R, as we will see later on.
-
-### Request to an API using a wrapper
-
-Finally, we will look at how open data can be obtained using API wrappers.
-Often, developers who work with APIs will share their code, and release them in the form of a package or module, so that other people can use it.
-This is called a _wrapper_ because it uses code that 'wraps' around the API to make it a neater, more usable package.
-Wrappers remove (or at least lower) many of the obstacles to accessing open data noted earlier.
-The wrapper can take many forms, such as a Python module or an R package.
-It could even be a web interface that provides a graphical user interface (GUI) for accessing the API in question.
-
-To demonstrate this, we will be accessing data from Open Street Map, a database of geospatial information built by a community of mappers, enthusiasts and members of the public, who contribute and maintain data about all sorts of environmental features, such as roads, green spaces, restaurants and railway stations, amongst many other things, all over the world.
-As such, it is a prime example of 'crowdsourced' open data.
-You can view the information contributed to Open Street Map using their online mapping platform (https://www.openstreetmap.org/).
-The result of people's contributions is a database of spatial information rich in local knowledge which provides invaluable information about places and their features, without being subject to strict terms on usage.
-
-Open Street Map has two main types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
-
-If we load the package `osmdata` we can use its functions to query the Open Street Map API, rather than the API query being made directly (like we did for TfL, using the URL).
-Once again, we will want to identify documentation which can help us understand and critique our data, and learn how to query it.
-When it comes to Open Street Map, you can read all about it on their community page: [https://www.openstreetmap.org/about](https://www.openstreetmap.org/about).
-Using the wrapper in R, we can refer to the package documentation and the associated vignette online: [https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html).
-
-Unlike the TfL API, `osmdata` is an international database, and has lots of data that we might not necessarily need.
-As such, the first thing we need to specify is our study region.
-To do this, we need to specify a bounding box.
-You can think of the bounding box as a box drawn around the area that we are interested in (in this case, London, England) which tells the Open Street Map API that we want everything *inside* the box, but nothing *outside* the box.
-
-So, how can we name a bounding box specification to define the study region?
-One way to do this is through a search term.
-Here, we want to select Greater London, so we can use the search term "greater london united kingdom" within the `getbb()` function.
-Using this function we can also specify what format we want the data to be in.
-In this case, we want a spatial object, specifically an `sf` polygon object (from the `sf` package), which we name `bb_sf`.
-
-```{r bb, message=F, warning=F}
-library(osmdata)
-library(sf)
-
-bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon")
-```
-
-One downside of using search terms to define the bounding box, as with any search term, is that inconsistencies in terminology, wordings and spelling might lead to unexpected outputs (or none at all).
-Another way to obtain the bounding box is to manually specify the latitude and longitude coordinates.
-This requires a bit of pre-existing knowledge about your study region.
-A quick internet search, or exploration of the interactive Open Street Map platform (https://www.openstreetmap.org/), will give you an idea of the coordinates for your study region.
-For Greater London, we can specify our bounding box as follows, saving the coordinates to the object `bb_gl` for later use.
-
-```{r bbmanual}
-bb_gl <- c(-0.51037, 51.28676, 0.33401, 51.69187) # xmin, ymin, xmax, ymax
-```
-
-We can now move on to query data from the Open Street Map API using the `opq()` function.
-The function name is short for 'Overpass query', which is how users can query the Open Street Map API using search criteria.
-
-Besides specifying what area we want to query with our bounding box object(s) in the `opq()` function, we must also define the feature which we want to pull from the API.
-Features in Open Street Map are defined through 'keys' and 'values'.
-Keys are used to describe a broad category of features (e.g.
-highway, amenity), and values are more specific descriptions (e.g.
-cycleway, bar).
-These are tags which contributors to Open Street Map have defined.
-A useful way to explore these is by using the comprehensive Open Street Map Wiki page on map features (https://wiki.openstreetmap.org/wiki/Map_Features).
-
-We can select what features we want using the `add_osm_feature()` function, specifying our key as 'public transport' and our value as 'station'.
-We also want to specify what sort of object (what class) to get our data into, and as we are still working with spatial data, we stick to the `sf` format, for which the function is `osmdata_sf()`.
-Here, we specify our bounding box as the manual coordinates contained in `bb_gl`.
-^[If you use the bounding box obtained through `getbb()` one can subsequently trim down the outputs from `add_osm_feature()` using `trim_osmdata()`.
-For instance, we could add `trim_osmdata(bb_poly = bb_sf)` to our initial query.]
-
-```{r opq}
-osm_stat_sf <- opq(bbox = bb_gl) %>%                               # select bounding box
-  add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
-  osmdata_sf()                                                     # specify class
-```
-
-The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the **Console**.
-
-```{r printstations}
-osm_stat_sf
-```
-
-This confirms details like the bounding box coordinates, but also provides information on the features collected from the query.
-As one might expect, most information relating to public transport station locations has been recorded using points (i.e.
-two-dimensional vertices, coordinates) of which we have `r nrow(osm_stat_sf$osm_points)` at the time of writing.
-We also have around fifty polygons.
-For now, let's extract the point information.
-
-```{r stationpoints}
-osm_stat_sf <- osm_stat_sf$osm_points 
-```
-
-We now have an `sf` object with all the public transport stations in our study region mapped by Open Street Map volunteers, along with  ~130 variables of auxiliary data, such as the `fare_zone` the station is in, what `amenity` it may have and whether it has `toilets`, amongst many others.
-Of course, it is up to the volunteers whether they collect all these data, and in many cases, they have not added information.
-Nevertheless, when the details are recorded, they provide rich insight and local knowledge that we may otherwise be unable to obtain.
-
-One additional step needed is to select the stations which are relevant to us (i.e.
-they fall along the Jubilee Line).
-The variable `line` can help us with this.
-First, let's explore this variable by using the `table()` function.
-
-```{r valuesofline, eval = F}
-table(osm_stat_sf$line)
-```
-
-You will see that "Jubilee" appears several times under different line names.
-The main category ("Jubilee") is the most common, but many other combinations exist (e.g.
-"Central;Jubilee") for stations which sit on multiple lines.
-To get round this, and select all stations with "Jubilee" somewhere in the name, we can use the `grepl()` function from base R.
-Note that we also make use of `filter()` from within `dplyr`, so we first need to load that package.
-
-```{r filterline}
-library(dplyr)
-
-osm_jub_sf <- osm_stat_sf %>% 
-  filter(grepl("Jubilee", line))
-```
-
-This gives us Open Street Map data on all 27 stations on the Jubilee underground line. Now we have all our open datasets loaded into the R environment!
-
-## Cleaning, wrangling, and visualising our data
-
-Obtaining open data from a direct download or API is often only half the battle.
-It doesn't necessarily mean that the data is in the shape needed to conduct analysis.
-The first thing to do once we have acquired open data is investigate it using data cleaning, wrangling, and visualization.
-We've already loaded the `osmdata` and `dplyr` packages, but here, we will also make use of `ggplot2`, `sf`, `tidyr` and `patchwork`, so load these now.
-Remember, if you do not have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.
-
-```{r packages}
-library(ggplot2)
-library(sf)
-library(tidyr)
-library(patchwork)
-```
-
-### From JSON to data frame
-
-We usually want to work with data that is in the format of a data frame, that is, all your observations are rows and all your variables are columns.
-While this is the case for our police data and our Open Street Map data, recall that the TfL data came in the nested JSON format.
-
-Fortunately for us, R is more than capable of dealing with nested data.
-Moreover, we can see that the information we are interested in (namely, the location of underground stations) has already been successfully parsed into columns called **lon** (longitude) and **lat** (latitude), along with an identification column of the station names called **commonName**.
-We can extract these columns, and create an `sf` (spatial) point object using these coordinates, all in one chunk of code, using the piping operator (`%>%`).
-Note that we define the Coordinate Reference System (CRS) which the data has come in (WGS 84), and then transform it into something more appropriate (the British National Grid, a projected CRS used in the United Kingdom).
-
-```{r apispatial}
-tfl_jub_sf <- api_call %>% 
-  select(commonName, lat, lon) %>% 
-  st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326) %>% 
-  st_transform(27700)
-```
-
-We can now easily visualize this data using the `ggplot2` package, which is compatible with `sf` objects.
-
-```{r tflmap}
-ggplot(tfl_jub_sf) +
-  geom_sf()
-```
-
-There we have it: with just a few lines of code in R, we have queried the TfL API, a public sector source of open data, and plotted a basic map of stations on the Jubilee underground line.
-
-```{r savedata, echo = F, message = F, warning = F, results = F, eval = F}
-# Select important columns only, as there are non-unique field names which stops save.
-temp_osm_jub_sf <- osm_jub_sf %>% 
-  select(name, geometry)
-
-st_write(obj = temp_osm_jub_sf, dsn = "data/jubilee_osm.shp")
-st_write(obj = tfl_jub_sf, dsn = "data/jubilee_tfl.shp")
-```
-
-### Visualise to check and compare our datasets
-
-We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two datasets compare.
-
-```{r mapline}
-ggplot() +
-  geom_sf(data = tfl_jub_sf) +
-  geom_sf(data = osm_jub_sf, color = "red", alpha = 0.5) 
-```
-
-As we can see, the two data sources appear to be broadly comparable in terms of their spatial patterning.
-However, the exact location of stations differs between Open Street Map and TfL.
-Determining a single pinpoint location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels.
-For operational purposes, TfL might have a specific definition for the location of a station, such as its centroid, which has been systematically applied to create the coordinates.
-The data pulled from Open Street Map might have been compiled by a volunteer contributor with their own perception of how to define the location of a station, such as the ticket booth or the main entrance.
-This definition _might_ be more theoretically meaningful than that used by TfL, since it could reflect a realistic 'on the ground' perception of the station's built environment, and thus opportunities for crime.
-On the other hand, if the locations from Open Street Map were collated by multiple different contributors, each with their own idea of where each station lies on the line, we might conclude that the measure is unreliable.
-These kinds of questions demonstrate why critically engaging with open data sources is so crucial, but also challenging.
-We return to this at the end of the exercise.
-
-Given the differences observed between the two open data sources, what might the impact be when studying crime in and around London Underground stations?
-For that, we can turn to our British Transport Police data.
-In its raw form, this data is a standard data frame, but we can make it spatial using some of the functions from the `sf` package that we used earlier, after first dropping crimes that have incomplete coordinates.
-
-```{r crimespatial}
-btp_sf <- btp_df %>%
-  drop_na(Longitude, Latitude) %>%
-  st_as_sf(coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
-  st_transform(27700)
-```
-
-We now have the recorded location of all crimes recorded by the British Transport Police during January, 2020.
-For the purposes of this demonstration, we can define crimes occurring as 'in and around' Jubilee line tube stations by creating a 50 meter buffer around each station, for each source of data, and counting the number of points falling within each.
-It is worth noting that this definition is somewhat arbitrary, and is subject to the spatial inaccuracy in open police-recorded crime data [see @tompson2015uk], but it does serve to facilitate this demonstration.
-
-```{r buffer}
-# Ensure each has the same CRS (BNG, 27700).
-osm_jub_sf <- st_transform(osm_jub_sf, st_crs(tfl_jub_sf))
-
-# Create buffers to define 'in and around'.
-osm_buff_sf <- st_buffer(osm_jub_sf, dist = 50)
-tfl_buff_sf <- st_buffer(tfl_jub_sf, dist = 50)
-
-# Count number of crimes recorded within each buffer.
-osm_jub_sf <- osm_buff_sf %>% 
-  mutate(crimes = lengths(st_intersects(osm_buff_sf, btp_sf)))
-tfl_jub_sf <- tfl_buff_sf %>% 
-  mutate(crimes = lengths(st_intersects(tfl_buff_sf, btp_sf)))
-```
-
-We can then visualize these counts and compare the two sources of data by coloring in our buffers according to the crime count.
-Note that we arrange the plots using syntax available using the `patchwork` package, and set the scales to be comparable using the `lim` argument within `scale_color_viridis_c()`.
-
-```{r crimemapcomparesave, echo = F}
-p1 <- ggplot(data = osm_jub_sf) +
-  geom_sf(mapping = aes(color = crimes), size = 2) +
-  labs(title = "Open Street Map") +
-  scale_color_viridis_c(lim = c(0,42))
-
-p2 <- ggplot(data = tfl_jub_sf) +
-  geom_sf(mapping = aes(color = crimes), size = 2) +
-  labs(title = "Transport for London") +
-  scale_color_viridis_c(lim = c(0,42))
-
-p12 <- p1 / p2
-
-ggsave(filename = "img/maps.png", height = 16, width = 24, unit = "cm")
-```
-
-```{r crimemapcompare, eval = F}
-p1 <- ggplot(data = osm_jub_sf) +
-  geom_sf(mapping = aes(color = crimes), size = 2) +
-  labs(title = "Open Street Map") +
-  scale_color_viridis_c(lim = c(0,42))
-
-p2 <- ggplot(data = tfl_jub_sf) +
-  geom_sf(mapping = aes(color = crimes), size = 2) +
-  labs(title = "Transport for London") +
-  scale_color_viridis_c(lim = c(0,42))
-
-p1 / p2
-```
-
-```{r mapfig, out.width = "\\textwidth", echo = F}
-knitr::include_graphics("img/maps.png")
-```
-
-A number of factors emerge from this investigation.
-First, there is a disparity between the total number of crimes captured by the buffers in each dataset.
-Using the Open Street Map data, a total of `r sum(osm_jub_sf$crimes)` crimes were recorded across all stations on the line.
-Using the TfL data, only `r sum(tfl_jub_sf$crimes)` were recorded.
-Thus, the volume of crime observed in and around Jubilee stations during January will differ depending on the open data source being used.
-This could have important implications for how community and policing resources are distributed between underground lines.
-Second, the spatial patterning of crime hotspots on the line differs considerably between the two data sources.
-Using the Open Street Map data, the final station on the line, Stratford, has the greatest concentration of crimes, whereas the TfL data suggests that the most problematic station is Green Park, in the city centre.
-Despite the minor differences in the locations of each station, the conclusions drawn are starkly different, and would determine the consistency of findings with existing research examining crime at underground stations [@ceccato2013security].
-
-We can explore Stratford station as a potential crime hotspot using some of the skills we have learnt so far.
-One useful first step might to visualize the built environment around the station, along with the crime buffers from Open Street Map and TfL respectively, and the crime locations themselves.
-This way, we might find out why the TfL buffer is failing to capture any crimes, while the Open Street Map buffer is capturing an apparent crime hotspot.
-In doing so, it is worth remembering that crime locations in open police data are approximations of the real geography in order to ensure anonymity.
-
-```{r, eval = F}
-# Retrieve Stratford buffer from Open Street Map.
-stratford_osm_sf <- osm_jub_sf %>% 
-  filter(name == "Stratford")
-
-# Do the same using TfL.
-stratford_tfl_sf <- tfl_jub_sf %>% 
-  filter(commonName == "Stratford Underground Station")
-
-# Create a mini study area for our visualization, by getting the bounding box for the
-# Open Street Map (or TfL) buffer. Note that we transform to WGS 84 first in preperation
-# for the Open Street Map API.
-bb_sf <- stratford_osm_sf %>%
-  st_transform(crs = 4326) %>% 
-  st_bbox()
-
-# Retrieve the coordinates of the bounding box, and create a numeric list for use with
-# osmdata package.
-strat_bb <- c(bb_sf[[1]], bb_sf[[2]], bb_sf[[3]], bb_sf[[4]])
-
-# Pull building footprints in this bounding box from the Open Street Map API.
-strat_fp_sf <- opq(strat_bb) %>%
-    add_osm_feature(key = 'building') %>% 
-    osmdata_sf()
-
-# Get the polygons only.
-strat_fp_poly_sf <- strat_fp_sf$osm_polygons
-
-# Transform back to British National Grid to match our other data.
-strat_fp_poly_sf <- st_transform(strat_fp_poly_sf, 27700)
-
-# Clip the crime data to the Open Street Map buffer. Remember that the TfL buffer
-# did not capture any crimes.
-strat_osm_btp_sf <- st_intersection(stratford_osm_sf, btp_sf)
-
-# Visualize the building footprints, Open Street Map buffer, TfL buffer, and crime
-# locations. Note that the single point for the crimes is due to the approximation!
-# There are actually multiple crimes.
-ggplot() +
-  theme_bw() +
-  geom_sf(data = strat_fp_poly_sf, fill = "black") +
-  geom_sf(data = stratford_osm_sf, fill = "pink", alpha = 0.3) +
-  geom_sf(data = stratford_tfl_sf, fill = "blue", alpha = 0.3) +
-  geom_sf(data = strat_osm_btp_sf, col = "red") 
-```
-
-```{r strat, echo = F}
-knitr::include_graphics("img/stratford.png")
-```
-
-This quick check has shed some light on what is happening.
-The buffer generated from the TfL station location (in blue) covers a different part of the station to the Open Street Map buffer (in pink).
-The latter covers most of the main station building, while the former centers around the platforms to the north.
-Because the British Transport Police crime locations have been geocoded to a location in the main station building (in red), the TfL buffer has missed them.
-If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station.
-In many cases, there will not be an obvious 'right or wrong' answer to determine which open data source is most appropriate, but investigations like this one will help inform the decision.
-We might consider investigating other stations before deciding on which data to use, and ensure that findings are reported transparently with reference to these discussions.
-
-Engaging with the creators of the data would also be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
-For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations.
-For Open Street Map, its online transparency means that we could find out which user(s) recorded the Jubilee station locations, and directly contact this community of volunteers to find out how station locations were recorded and why.
+With this in mind, we encourage readers to complete the practical exercise in which we will utilize multiple sources of open data to explore police-recorded crime on and around public transport in London, England.
 
@@
 Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources.
-By way of an example, we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
+The suppmentary material contained a practical exercise in which we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
 Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line.

### 2020-11-18 reference_doc.Rmd
revisions v1 no tiffs

---
title: "Reference document"
author: "Samuel Langton"
date: "18/11/2020"
output: word_document
bibliography: refs.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document [@kitchin2014data]. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### One more title

Here is some text.

### 2020-11-18 supplementary_materials.Rmd
revisions v1 no tiffs

--- 2020-11-16 draft_chapter.Rmd
@@
 title: "Open Data for _Crime and Place_ Research: A Practical Guide in R."
+subtitle: "Supplementary materials"
 author:
-  - Samuel Langton\dag
-  - Reka Solymosi\ddag
+  - Samuel Langton (University of Leeds)
+  - Reka Solymosi (University of Manchester)
 editor_options:
@@
 knitr::opts_chunk$set(message = F, warning = F, eval = T, echo = T)
-# library(jsonlite)
-# library(osmdata)
-# library(ggplot2)
-# library(sf)
-# library(readr)
-# library(tidyr)
-# library(patchwork)
-# library(dplyr)
 ```
-
-\enspace
-
-This is an unedited draft, not for citation, of a book chapter that has been accepted for publication in E.
-Groff and C.
-Haberman (Eds.), The study of crime and place: A methods handbook (Temple University Press, 2021).
-Temple University Press retains all rights and subrights of this document.
-This pre-print is not published under a CC license.
-
-\enspace
-
-**Keywords:** crime, open data, Open Street Map, transport, geography.
-
-**Corresponding author:** Samuel Langton, University of Leeds, School of Law, Leeds, United Kingdom. Email: s.langton@leeds.ac.uk.
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-\enspace
-
-***
-
-\dag School of Law, University of Leeds, Leeds, United Kingdom. ORCID: 0000-0002-1322-1553.
-
-\ddag School of Social Sciences, University of Manchester, Manchester, United Kingdom. ORCID: 0000-0001-8689-1526.
-
-\newpage
-
-# Introduction
-
-Access to data in crime and place research has traditionally been reserved for those who have the means to collect fresh data themselves, pay for access, or obtain data through formal data sharing agreements.
-Even when access is granted, the usage of these data often comes with conditions that circumscribe how the data can be used through licensing or policy [@kitchin2014data].
-Even the public dissemination of findings which emerge from analysis might be subject to restrictions.
-This can lead to unequal access, controlled usage and curb the diffusion of findings, severely limiting the insight that can be obtained from data.
-
-Open data initiatives provide a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements, and negotiations [@manovich2011trending].
-Open data can lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
-As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyse open datasets.
-
-This chapter will outline how to develop such skills by providing a framework to approach and meaningfully interpret open data.
-The chapter also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.
-
-# Background
-
-## What is open data?
-
-First, it is useful to clarify what we mean when we refer to 'open data'.
-The _Open Data Handbook_, compiled by the Open Knowledge Foundation, states that open data are "data that can be freely used, re-used and redistributed by anyone - subject only, at most, to the requirement to attribute and sharealike" [@dietrich2009open, para.
-3].
-Specifically, open data can be defined along three key domains [@knowledge2016open].
-
-- **Availability and Access:** the data must be accessible via a public domain, at no more than a reasonable reproduction cost, and through a convenient medium, such as the internet.
-Ideally, it should be machine-readable and modifiable.
-For example, it should be downloadable from a website as an unlocked spreadsheet, or JSON file, rather than presented as a summary table in a PDF document.
-- **Re-use and Redistribution:** the data must be provided under conditions that allow free use, such as modification, separation, or compilation of the data, and permit re-use and redistribution including the intermixing (merging) with other datasets.
-- **Universal Participation:** everyone should be allowed to use, re-use and redistribute the data and its derivatives without restriction.
-For instance, restrictions that only allow non-commercial or educational usage would *not* constitute open data.
-
-In short, 'open data' should be data that are readily and publicly available, in a usable format, and without restrictions on usage, modification and re-distribution.
-
-Open data practices, and the extent to which it is a reality, vary both within and between countries.
-For example, the United States has a history of making public sector datasets openly available.
-The United Kingdom can be more strict, releasing data under a licence ('Open Government Licence'), with some data requiring a fee [@kitchin2014data].
-The _Global Open Data Index_ [@globopendata] is one tool for tracking the annual global benchmark for publication of open government data by country.
-It provides a rank which indicates how well governments across the world follow open data practices across various domains.
-Many of these have direct relevance to crime and place research, such as administrative boundaries and geographic locations (e.g.
-coordinates).
-
-The drive towards transparency and scrutiny of public information is further enhanced by an increasing need for replication - a hallmark for open science research practice.
-Opening up datasets used in criminology publications and wider social sciences fosters and facilitates a culture of replication [@pridemore2018replication].
-As such, open data fits into the wider discourse around transparency, open review and open scrutiny, which (in time) may even become a policy requirement irrespective of the research field or the source of funding [@vuong2017open].
-Increasingly, researchers are making use of online repositories for their papers (e.g.
-https://arxiv.org/), data (e.g.
-https://www.ukdataservice.ac.uk/deposit-data), code (e.g.
-https://github.com/), or whole projects (e.g.
-https://osf.io/) all in the name of transparency and replication.
-These are guidelines and tools worth considering, not just when conducting analysis on secondary data sources, but also collecting and sharing your own data.
-
-## What are types of open data?
-
-Now we have a reasonable idea about what open data is in a broad sense, we can consider the different types of open data which you might come across.
-Here, we formulate a typology of open datasets based on their origin.
-
-### Public sector
-
-The 'public sector' refers to organizations owned and operated by local or central government with the core aim of providing services to the public.
-Much of the open data movement has focused on opening up information generated by these local and national state agencies [@kitchin2014data].
-These can include national statistics or administrative data, but also data collected by publicly funded research projects such as victimization and household surveys, amongst others.
-The _Global Open Data Index_, introduced earlier, focuses on identifying the gaps in governmental organizations, encouraging them to think about how public sector information can become more usable, and ultimately, more impactful [@globopendata].
-
-Specific to crime and place research, datasets of interest in this particular domain might include data describing urban structure, such as street networks, to assess whether the configuration of roads dictates things like violent crime victimization [@streetsyntax2017], or victimisation surveys to quantify citizens' perceived safety and security.
-
-### Private sector
-
-In contrast to the public sector, 'private sector' tends to refer to parts of the economy which are not under direct state control, and often operate for profit.
-Opening up data generated by the private sector represents a significant challenge, largely due to the proprietary value to its creators [@kitchin2014data].
-The aims and objectives of private companies, who have a duty to shareholders and operate in a competitive market environment, differ considerably from local and central government.
-
-That said, some datasets are released openly by private sector organizations, albeit often only a subset of what would be available as a paying customer.
-The website for ArcGIS, a piece of software maintained by the Environmental Systems Research Institute (ESRI), a private company, host a number of open geographic datasets on their online cloud platform (https://hub.arcgis.com/search).
-Indeed, many papers have made use of data collected or distributed by private organizations to explore crime and place, such as Google Street View images [@langton2017residential] and Twitter [@malleson2015impact].
-In fact, you will learn how to obtain free Twitter data for such purposes in this book (see Chapter 6, Topic 7).
-
-### Open crowdsourced data
-
-Finally, there is a specific group of open data sources that collate data collectively, generated by large groups of individuals who do not specifically belong to any organization, but instead work together on a collaborative project.
-These are crowdsourced datasets.
-Examples include Wikipedia (https://www.wikipedia.org/), an online encyclopedia where anyone can contribute, or Flickr (https://www.flickr.com/) an online photo gallery where people can upload and tag their photos.
-Often, the organizations who maintain and monitor these data collection activities are charities or non-governmental organizations that do not operate for profit, but instead provide some form of social good.
-An example is the online reporting platform FixMyStreet (https://www.fixmystreet.com/), where people can report problems such as instances of graffiti, vandalism or environmental issues.
-This data has been used to explore signal crimes theory, for instance, offering insight into people's experiences with incivilities [@solymosi2018crowdsourcing].
-Another topic in this book (see Chapter 6, Topic 6) discusses the merits and pitfalls of crowdsourced data, and what to watch out for when analyzing open data of this type.
-
-## Strengths and limitations
-
-The defining characteristics of open data, namely, that of availability, re-usability and universal participation, outlined earlier, represent its greatest strength.
-But besides from this, there are other advantages.
-One specific motivation for using open data comes from its potential to address many of the limitations associated with traditional surveying methods, such as social desirability bias, or issues associated with memory and recall [@mayer2013big].
-This advantage exists largely due to the organic way in which open data is generated.
-It is often a by-product of other activities, and as such, we can gain an honest insight into people's everyday lives and associated social processes [@solymosi2018role].
-As noted, open data also means open research, facilitating transparency and reproducibility.
-That said, there are a number of shortcomings which are worthy of consideration.
-
-Firstly, one of the biggest obstacles to collecting open data is the ability to interpret what the data means (e.g.
-applying a framework for its analysis), and computational skills in data scraping, wrangling and cleaning, in order to transform it into a usable format for research [@boyd2012critical].
-All sorts of open data remain inaccessible to people who may lack the skills and know-how to acquire them.
-Although this is certainly an obstacle for data usage more generally, it represents a key challenge to governmental bodies, in particular, on which there is an onus to ensure transparency and facilitate scrutiny.
-Private companies, especially those who generate open data as a by-product of their primary activity (e.g.
-Twitter), have little responsibility or pressure to ensure that their data is accessible and usable to the general public.
-Moreover, in both public and private spheres, licences and conditions on open data are not necessarily concrete, and might be subject to change with little or no notice.
-This is an important consideration when planning and running long-term research projects which involve open data.
-
-Secondly, despite the merits of open data in terms of remedying pitfalls in traditional survey methods, there are other threats to validity that may emerge.
-For instance, whilst many online resources offer an 'honest' depiction of society, uncaptured by surveys, researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
-This could be a result of individuals willingly managing and 'curating' their online presence [@ellison2006managing], or because of wider issues such as government censorship, or cultural norms around particular topics, particularly sensitive topics of interest to researchers of crime, such as sexual assault or drug use.
-
-Relatedly, researchers should be aware of issues over sampling, and ultimately, the generalizability of findings that emerge from the analysis of open data.
-In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduces a degree of bias [@longley2012geodemographics].
-Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
-Contributions to resources such as Open Street Map (which we look at in the practical exercise later) are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
-These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
-
-### What can be done?
-
-Open data can be messy, biased and noisey, but criminology (and social sciences more generally) can benefit immeasurably from its use.
-Only through open data can public sector bodies be held to account, research be transparent and reproducible, and participation in data analysis universal.
-In crime and place research, both dependent variables (e.g.
-police-recorded crime incidents, victimization rates) and independent variables (e.g.
-demographic characteristics, ambient population estimates) can be sourced from open data, whether public, private or crowdsourced.
-
-So, what can we do to make sure we make good use of these data?
-With a critical, engaged and considered approach to conducting research with open data, criminology can become a leading force in open and reproducible social science.
-In sum, researchers and analysts using open data must ask critical questions:
-
-- **Where** does this data come from? 
-- **Why** was the data collected in the first place?
-- **Who** is represented in the data, and who is excluded?
-- **What** concepts and constructs can and cannot be operationalized with this data?
-- **When** did data collection take place, and how might have that influenced results?
-
-The answers to these questions will put the data sources in context, and ensure that researchers use them appropriately and with due care.
-
-With this in mind, we now move on to a practical exercise in which we will utilize multiple sources of open data to explore police-recorded crime on and around public transport in London, England.
 
@@
 
-> Location anonymization.
-The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened.
+_Location anonymization._
+_The latitude and longitude locations of Crime and ASB incidents published on this site always represent the approximate location of a crime — not the exact place that it happened._
 
@@
 
-```{r tfl, out.width = "\\textwidth", fig.cap="example of data from Transport for London", echo = F}
-knitr::include_graphics("img/tfl_screenshot.png")
-```
+[INSERT FIGURE 1 HERE]
 
@@
 
-bb_sf <- getbb(place_name = "greater london united kingdom", format_out = "sf_polygon")
+bb_sf <- getbb(place_name = "greater london united kingdom",
+               format_out = "sf_polygon")
 ```
@@
 Here, we specify our bounding box as the manual coordinates contained in `bb_gl`.
-^[If you use the bounding box obtained through `getbb()` one can subsequently trim down the outputs from `add_osm_feature()` using `trim_osmdata()`.
-For instance, we could add `trim_osmdata(bb_poly = bb_sf)` to our initial query.]
 
 ```{r opq}
-osm_stat_sf <- opq(bbox = bb_gl) %>%                               # select bounding box
+osm_stat_sf <- opq(bbox = bb_gl) %>%                               # bounding box
   add_osm_feature(key = 'public_transport', value = 'station') %>% # select features
@@
 
-We can now easily visualize this data using the `ggplot2` package, which is compatible with `sf` objects.
+We can now easily visualize this data (see Figure 2) using the `ggplot2` package, which is compatible with `sf` objects.
 
-```{r tflmap}
+```{r tflmap, eval = F}
 ggplot(tfl_jub_sf) +
@@
 ```
+
+[INSERT FIGURE 2 HERE]
 
@@
 
-We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two datasets compare.
+We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two datasets compare (see Figure 3).
 
-```{r mapline}
+```{r mapline, eval = F}
 ggplot() +
@@
 ```
+
+[INSERT FIGURE 3 HERE]
 
@@
 
-We can then visualize these counts and compare the two sources of data by coloring in our buffers according to the crime count.
+We can then visualize these counts and compare the two sources of data by coloring in our buffers according to the crime count (see Figure 4).
 Note that we arrange the plots using syntax available using the `patchwork` package, and set the scales to be comparable using the `lim` argument within `scale_color_viridis_c()`.
-
-```{r crimemapcomparesave, echo = F}
-p1 <- ggplot(data = osm_jub_sf) +
-  geom_sf(mapping = aes(color = crimes), size = 2) +
-  labs(title = "Open Street Map") +
-  scale_color_viridis_c(lim = c(0,42))
-
-p2 <- ggplot(data = tfl_jub_sf) +
-  geom_sf(mapping = aes(color = crimes), size = 2) +
-  labs(title = "Transport for London") +
-  scale_color_viridis_c(lim = c(0,42))
-
-p12 <- p1 / p2
-
-ggsave(filename = "img/maps.png", height = 16, width = 24, unit = "cm")
-```
 
@@
 
-```{r mapfig, out.width = "\\textwidth", echo = F}
-knitr::include_graphics("img/maps.png")
-```
+[INSERT FIGURE 4 HERE]
 
@@
 We can explore Stratford station as a potential crime hotspot using some of the skills we have learnt so far.
-One useful first step might to visualize the built environment around the station, along with the crime buffers from Open Street Map and TfL respectively, and the crime locations themselves.
+One useful first step might to visualize the built environment around the station, along with the crime buffers from Open Street Map and TfL respectively, and the crime locations themselves (see Figure 5).
 This way, we might find out why the TfL buffer is failing to capture any crimes, while the Open Street Map buffer is capturing an apparent crime hotspot.
@@
 
-```{r strat, echo = F}
-knitr::include_graphics("img/stratford.png")
-```
+[INSERT FIGURE 5 HERE]
 
@@
 
-# Discussion
+\newpage
 
-## Summary
-
-This chapter has sought to introduce open data as a novel, growing and invaluable tool in crime and place research through a review of key fundamentals and a substantive demonstration within R.
-We have defined open data, and outlined some key types of open information available, along with their respective strengths and weaknesses.
-An important component of this review has been to encourage an engaged and critical research approach to using open data.
-Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources.
-By way of an example, we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
-Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line.
-The demonstration highlighted both the power of open data and the need to critically engage with sources.
-
-## Future of open data
-
-As it stands, open data represents a key resource in crime and place research.
-It is a fundamental component in the movement towards transparent and reproducible scientific research.
-The skills required to effectively access and use open data, such as those demonstrated in this chapter, are increasingly being taught and deployed in social science.
-The number and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology.
-For instance, information about building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
-But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go.
-As with any emerging domain, open data sources vary considerably in their degree of completeness and reliability.
-Whilst this can offer insight, for instance, in terms of public participation and sense of community, it can also represent a significant obstacle to empirical analysis.
-A key challenge for the future will be to ensure that students and research practitioners ask critical questions of their data before blindly delving into analysis and public dissemination.
-Moreover, the longevity of open data licences remains an unknown parameter.
-With so much open data being distributed by the public sector, the accessibility of open information is subject to fluctuations in the amount of resource (and goodwill) available to safe-guard continuity.
-One way of justifying the continued collection and distribution of open public sector data is simply to use the data for public good, such as informing interventions which help reduce crime victimization.
-
-\newpage
+# Learn more about it
 

### 2020-11-18 scripts/osm_explore_tube.R

@@ -45,8 +45,11 @@ osm_jub_sf <- osm_stat_sf %>%
 osm_jub_sf <- st_transform(osm_jub_sf, 27700)
 # bb_sf      <- st_transform(bb_sf, 27700)
 
-ggplot() +
+plot_1 <- ggplot() +
   geom_sf(data = osm_jub_sf, size = 0.8)
+
+ggsave(plot = plot_1, filename = "img/fig2.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
+# ggsave(plot = plot_1, filename = "img/fig1.eps", device = "eps", dpi = 1200)
 
 # TfL scrape ==============================================================================================
 
@@ -60,9 +63,11 @@ tfl_jub_sf <- api_call %>%
   st_transform(27700)
 
 # Plot difference.
-ggplot() +
+plot_2 <- ggplot() +
   geom_sf(data = osm_jub_sf) +
   geom_sf(data = tfl_jub_sf, color = "red", alpha = 0.5) 
+
+ggsave(plot = plot_2, filename = "img/fig3.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
 
 # Save before someone changes the lines.
 # temp_osm_jub_sf <- osm_jub_sf %>% 
@@ -103,8 +108,9 @@ p2 <- ggplot(data = tfl_jub_sf) +
   labs(title = "Transport for London") +
   scale_colour_viridis_c(lim = c(0,42))
 
-p1 / p2
+plot_3 <- p1 / p2
 
+ggsave(plot = plot_3, filename = "img/fig4.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
 
 sum(osm_jub_sf$crimes) # 168
 sum(tfl_jub_sf$crimes) # 136
@@ -154,16 +160,15 @@ strat_osm_btp_sf <- st_intersection(stratford_osm_sf, btp_sf)
 strat_fp_poly_sf <- st_transform(strat_fp_poly_sf, 27700)
 
 # Quick plot.
-strat_gg <- ggplot() +
+plot_4 <- ggplot() +
   geom_sf(data = strat_fp_poly_sf, fill = "black") +
   geom_sf(data = stratford_osm_sf, fill = "#440154FF", alpha = 0.3) +
   geom_sf(data = stratford_tfl_sf, fill = "#FDE725FF", alpha = 0.3) +
   geom_sf(data = strat_osm_btp_sf, col = "#3CBC75FF") +
[…]

### 2020-11-18 supplementary_materials.Rmd
learn more about it section

@@
 
-# Learn more about it
-
 # Software
@@
 Packages _readr_ [@readrpack] and _jsonlite_ [@jsonlitepack] were used for reading and parsing data, _osmdata_ for querying the Open Street Map API [@osmdatapack], _sf_ for handling and visualizing spatial data [@sfpack], _dplyr_ and _tidyr_ for wrangling data [@dplyrpack; @tidyrpack], _ggplot2_ for visualization [@ggplotpack] and _patchwork_ for arranging plots [@patchworkpack].
-The document itself was prepared using _rmarkdown_ [@rmarkdownpack].
+The draft versions of the book chapter and supplementary material were prepared using _rmarkdown_ [@rmarkdownpack].
+
+# Learn more about it
+
+A useful resource going forward is the _Open Data Handbook_ which can be found online for free (https://opendatahandbook.org/).
+It provides introductory information (e.g.
+definitions), advice (e.g.
+legal, technical details), case studies which showcase the societal value of open data, and an e-library of papers, presentations and videos on open data.
+It is one of many projects organized by the _Open Knowledge Foundation_.
+Another comprehensive and free online resource is the _Open Data Institute_ website (https://theodi.org/).
+Their mission is to demonstrate the positive value of open data and advocate for its ethical creation and usage by working with companies and governments around the globe.
+For resources geared towards R, _rOpenSci_ (https://ropensci.org/) develop packages and offer an infrastructure and community to promote the accessible usage of R for open scientific research.
+The official _RStudio_ webpage (https://rstudio.com/about/) also provides a wealth of information and freely available resources for making the most of the software.
 

### 2020-11-18 manuscript.Rmd
revisions update

@@
 
-The drive towards transparency and scrutiny of public information is further enhanced by an increasing need for replication - a hallmark for open science research practice.
+The drive towards transparency and scrutiny of public information is further enhanced by an increasing need for replication.
+This is a key component of open science: transparent, reusable, reproducible and accessible research practices.
 Opening up datasets used in criminology publications and wider social sciences fosters and facilitates a culture of replication [@pridemore2018replication].
@@
 Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources.
-The suppmentary material contained a practical exercise in which we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
+The supplementary material contained a practical exercise in which we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper.
 Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line.

### 2020-11-18 supplementary_materials.Rmd
revisions update

@@
 For this exercise, we are going to use British Transport Police data, a force which operates on railways and light-rail systems across the country, for the month of January in 2020.
-Select the time period using the dropdown menus, the force using the tickboxes, and then generate and download the file.
+Select the time period using the dropdown menus, the force using the tickboxes (see Figure 1), and then generate and download the file.
+
+[INSERT FIGURE 1 HERE]
 
@@
 Copy their demonstration URL for the bakerloo line (https://api.tfl.gov.uk/line/bakerloo/stoppoints) and paste it into your web browser.
-When you visit this page, you should see something like Figure 1.
+When you visit this page, you should see something like Figure 2.
 
-[INSERT FIGURE 1 HERE]
+[INSERT FIGURE 2 HERE]
 
@@
 
-We can now easily visualize this data (see Figure 2) using the `ggplot2` package, which is compatible with `sf` objects.
+We can now easily visualize this data (see Figure 3) using the `ggplot2` package, which is compatible with `sf` objects.
 
@@
 
-[INSERT FIGURE 2 HERE]
+[INSERT FIGURE 3 HERE]
 
@@
 
-We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two datasets compare (see Figure 3).
+We can now plot the Open Street Map station points, which are already spatial, over these station locations pulled from the TfL API, to see how the two datasets compare (see Figure 4).
 
@@
 
-[INSERT FIGURE 3 HERE]
+[INSERT FIGURE 4 HERE]
 
@@
 
-We can then visualize these counts and compare the two sources of data by coloring in our buffers according to the crime count (see Figure 4).
+We can then visualize these counts and compare the two sources of data by coloring in our buffers according to the crime count (see Figure 5).
 Note that we arrange the plots using syntax available using the `patchwork` package, and set the scales to be comparable using the `lim` argument within `scale_color_viridis_c()`.
@@
 
-[INSERT FIGURE 4 HERE]
+[INSERT FIGURE 5 HERE]
 
@@
 We can explore Stratford station as a potential crime hotspot using some of the skills we have learnt so far.
-One useful first step might to visualize the built environment around the station, along with the crime buffers from Open Street Map and TfL respectively, and the crime locations themselves (see Figure 5).
+One useful first step might to visualize the built environment around the station, along with the crime buffers from Open Street Map and TfL respectively, and the crime locations themselves (see Figure 6).
 This way, we might find out why the TfL buffer is failing to capture any crimes, while the Open Street Map buffer is capturing an apparent crime hotspot.
@@
 
-[INSERT FIGURE 5 HERE]
+[INSERT FIGURE 6 HERE]
 

### 2020-11-19 manuscript.Rmd
revisions update scholar

@@
 
-Secondly, despite the merits of open data in terms of remedying pitfalls in traditional survey methods, there are other threats to validity that may emerge.
-For instance, whilst many online resources offer an 'honest' depiction of society, uncaptured by surveys, researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
-This could be a result of individuals willingly managing and 'curating' their online presence [@ellison2006managing], or because of wider issues such as government censorship, or cultural norms around particular topics, particularly sensitive topics of interest to researchers of crime, such as sexual assault or drug use.
+Secondly, the validity and reliability of open data sources can be questioned, and some cases, difficult to verify.
+For instance, Open Street Map (which we look at in the practical exercise) contains a vast array of data about its features, from whether a bar is LGBTQ+ friendly, to whether a train station is wheelchair accessible.
+Often, there is no method for cross-validating this information with other data sources.
+Similarly, whilst many open resources offer a depiction of society uncaptured by traditional surveys (e.g.
+social media) researchers should be careful in interpreting people's communication online as completely authentic [@manovich2011trending].
+This could be a result of individuals willingly managing and 'curating' their online presence [@ellison2006managing], or because of wider issues such as government censorship, or cultural norms around certain topics, particularly sensitive topics of interest to researchers of crime, such as sexual assault or drug use.
 
-Relatedly, researchers should be aware of issues over sampling, and ultimately, the generalizability of findings that emerge from the analysis of open data.
+Relatedly, researchers should be aware of issues over sampling and selection, and ultimately, the generalizability of findings that emerge from the analysis of open data.
 In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduces a degree of bias [@longley2012geodemographics].
 Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good].
-Contributions to resources such as Open Street Map (which we look at in the practical exercise) are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
+Contributions to resources such as Open Street Map are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
 These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.

### 2020-11-19 supplementary_materials.Rmd
revisions update scholar

@@
 
-```{r, eval = F}
+```{r stratford, eval = F}
 # Retrieve Stratford buffer from Open Street Map.
@@
 # There are actually multiple crimes. Colors have been manually selected
-# from the viridis package to be easier to read by those with colorblindness.
+# from the viridis package simply to identify each object individually.
 ggplot() +
@@
 If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station.
+This also serves to highlight the limitations of the open crime data, because we cannot discern the _exact_ location of these crimes in the station.
+It is in fact plausible that some of these crimes did occur in the TfL buffer, but due to the process of anonymization, the points have been snapped the main station building.
+
 In many cases, there will not be an obvious 'right or wrong' answer to determine which open data source is most appropriate, but investigations like this one will help inform the decision.

### 2020-11-19 manuscript.Rmd
revisions update

@@
 Open data can lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role].
-As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyse open datasets.
+As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyze open datasets.
 
 This chapter will outline how to develop such skills by providing a framework to approach and meaningfully interpret open data.
-The supplementary material also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.
+The supplementary material also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyze different sources of open data in order to draw conclusions about crime and place.
 
@@
 Contributions to resources such as Open Street Map are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting].
-These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
+These issues are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.
 
@@
 - **What** concepts and constructs can and cannot be operationalized with this data?
-- **When** did data collection take place, and how might have that influenced results?
+- **When** did data collection take place, and how might that have influenced results?
 
@@
 
-With this in mind, we encourage readers to complete the practical exercise in which we will utilize multiple sources of open data to explore police-recorded crime on and around public transport in London, England.
+With this in mind, we encourage readers to complete the practical exercise in which we utilize multiple sources of open data to explore police-recorded crime in and around public transport stations in London, England.
 
@@
 
-This chapter has sought to introduce open data as a novel, growing and invaluable tool in crime and place research through a review of key fundamentals and a substantive demonstration within R.
+This chapter has sought to introduce open data as a novel, growing and invaluable tool in crime and place research through a review of key fundamentals.
+The corresponding supplementary material also includes a substantive demonstration using R.
 We have defined open data, and outlined some key types of open information available, along with their respective strengths and weaknesses.
@@
 It is a fundamental component in the movement towards transparent and reproducible scientific research.
-The skills required to effectively access and use open data, such as those demonstrated in this chapter, are increasingly being taught and deployed in social science.
-The number and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology.
-For instance, information about building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time.
+The skills required to effectively access and use open data, such as those demonstrated in the practical exercise, are increasingly being taught and deployed in social science.
+The scale and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology.
+Open Street Map represents an ever-growing database to extract theoretically relevant information on the built environment (e.g.
+land-use, space syntax) and the opportunity structures of micro-places (e.g.
+locations of bars, gas stations) to study crime and place.
+New resources are also emerging.
+For instance, information about historical building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime in the capital.
 But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go.
-As with any emerging domain, open data sources vary considerably in their degree of completeness and reliability.
+As with many domains, open data sources vary considerably in their degree of completeness, validity and reliability.
 Whilst this can offer insight, for instance, in terms of public participation and sense of community, it can also represent a significant obstacle to empirical analysis.

### 2020-11-19 supplementary_materials.Rmd
revisions update

@@
 
-Open Street Map has two main types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.
+Two main types of wrappers available for the Open Street Map API are a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/) and an R package called `osmdata`.
 
@@
 
-The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the **Console**.
+The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the R Console.
 
@@
 ggplot() +
-  geom_sf(data = tfl_jub_sf) +
-  geom_sf(data = osm_jub_sf, col = "red", alpha = 0.5) 
+  geom_sf(data = osm_jub_sf) +
+  geom_sf(data = tfl_jub_sf, color = "red", alpha = 0.5)  
 ```
@@
 # Create a mini study area for our visualization, by getting the bounding box for the
-# Open Street Map (or TfL) buffer. Note that we transform to WGS 84 first in preperation
-# for the Open Street Map API.
+# Open Street Map (or TfL) buffer. Note that we transform to WGS 84 first in
+# preperation for the Open Street Map API.
 bb_sf <- stratford_osm_sf %>%
@@
 
-# Retrieve the coordinates of the bounding box, and create a numeric list for use with
-# osmdata package.
+# Retrieve the coordinates of the bounding box, and create a numeric list for use
+# with osmdata package.
 strat_bb <- c(bb_sf[[1]], bb_sf[[2]], bb_sf[[3]], bb_sf[[4]])
@@
 
-Engaging with the creators of the data would also be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter.
+Engaging with the creators of the data would also be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed in the book chapter.
 For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations.
@@
 legal, technical details), case studies which showcase the societal value of open data, and an e-library of papers, presentations and videos on open data.
-It is one of many projects organized by the _Open Knowledge Foundation_.
+It is one of many projects organized by the Open Knowledge Foundation.
 Another comprehensive and free online resource is the _Open Data Institute_ website (https://theodi.org/).

### 2020-11-20 scripts/osm_explore_tube.R
revisions as submitted

@@ -45,10 +45,10 @@ osm_jub_sf <- osm_stat_sf %>%
 osm_jub_sf <- st_transform(osm_jub_sf, 27700)
 # bb_sf      <- st_transform(bb_sf, 27700)
 
-plot_1 <- ggplot() +
-  geom_sf(data = osm_jub_sf, size = 0.8)
+# plot_1 <- ggplot() +
+#   geom_sf(data = osm_jub_sf, size = 0.8)
 
-ggsave(plot = plot_1, filename = "img/fig2.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
+# ggsave(plot = plot_1, filename = "img/fig2.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
 # ggsave(plot = plot_1, filename = "img/fig1.eps", device = "eps", dpi = 1200)
 
 # TfL scrape ==============================================================================================
@@ -62,12 +62,22 @@ tfl_jub_sf <- api_call %>%
   st_as_sf(coords = c(x = "lon", y = "lat"), crs = 4326) %>% 
   st_transform(27700)
 
-# Plot difference.
-plot_2 <- ggplot() +
-  geom_sf(data = osm_jub_sf) +
-  geom_sf(data = tfl_jub_sf, color = "red", alpha = 0.5) 
+# Plot TfL.
+fig3 <- ggplot() +
+  geom_sf(data = tfl_jub_sf)
 
-ggsave(plot = plot_2, filename = "img/fig3.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
+ggsave(plot = fig3, filename = "img/fig3.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
+
+
+
+
+# Plot difference.
+fig4 <- ggplot() +
+  geom_sf(data = tfl_jub_sf) +
+  geom_sf(data = osm_jub_sf, color = "red", alpha = 0.5) 
+fig4
+
+ggsave(plot = fig4, filename = "img/fig4.tiff", device = "tiff", dpi = 1200, height = 7, width = 7)
 
 # Save before someone changes the lines.
 # temp_osm_jub_sf <- osm_jub_sf %>% 
@@ -108,9 +118,9 @@ p2 <- ggplot(data = tfl_jub_sf) +
   labs(title = "Transport for London") +
   scale_colour_viridis_c(lim = c(0,42))
 
-plot_3 <- p1 / p2
[…]
